---
title: "MIMIC-IV Data Profiler"
format: html
editor: visual
params:
  db_config: "databases/mimic_iv.yaml"
---

# MIMIC-IV Data Profiler

Profiler for the [MIMIC-IV critical care
dataset](https://physionet.org/content/mimiciv/) loaded into DuckDB.
Produces a schema dump and a PHI-safe data profile at the paths declared in
the YAML config pointed to by the `db_config` param.

**SQL dialect:** DuckDB (`LIMIT`, `EXTRACT(year FROM …)`,
`strftime(date,'%Y-%m')`).

**MIMIC-IV de-id notes baked in:**
- `anchor_age` is already top-coded at 91 for ages > 89 — we still emit
  90+ buckets to keep parity with the other CDM profilers.
- Calendar dates are shifted by per-patient random offsets; intervals are
  preserved. Year-by-year breakdowns reflect shifted years, not real time.

```{r load-libraries}
library(tidyverse)
library(DBI)
library(duckdb)
library(yaml)
library(here)
options(encoding = "UTF-8")
```

## Load configuration and shared scaffolding

```{r load-config}
#| code-fold: true
source(here::here("profilers/_helpers.R"))

db_config_path <- if (exists("params") && !is.null(params$db_config)) {
  params$db_config
} else {
  "databases/mimic_iv.yaml"
}

config <- load_profiler_config(db_config_path)
cfg     <- config$cfg
paths   <- config$paths

# MIMIC-IV is typically loaded in DuckDB at the root namespace (no schema)
# OR under mimiciv_hosp / mimiciv_icu. Build a qualifier helper that handles
# both: if schema_prefix is empty, table names are bare; otherwise prefixed.
qualify <- function(tbl) {
  if (is.null(cfg$schema_prefix) || !nzchar(cfg$schema_prefix)) tbl
  else paste0(cfg$schema_prefix, ".", tbl)
}

suppress_threshold <- 11
sup <- make_suppress(suppress_threshold)
```

## Preflight

```{r preflight}
preflight_check(cfg$connection$r_code, label = cfg$id %||% "db")
```

## Schema extraction (DuckDB dialect)

```{r fn-extract-schema}
#| code-fold: true

extract_schema_mimic <- function(con, output_file = "schema.txt") {
  # DuckDB exposes information_schema. List both base tables and views.
  tables <- DBI::dbGetQuery(con, "
    SELECT table_schema, table_name, table_type
    FROM information_schema.tables
    WHERE table_schema NOT IN ('information_schema', 'pg_catalog')
    ORDER BY table_schema, table_name
  ")

  out <- c(
    paste0("# MIMIC-IV Schema (", cfg$id %||% "db", ")"),
    paste0("# Generated: ", Sys.time()),
    ""
  )

  current_schema <- ""
  for (i in seq_len(nrow(tables))) {
    row <- tables[i, ]
    if (row$table_schema != current_schema) {
      out <- c(out, paste0("## Schema: ", row$table_schema), "")
      current_schema <- row$table_schema
    }

    columns <- DBI::dbGetQuery(con, sprintf("
      SELECT column_name, data_type, is_nullable
      FROM information_schema.columns
      WHERE table_schema = '%s' AND table_name = '%s'
      ORDER BY ordinal_position
    ", row$table_schema, row$table_name))

    out <- c(out,
             paste0("### Table: ", row$table_name, " (", row$table_type, ")"),
             "")
    for (j in seq_len(nrow(columns))) {
      col <- columns[j, ]
      nullable <- ifelse(col$is_nullable == "YES", "NULL", "NOT NULL")
      out <- c(out, paste0("- ", col$column_name, ": ",
                           col$data_type, " (", nullable, ")"))
    }
    out <- c(out, "")
  }
  out <- c(out, "---", "## Summary",
           paste0("Total tables/views: ", nrow(tables)))
  writeLines(out, output_file)
  message("Schema written to: ", output_file)
}
```

## Data profile

```{r fn-profile-mimic}
#| code-fold: true

profile_mimic <- function(con,
                          output_file,
                          suppress_threshold = 11) {

  safe_query <- make_safe_query(con)
  sup <- make_suppress(suppress_threshold)

  out <- character()
  ln  <- function(...) out <<- c(out, paste0(...))

  ln("# MIMIC-IV Data Profile")
  ln("# Generated: ", Sys.time())
  ln("# This file contains ONLY aggregate counts and de-identified")
  ln("# distributions. Counts < ", suppress_threshold, " are suppressed.")
  ln("# Ages \u2265 90 are top-coded per HIPAA Safe Harbor.")
  ln("# (MIMIC-IV's `anchor_age` already top-codes >89 to 91.)")
  ln("")

  # ---------------------------------------------------------------
  # 1. Overall size
  # ---------------------------------------------------------------
  ln("---")
  ln("## 1. Overall Size")
  ln("")
  pts <- safe_query(sprintf("
    SELECT COUNT(DISTINCT subject_id) AS n FROM %s
  ", qualify("patients")), "patients")
  if (nrow(pts) > 0) ln("Distinct patients (hosp): ", sup(pts$n[1]))

  adm <- safe_query(sprintf("
    SELECT COUNT(DISTINCT subject_id) AS n_pat,
           COUNT(*) AS n_adm
    FROM %s
  ", qualify("admissions")), "admissions")
  if (nrow(adm) > 0) {
    ln("Distinct patients with admissions: ", sup(adm$n_pat[1]))
    ln("Total admissions: ", format(adm$n_adm[1], big.mark = ","))
  }

  icu <- safe_query(sprintf("
    SELECT COUNT(DISTINCT subject_id) AS n_pat,
           COUNT(*) AS n_stays
    FROM %s
  ", qualify("icustays")), "icustays")
  if (nrow(icu) > 0) {
    ln("Distinct patients with ICU stays: ", sup(icu$n_pat[1]))
    ln("Total ICU stays: ", format(icu$n_stays[1], big.mark = ","))
  }
  ln("")

  # ---------------------------------------------------------------
  # 2. Year-by-year admission volume (shifted years)
  # ---------------------------------------------------------------
  ln("---")
  ln("## 2. Year-by-Year Admission Volume (shifted years)")
  ln("")
  ln("MIMIC-IV dates are randomly shifted per patient — these years do not")
  ln("correspond to real calendar time. Use `anchor_year_group` for the")
  ln("real-world era an anchor maps to.")
  ln("")
  yearly <- safe_query(sprintf("
    SELECT EXTRACT(year FROM admittime)::int AS yr,
           COUNT(DISTINCT subject_id) AS n_patients,
           COUNT(*) AS n_admissions
    FROM %s
    WHERE admittime IS NOT NULL
    GROUP BY EXTRACT(year FROM admittime)
    ORDER BY yr
  ", qualify("admissions")), "yearly admissions")
  if (nrow(yearly) > 0) {
    ln("| Year (shifted) | Distinct Patients | Admissions |")
    ln("|----------------|-------------------|------------|")
    for (i in seq_len(nrow(yearly))) {
      ln("| ", yearly$yr[i], " | ", sup(yearly$n_patients[i]), " | ",
         format(yearly$n_admissions[i], big.mark = ","), " |")
    }
    ln("")
  }

  ln("### `anchor_year_group` Distribution (real-world era)")
  ln("")
  ayg <- safe_query(sprintf("
    SELECT anchor_year_group, COUNT(DISTINCT subject_id) AS n
    FROM %s
    GROUP BY anchor_year_group
    ORDER BY anchor_year_group
  ", qualify("patients")), "anchor_year_group")
  if (nrow(ayg) > 0) {
    ln("| Era | Patients |")
    ln("|-----|----------|")
    for (i in seq_len(nrow(ayg))) {
      ln("| ", ayg$anchor_year_group[i], " | ", sup(ayg$n[i]), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 3. Age distributions
  # ---------------------------------------------------------------
  ln("---")
  ln("## 3. Age Distributions")
  ln("")
  ln("`anchor_age` is age at the patient's anchor year. MIMIC-IV already")
  ln("top-codes ages > 89 to 91. We bucket into the same bins used by")
  ln("the PCORnet/OMOP profilers for cross-CDM comparability.")
  ln("")

  ln("### 3a. Anchor Age Distribution")
  ln("")
  age_anchor <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT subject_id) AS n
    FROM (
      SELECT subject_id,
        CASE
          WHEN anchor_age < 18 THEN '0-17'
          WHEN anchor_age < 45 THEN '18-44'
          WHEN anchor_age < 65 THEN '45-64'
          WHEN anchor_age < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s
      WHERE anchor_age IS NOT NULL
    ) t
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", qualify("patients")), "anchor age")
  if (nrow(age_anchor) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_anchor))) {
      ln("| ", age_anchor$age_bin[i], " | ", sup(age_anchor$n[i]), " |")
    }
    ln("")
  }

  ln("### 3b. Age at First Admission")
  ln("")
  ln("Approximated as `anchor_age + (admit_year - anchor_year)` for each")
  ln("patient's earliest admission.")
  ln("")
  age_first <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT subject_id) AS n
    FROM (
      SELECT p.subject_id,
        CASE
          WHEN (p.anchor_age +
                (EXTRACT(year FROM f.first_admit)::int - p.anchor_year)) < 18 THEN '0-17'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM f.first_admit)::int - p.anchor_year)) < 45 THEN '18-44'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM f.first_admit)::int - p.anchor_year)) < 65 THEN '45-64'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM f.first_admit)::int - p.anchor_year)) < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s p
      INNER JOIN (
        SELECT subject_id, MIN(admittime) AS first_admit
        FROM %s
        WHERE admittime IS NOT NULL
        GROUP BY subject_id
      ) f ON p.subject_id = f.subject_id
      WHERE p.anchor_age IS NOT NULL
    ) t
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", qualify("patients"), qualify("admissions")), "age at first admit")
  if (nrow(age_first) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_first))) {
      ln("| ", age_first$age_bin[i], " | ", sup(age_first$n[i]), " |")
    }
    ln("")
  }

  ln("### 3c. Age at Death (in-hospital)")
  ln("")
  ln("Among admissions where `hospital_expire_flag = 1`.")
  ln("")
  age_death <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT subject_id) AS n
    FROM (
      SELECT p.subject_id,
        CASE
          WHEN (p.anchor_age +
                (EXTRACT(year FROM a.deathtime)::int - p.anchor_year)) < 18 THEN '0-17'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM a.deathtime)::int - p.anchor_year)) < 45 THEN '18-44'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM a.deathtime)::int - p.anchor_year)) < 65 THEN '45-64'
          WHEN (p.anchor_age +
                (EXTRACT(year FROM a.deathtime)::int - p.anchor_year)) < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s p
      INNER JOIN %s a ON p.subject_id = a.subject_id
      WHERE p.anchor_age IS NOT NULL
        AND a.hospital_expire_flag = 1
        AND a.deathtime IS NOT NULL
    ) t
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", qualify("patients"), qualify("admissions")), "age at death")
  if (nrow(age_death) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_death))) {
      ln("| ", age_death$age_bin[i], " | ", sup(age_death$n[i]), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 4. Demographics (admissions table — race / insurance / language)
  # ---------------------------------------------------------------
  ln("---")
  ln("## 4. Demographic Distributions")
  ln("")
  for (col in c("gender")) {
    r <- safe_query(sprintf("
      SELECT COALESCE(%s, '(NULL)') AS val, COUNT(DISTINCT subject_id) AS n
      FROM %s
      GROUP BY %s
      ORDER BY n DESC
    ", col, qualify("patients"), col), paste0("patients.", col))
    if (nrow(r) > 0) {
      ln("### patients.", col)
      ln("")
      ln("| Value | Patients |")
      ln("|-------|----------|")
      for (i in seq_len(nrow(r))) {
        ln("| ", r$val[i], " | ", sup(r$n[i]), " |")
      }
      ln("")
    }
  }

  for (col in c("race", "insurance", "marital_status", "language")) {
    r <- safe_query(sprintf("
      SELECT COALESCE(%s, '(NULL)') AS val, COUNT(DISTINCT subject_id) AS n
      FROM %s
      GROUP BY %s
      ORDER BY n DESC
      LIMIT 30
    ", col, qualify("admissions"), col), paste0("admissions.", col))
    if (nrow(r) > 0) {
      ln("### admissions.", col)
      ln("")
      ln("| Value | Patients |")
      ln("|-------|----------|")
      for (i in seq_len(nrow(r))) {
        ln("| ", r$val[i], " | ", sup(r$n[i]), " |")
      }
      ln("")
    }
  }

  # ---------------------------------------------------------------
  # 5. Admission types and dispositions
  # ---------------------------------------------------------------
  ln("---")
  ln("## 5. Admission Type and Discharge Disposition")
  ln("")
  for (col in c("admission_type", "admission_location", "discharge_location")) {
    r <- safe_query(sprintf("
      SELECT COALESCE(%s, '(NULL)') AS val, COUNT(*) AS n
      FROM %s
      GROUP BY %s
      ORDER BY n DESC
    ", col, qualify("admissions"), col), paste0("admissions.", col))
    if (nrow(r) > 0) {
      ln("### admissions.", col)
      ln("")
      ln("| Value | Admissions |")
      ln("|-------|------------|")
      for (i in seq_len(nrow(r))) {
        ln("| ", r$val[i], " | ", sup(r$n[i]), " |")
      }
      ln("")
    }
  }

  # ---------------------------------------------------------------
  # 6. ICU stay characteristics
  # ---------------------------------------------------------------
  ln("---")
  ln("## 6. ICU Stay Characteristics")
  ln("")
  cu <- safe_query(sprintf("
    SELECT first_careunit, COUNT(*) AS n_stays,
           COUNT(DISTINCT subject_id) AS n_patients,
           ROUND(AVG(los)::numeric, 2) AS mean_los_days
    FROM %s
    GROUP BY first_careunit
    ORDER BY n_stays DESC
  ", qualify("icustays")), "ICU careunit")
  if (nrow(cu) > 0) {
    ln("| First Care Unit | Stays | Patients | Mean LOS (days) |")
    ln("|------------------|-------|----------|-----------------|")
    for (i in seq_len(nrow(cu))) {
      ln("| ", cu$first_careunit[i], " | ",
         sup(cu$n_stays[i]), " | ", sup(cu$n_patients[i]), " | ",
         cu$mean_los_days[i], " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 7. Top diagnoses (ICD)
  # ---------------------------------------------------------------
  ln("---")
  ln("## 7. Top 30 Diagnoses (ICD)")
  ln("")
  ln("Joined to `d_icd_diagnoses` for human-readable titles where possible.")
  ln("")
  top_dx <- safe_query(sprintf("
    SELECT d.icd_code,
           d.icd_version,
           dd.long_title,
           COUNT(DISTINCT d.subject_id) AS n_patients,
           COUNT(*) AS n_rows
    FROM %s d
    LEFT JOIN %s dd
      ON d.icd_code = dd.icd_code AND d.icd_version = dd.icd_version
    GROUP BY d.icd_code, d.icd_version, dd.long_title
    ORDER BY n_patients DESC
    LIMIT 30
  ", qualify("diagnoses_icd"), qualify("d_icd_diagnoses")), "top diagnoses")
  if (nrow(top_dx) > 0) {
    ln("| ICD Code | Version | Title | Patients | Rows |")
    ln("|----------|---------|-------|----------|------|")
    for (i in seq_len(nrow(top_dx))) {
      r <- top_dx[i, ]
      ln("| ", r$icd_code, " | ", r$icd_version, " | ",
         r$long_title %||% "—", " | ",
         sup(r$n_patients), " | ",
         format(r$n_rows, big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 8. Top labs (with d_labitems lookup)
  # ---------------------------------------------------------------
  ln("---")
  ln("## 8. Top 50 Lab Items by Patient Count")
  ln("")
  top_labs <- safe_query(sprintf("
    SELECT le.itemid,
           dl.label,
           dl.category,
           COUNT(DISTINCT le.subject_id) AS n_patients,
           COUNT(*) AS n_results
    FROM %s le
    LEFT JOIN %s dl ON le.itemid = dl.itemid
    GROUP BY le.itemid, dl.label, dl.category
    ORDER BY n_patients DESC
    LIMIT 50
  ", qualify("labevents"), qualify("d_labitems")), "top labs")
  if (nrow(top_labs) > 0) {
    ln("| ItemID | Label | Category | Patients | Results |")
    ln("|--------|-------|----------|----------|---------|")
    for (i in seq_len(nrow(top_labs))) {
      r <- top_labs[i, ]
      ln("| ", r$itemid, " | ", r$label %||% "—", " | ",
         r$category %||% "—", " | ",
         sup(r$n_patients), " | ",
         format(r$n_results, big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 9. Top medications
  # ---------------------------------------------------------------
  ln("---")
  ln("## 9. Top 30 Medications by Patient Count")
  ln("")
  top_rx <- safe_query(sprintf("
    SELECT drug,
           COUNT(DISTINCT subject_id) AS n_patients,
           COUNT(*) AS n_rx
    FROM %s
    WHERE drug IS NOT NULL
    GROUP BY drug
    ORDER BY n_patients DESC
    LIMIT 30
  ", qualify("prescriptions")), "top prescriptions")
  if (nrow(top_rx) > 0) {
    ln("| Drug | Patients | Prescriptions |")
    ln("|------|----------|---------------|")
    for (i in seq_len(nrow(top_rx))) {
      r <- top_rx[i, ]
      ln("| ", r$drug, " | ", sup(r$n_patients), " | ",
         format(r$n_rx, big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 10. Mortality summary
  # ---------------------------------------------------------------
  ln("---")
  ln("## 10. Mortality Summary")
  ln("")
  m <- safe_query(sprintf("
    SELECT
      COUNT(DISTINCT subject_id) AS n_total_pts,
      COUNT(DISTINCT CASE WHEN dod IS NOT NULL THEN subject_id END) AS n_with_dod
    FROM %s
  ", qualify("patients")), "mortality")
  if (nrow(m) > 0) {
    ln("Patients with any recorded date of death (`dod`): ",
       sup(m$n_with_dod[1]), " of ", sup(m$n_total_pts[1]))
  }
  in_hosp <- safe_query(sprintf("
    SELECT
      COUNT(*) AS n_admissions,
      SUM(hospital_expire_flag) AS n_in_hosp_deaths
    FROM %s
  ", qualify("admissions")), "in-hospital deaths")
  if (nrow(in_hosp) > 0) {
    ln("In-hospital deaths (admissions with hospital_expire_flag=1): ",
       sup(in_hosp$n_in_hosp_deaths[1]), " of ",
       format(in_hosp$n_admissions[1], big.mark = ","), " admissions")
    ln("")
  }

  # ---------------------------------------------------------------
  # 11. Column completeness
  # ---------------------------------------------------------------
  ln("---")
  ln("## 11. Column Completeness — Key Tables")
  ln("")
  completeness <- list(
    list(tbl = "admissions", cols = c("admittime", "dischtime", "deathtime",
                                       "admission_type", "race", "insurance")),
    list(tbl = "labevents",  cols = c("charttime", "valuenum", "valueuom",
                                       "ref_range_lower", "ref_range_upper")),
    list(tbl = "prescriptions", cols = c("starttime", "stoptime", "drug",
                                          "dose_val_rx", "ndc")),
    list(tbl = "icustays",   cols = c("intime", "outtime", "los",
                                       "first_careunit", "last_careunit"))
  )
  for (ct in completeness) {
    total <- safe_query(sprintf("SELECT COUNT(*) AS n FROM %s",
                                qualify(ct$tbl)),
                        paste0(ct$tbl, " total"))
    if (nrow(total) == 0 || total$n[1] == 0) next
    ln("### ", ct$tbl, " (", format(total$n[1], big.mark = ","), " rows)")
    ln("")
    ln("| Column | % Non-NULL |")
    ln("|--------|-----------|")
    for (col in ct$cols) {
      r <- safe_query(sprintf("
        SELECT COUNT(*) AS n FROM %s WHERE %s IS NOT NULL
      ", qualify(ct$tbl), col), paste0(ct$tbl, ".", col))
      if (nrow(r) > 0) {
        pct <- round(100 * r$n[1] / total$n[1], 1)
        ln("| ", col, " | ", pct, "% |")
      }
    }
    ln("")
  }

  writeLines(out, output_file)
  message("Data profile written to: ", output_file)
  invisible(out)
}
```

## Run everything

```{r run-schema}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  extract_schema_mimic(con, output_file = paths$schema)
})
```

```{r run-profile}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  profile_mimic(con, output_file = paths$data_profile,
                suppress_threshold = suppress_threshold)
})
```

```{r done-banner}
#| echo: false
done_banner(`schema  ` = paths$schema,
            `profile ` = paths$data_profile)
```
