---
title: "PCORnet CDM Data Profiler"
format: html
editor: visual
params:
  db_config: "databases/secure_pcornet_cdw.yaml"
---

# PCORnet CDM Data Profiler

Generic profiler for any database conforming to the [PCORnet Common Data
Model](https://pcornet.org/data/). Produces a schema dump and a PHI-safe
data profile at the paths declared in the YAML config pointed to by the
`db_config` param above.

**To use for a different PCORnet site:** edit `params.db_config` to point at
your DB's YAML config, then **Restart R + Run All Chunks**.

**SQL dialect:** queries assume **MS SQL Server** (the most common engine for
production PCORnet CDWs). For PCORnet on DuckDB or Postgres, swap `TOP N` →
`LIMIT N`, `FORMAT(date,'yyyy-MM')` → `strftime`/`to_char`, `ISNULL` →
`COALESCE`, `DATEDIFF(year,a,b)` → `EXTRACT(year FROM age(b,a))`.

```{r load-libraries}
library(tidyverse)
library(DBI)
library(yaml)
library(here)
options(encoding = "UTF-8")
```

## Load configuration and shared scaffolding

```{r load-config}
#| code-fold: true
source(here::here("profilers/_helpers.R"))

# `params` is only defined during `quarto render`. Fall back to the default
# for interactive "Run All Chunks" in RStudio.
db_config_path <- if (exists("params") && !is.null(params$db_config)) {
  params$db_config
} else {
  "databases/secure_pcornet_cdw.yaml"
}

config <- load_profiler_config(db_config_path)
cfg     <- config$cfg
paths   <- config$paths
schema  <- schema_name_from_cfg(cfg)
suppress_threshold <- 11
sup     <- make_suppress(suppress_threshold)
```

## Preflight

```{r preflight}
preflight_check(cfg$connection$r_code, label = cfg$id %||% "db")
```

## Schema extraction

```{r fn-extract-schema}
#| code-fold: true

# MS SQL Server schema introspection. Produces a markdown-ish text file with
# tables, columns, types, nullability, primary keys, and FKs.
extract_schema_pcornet <- function(con,
                                   schema_name = "dbo",
                                   output_file = "schema.txt") {
  tables <- DBI::dbGetQuery(con, sprintf("
    SELECT TABLE_NAME
    FROM INFORMATION_SCHEMA.TABLES
    WHERE TABLE_SCHEMA = '%s' AND TABLE_TYPE = 'BASE TABLE'
    ORDER BY TABLE_NAME
  ", schema_name))

  out <- c(
    paste0("# PCORnet CDM Schema (", cfg$id %||% "db", ")"),
    paste0("# Generated: ", Sys.time()),
    paste0("# Schema: ", schema_name),
    ""
  )

  for (table_name in tables$TABLE_NAME) {
    columns <- DBI::dbGetQuery(con, sprintf("
      SELECT COLUMN_NAME, DATA_TYPE, CHARACTER_MAXIMUM_LENGTH, IS_NULLABLE
      FROM INFORMATION_SCHEMA.COLUMNS
      WHERE TABLE_SCHEMA = '%s' AND TABLE_NAME = '%s'
      ORDER BY ORDINAL_POSITION
    ", schema_name, table_name))

    pks <- DBI::dbGetQuery(con, sprintf("
      SELECT COLUMN_NAME
      FROM INFORMATION_SCHEMA.KEY_COLUMN_USAGE
      WHERE TABLE_SCHEMA = '%s' AND TABLE_NAME = '%s'
        AND CONSTRAINT_NAME LIKE 'PK_%%'
    ", schema_name, table_name))

    fks <- DBI::dbGetQuery(con, sprintf("
      SELECT fk.COLUMN_NAME, pk.TABLE_NAME AS REFERENCES_TABLE,
             pk.COLUMN_NAME AS REFERENCES_COLUMN
      FROM INFORMATION_SCHEMA.REFERENTIAL_CONSTRAINTS rc
      JOIN INFORMATION_SCHEMA.KEY_COLUMN_USAGE fk
        ON rc.CONSTRAINT_NAME = fk.CONSTRAINT_NAME
      JOIN INFORMATION_SCHEMA.KEY_COLUMN_USAGE pk
        ON rc.UNIQUE_CONSTRAINT_NAME = pk.CONSTRAINT_NAME
      WHERE fk.TABLE_SCHEMA = '%s' AND fk.TABLE_NAME = '%s'
    ", schema_name, table_name))

    out <- c(out, paste0("## Table: ", table_name), "", "### Columns:")
    for (i in seq_len(nrow(columns))) {
      col <- columns[i, ]
      type_info <- col$DATA_TYPE
      if (!is.na(col$CHARACTER_MAXIMUM_LENGTH))
        type_info <- paste0(type_info, "(", col$CHARACTER_MAXIMUM_LENGTH, ")")
      nullable  <- ifelse(col$IS_NULLABLE == "YES", "NULL", "NOT NULL")
      pk_marker <- ifelse(col$COLUMN_NAME %in% pks$COLUMN_NAME, " [PRIMARY KEY]", "")
      fk_info <- ""
      if (col$COLUMN_NAME %in% fks$COLUMN_NAME) {
        fk_row <- fks[fks$COLUMN_NAME == col$COLUMN_NAME, ]
        fk_info <- paste0(" [FK -> ", fk_row$REFERENCES_TABLE, ".",
                          fk_row$REFERENCES_COLUMN, "]")
      }
      out <- c(out, paste0("- ", col$COLUMN_NAME, ": ", type_info,
                           " (", nullable, ")", pk_marker, fk_info))
    }
    out <- c(out, "")
  }
  out <- c(out, "---", "## Summary", paste0("Total tables: ", nrow(tables)))
  writeLines(out, output_file)
  message("Schema written to: ", output_file)
}
```

## Data profile

```{r fn-profile-pcornet}
#| code-fold: true

profile_pcornet <- function(con,
                            schema_name = "dbo",
                            output_file,
                            suppress_threshold = 11) {

  safe_query <- make_safe_query(con)
  sup <- make_suppress(suppress_threshold)

  out <- character()
  ln  <- function(...) out <<- c(out, paste0(...))

  ln("# PCORnet CDM Data Profile")
  ln("# Generated: ", Sys.time())
  ln("# Schema: ", schema_name)
  ln("# This file contains ONLY aggregate counts and de-identified")
  ln("# distributions. Counts < ", suppress_threshold, " are suppressed.")
  ln("# Ages \u2265 90 are top-coded per HIPAA Safe Harbor.")
  ln("")

  # ---------------------------------------------------------------
  # 1. Overall size and temporal coverage
  # ---------------------------------------------------------------
  ln("---")
  ln("## 1. Overall Size and Temporal Coverage")
  ln("")
  total_pts <- safe_query(sprintf(
    "SELECT COUNT(DISTINCT PATID) AS n FROM %s.DEMOGRAPHIC", schema_name),
    "DEMOGRAPHIC")
  if (nrow(total_pts) > 0) {
    ln("Total distinct patients in DEMOGRAPHIC: ", sup(total_pts$n[1]))
    ln("")
  }

  table_dates <- list(
    list(tbl = "ENCOUNTER",     date_col = "ADMIT_DATE"),
    list(tbl = "DIAGNOSIS",     date_col = "ADMIT_DATE"),
    list(tbl = "PROCEDURES",    date_col = "PX_DATE"),
    list(tbl = "PRESCRIBING",   date_col = "RX_ORDER_DATE"),
    list(tbl = "DISPENSING",    date_col = "DISPENSE_DATE"),
    list(tbl = "LAB_RESULT_CM", date_col = "RESULT_DATE"),
    list(tbl = "VITAL",         date_col = "MEASURE_DATE"),
    list(tbl = "MED_ADMIN",     date_col = "MEDADMIN_START_DATE"),
    list(tbl = "CONDITION",     date_col = "ONSET_DATE"),
    list(tbl = "DEATH",         date_col = "DEATH_DATE"),
    list(tbl = "ENROLLMENT",    date_col = "ENR_START_DATE")
  )

  ln("| Table | Distinct Patients | Rows | Earliest (YYYY-MM) | Latest (YYYY-MM) |")
  ln("|-------|-------------------|------|---------------------|-------------------|")
  for (td in table_dates) {
    res <- safe_query(sprintf("
      SELECT COUNT(DISTINCT PATID) AS n_patients,
             COUNT(*)              AS n_rows,
             FORMAT(MIN(%s), 'yyyy-MM') AS earliest,
             FORMAT(MAX(%s), 'yyyy-MM') AS latest
      FROM %s.%s
    ", td$date_col, td$date_col, schema_name, td$tbl), td$tbl)
    if (nrow(res) > 0 && !is.na(res$n_patients[1]) && res$n_patients[1] > 0) {
      ln("| ", td$tbl, " | ", sup(res$n_patients[1]), " | ",
         format(res$n_rows[1], big.mark = ","), " | ",
         res$earliest[1] %||% "—", " | ",
         res$latest[1]   %||% "—", " |")
    }
  }
  ln("")

  # ---------------------------------------------------------------
  # 2. Year-by-year encounter volume
  # ---------------------------------------------------------------
  ln("---")
  ln("## 2. Year-by-Year Patient Volume (ENCOUNTER)")
  ln("")
  yearly <- safe_query(sprintf("
    SELECT YEAR(ADMIT_DATE) AS yr,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_encounters
    FROM %s.ENCOUNTER
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY YEAR(ADMIT_DATE)
    ORDER BY yr
  ", schema_name), "yearly encounters")
  if (nrow(yearly) > 0) {
    ln("| Year | Distinct Patients | Encounters |")
    ln("|------|-------------------|------------|")
    for (i in seq_len(nrow(yearly))) {
      ln("| ", yearly$yr[i], " | ", sup(yearly$n_patients[i]), " | ",
         format(yearly$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 3. Age distributions (PHI-safe aggregates)
  # ---------------------------------------------------------------
  # DOB is queried but never written. Outputs are bins only, with 90+
  # top-coded per HIPAA Safe Harbor. See feedback_phi_in_profilers memory.
  ln("---")
  ln("## 3. Age Distributions")
  ln("")
  ln("Aggregate distributions derived from `DEMOGRAPHIC.BIRTH_DATE`. The DOB")
  ln("itself is never written to this profile. Ages \u2265 90 are top-coded.")
  ln("")

  ln("### 3a. Birth Decade Histogram")
  ln("")
  bd <- safe_query(sprintf("
    SELECT (YEAR(BIRTH_DATE) / 10) * 10 AS birth_decade,
           COUNT(DISTINCT PATID) AS n
    FROM %s.DEMOGRAPHIC
    WHERE BIRTH_DATE IS NOT NULL
    GROUP BY (YEAR(BIRTH_DATE) / 10) * 10
    ORDER BY birth_decade
  ", schema_name), "birth decade")
  if (nrow(bd) > 0) {
    ln("| Birth Decade | Patients |")
    ln("|--------------|----------|")
    for (i in seq_len(nrow(bd))) {
      ln("| ", bd$birth_decade[i], "s | ", sup(bd$n[i]), " |")
    }
    ln("")
  }

  ln("### 3b. Current Age Distribution")
  ln("")
  ln("Age in years from BIRTH_DATE to today, binned. 90+ top-coded.")
  ln("")
  age_now <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT PATID) AS n
    FROM (
      SELECT PATID,
        CASE
          WHEN DATEDIFF(year, BIRTH_DATE, GETDATE()) < 18 THEN '0-17'
          WHEN DATEDIFF(year, BIRTH_DATE, GETDATE()) < 45 THEN '18-44'
          WHEN DATEDIFF(year, BIRTH_DATE, GETDATE()) < 65 THEN '45-64'
          WHEN DATEDIFF(year, BIRTH_DATE, GETDATE()) < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s.DEMOGRAPHIC
      WHERE BIRTH_DATE IS NOT NULL
    ) t
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", schema_name), "age now")
  if (nrow(age_now) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_now))) {
      ln("| ", age_now$age_bin[i], " | ", sup(age_now$n[i]), " |")
    }
    ln("")
  }

  ln("### 3c. Age at First Encounter")
  ln("")
  ln("More relevant for TTE feasibility than current age — answers 'how old")
  ln("are patients when they first appear in this CDW?'. 90+ top-coded.")
  ln("")
  age_first <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT PATID) AS n
    FROM (
      SELECT d.PATID,
        CASE
          WHEN DATEDIFF(year, d.BIRTH_DATE, f.first_date) < 18 THEN '0-17'
          WHEN DATEDIFF(year, d.BIRTH_DATE, f.first_date) < 45 THEN '18-44'
          WHEN DATEDIFF(year, d.BIRTH_DATE, f.first_date) < 65 THEN '45-64'
          WHEN DATEDIFF(year, d.BIRTH_DATE, f.first_date) < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s.DEMOGRAPHIC d
      INNER JOIN (
        SELECT PATID, MIN(ADMIT_DATE) AS first_date
        FROM %s.ENCOUNTER
        WHERE ADMIT_DATE IS NOT NULL
        GROUP BY PATID
      ) f ON d.PATID = f.PATID
      WHERE d.BIRTH_DATE IS NOT NULL
    ) t
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", schema_name, schema_name), "age at first encounter")
  if (nrow(age_first) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_first))) {
      ln("| ", age_first$age_bin[i], " | ", sup(age_first$n[i]), " |")
    }
    ln("")
  }

  ln("### 3d. Age at Death")
  ln("")
  ln("Among patients with a DEATH record. 90+ top-coded.")
  ln("")
  age_death <- safe_query(sprintf("
    SELECT age_bin, COUNT(DISTINCT d.PATID) AS n
    FROM (
      SELECT d.PATID,
        CASE
          WHEN DATEDIFF(year, d.BIRTH_DATE, de.DEATH_DATE) < 18 THEN '0-17'
          WHEN DATEDIFF(year, d.BIRTH_DATE, de.DEATH_DATE) < 45 THEN '18-44'
          WHEN DATEDIFF(year, d.BIRTH_DATE, de.DEATH_DATE) < 65 THEN '45-64'
          WHEN DATEDIFF(year, d.BIRTH_DATE, de.DEATH_DATE) < 90 THEN '65-89'
          ELSE '90+'
        END AS age_bin
      FROM %s.DEMOGRAPHIC d
      INNER JOIN %s.DEATH de ON d.PATID = de.PATID
      WHERE d.BIRTH_DATE IS NOT NULL AND de.DEATH_DATE IS NOT NULL
    ) d
    GROUP BY age_bin
    ORDER BY CASE age_bin WHEN '0-17' THEN 1 WHEN '18-44' THEN 2
                          WHEN '45-64' THEN 3 WHEN '65-89' THEN 4
                          ELSE 5 END
  ", schema_name, schema_name), "age at death")
  if (nrow(age_death) > 0) {
    ln("| Age Bin | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(age_death))) {
      ln("| ", age_death$age_bin[i], " | ", sup(age_death$n[i]), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 4. Diagnosis coding by year (ICD-9 vs ICD-10)
  # ---------------------------------------------------------------
  ln("---")
  ln("## 4. Diagnosis Coding Systems by Year")
  ln("")
  ln("DX_TYPE: '09' = ICD-9-CM, '10' = ICD-10-CM, 'SM' = SNOMED, etc.")
  ln("US transitioned ICD-9 \u2192 ICD-10 on 2015-10-01.")
  ln("")
  dx <- safe_query(sprintf("
    SELECT YEAR(ADMIT_DATE) AS yr, DX_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rows
    FROM %s.DIAGNOSIS
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY YEAR(ADMIT_DATE), DX_TYPE
    ORDER BY yr, DX_TYPE
  ", schema_name), "DX by year")
  if (nrow(dx) > 0) {
    ln("| Year | DX_TYPE | Patients | Rows |")
    ln("|------|---------|----------|------|")
    for (i in seq_len(nrow(dx))) {
      ln("| ", dx$yr[i], " | ", dx$DX_TYPE[i], " | ",
         sup(dx$n_patients[i]), " | ",
         format(dx$n_rows[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 5. Procedure coding systems
  # ---------------------------------------------------------------
  ln("---")
  ln("## 5. Procedure Coding Systems")
  ln("")
  ln("PX_TYPE: 'CH' = CPT/HCPCS, '10' = ICD-10-PCS, '09' = ICD-9-CM, etc.")
  ln("")
  px <- safe_query(sprintf("
    SELECT PX_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rows,
           FORMAT(MIN(PX_DATE), 'yyyy-MM') AS earliest,
           FORMAT(MAX(PX_DATE), 'yyyy-MM') AS latest
    FROM %s.PROCEDURES
    WHERE PX_DATE IS NOT NULL
    GROUP BY PX_TYPE
    ORDER BY n_patients DESC
  ", schema_name), "PX_TYPE")
  if (nrow(px) > 0) {
    ln("| PX_TYPE | Patients | Rows | Earliest | Latest |")
    ln("|---------|----------|------|----------|--------|")
    for (i in seq_len(nrow(px))) {
      ln("| ", px$PX_TYPE[i], " | ", sup(px$n_patients[i]), " | ",
         format(px$n_rows[i], big.mark = ","), " | ",
         px$earliest[i], " | ", px$latest[i], " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 6. Top labs by patient count
  # ---------------------------------------------------------------
  ln("---")
  ln("## 6. Top 50 LAB_LOINCs by Patient Count")
  ln("")
  top_labs <- safe_query(sprintf("
    SELECT TOP 50 LAB_LOINC,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_results
    FROM %s.LAB_RESULT_CM
    WHERE LAB_LOINC IS NOT NULL
    GROUP BY LAB_LOINC
    ORDER BY n_patients DESC
  ", schema_name), "top labs")
  if (nrow(top_labs) > 0) {
    ln("| LAB_LOINC | Patients | Results |")
    ln("|-----------|----------|---------|")
    for (i in seq_len(nrow(top_labs))) {
      r <- top_labs[i, ]
      ln("| ", r$LAB_LOINC, " | ", sup(r$n_patients), " | ",
         format(r$n_results, big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 7. Top medications
  # ---------------------------------------------------------------
  ln("---")
  ln("## 7. Top 50 Medications by Patient Count")
  ln("")
  has_raw <- safe_query(sprintf("
    SELECT TOP 1 COLUMN_NAME
    FROM INFORMATION_SCHEMA.COLUMNS
    WHERE TABLE_SCHEMA = '%s' AND TABLE_NAME = 'PRESCRIBING'
      AND COLUMN_NAME = 'RAW_RX_MED_NAME'
  ", schema_name), "RAW_RX_MED_NAME check")
  med_col <- if (nrow(has_raw) > 0) "RAW_RX_MED_NAME" else "RXNORM_CUI"

  top_rx <- safe_query(sprintf("
    SELECT TOP 50 %s AS med,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rx
    FROM %s.PRESCRIBING
    WHERE %s IS NOT NULL AND %s <> ''
    GROUP BY %s
    ORDER BY n_patients DESC
  ", med_col, schema_name, med_col, med_col, med_col), "top Rx")
  if (nrow(top_rx) > 0) {
    ln("Search column: **", med_col, "**")
    ln("")
    ln("| Medication | Patients | Prescriptions |")
    ln("|------------|----------|---------------|")
    for (i in seq_len(nrow(top_rx))) {
      r <- top_rx[i, ]
      ln("| ", r$med, " | ", sup(r$n_patients), " | ",
         format(r$n_rx, big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 8. Demographic distributions
  # ---------------------------------------------------------------
  ln("---")
  ln("## 8. Demographic Distributions")
  ln("")
  for (col in c("SEX", "RACE", "HISPANIC")) {
    dist <- safe_query(sprintf("
      SELECT ISNULL(CAST(%s AS varchar(20)), '(NULL)') AS val,
             COUNT(DISTINCT PATID) AS n
      FROM %s.DEMOGRAPHIC
      GROUP BY %s
      ORDER BY n DESC
    ", col, schema_name, col), paste0("DEMOGRAPHIC.", col))
    if (nrow(dist) > 0) {
      ln("### ", col)
      ln("")
      ln("| Value | Patients |")
      ln("|-------|----------|")
      for (i in seq_len(nrow(dist))) {
        ln("| ", dist$val[i], " | ", sup(dist$n[i]), " |")
      }
      ln("")
    }
  }

  # ---------------------------------------------------------------
  # 9. Encounter type distribution
  # ---------------------------------------------------------------
  ln("---")
  ln("## 9. Encounter Types")
  ln("")
  ln("AV=Ambulatory, ED=Emergency, EI=ED-to-Inpt, IP=Inpatient, IS=Institutional,")
  ln("OS=Outpt Surgery, OA=Other Ambulatory, TH=Telehealth, NI/UN=No info, OT=Other.")
  ln("")
  enc <- safe_query(sprintf("
    SELECT ISNULL(ENC_TYPE, '(NULL)') AS ENC_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_encounters
    FROM %s.ENCOUNTER
    GROUP BY ENC_TYPE
    ORDER BY n_patients DESC
  ", schema_name), "ENC_TYPE")
  if (nrow(enc) > 0) {
    ln("| ENC_TYPE | Patients | Encounters |")
    ln("|----------|----------|------------|")
    for (i in seq_len(nrow(enc))) {
      ln("| ", enc$ENC_TYPE[i], " | ", sup(enc$n_patients[i]), " | ",
         format(enc$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 10. Death records
  # ---------------------------------------------------------------
  ln("---")
  ln("## 10. Death Records")
  ln("")
  death_res <- safe_query(sprintf("
    SELECT COUNT(DISTINCT PATID) AS n_patients, COUNT(*) AS n_rows
    FROM %s.DEATH
  ", schema_name), "DEATH count")
  if (nrow(death_res) > 0) {
    ln("Distinct patients with death records: ", sup(death_res$n_patients[1]))
    ln("Total rows: ", format(death_res$n_rows[1], big.mark = ","))
    ln("")
  }

  # ---------------------------------------------------------------
  # 11. Column completeness (NULL rates) for key tables
  # ---------------------------------------------------------------
  ln("---")
  ln("## 11. Column Completeness — Key Tables")
  ln("")
  ln("Percentage of rows where the column is NOT NULL.")
  ln("")
  completeness_tables <- list(
    list(tbl = "DEMOGRAPHIC", cols = c("BIRTH_DATE", "SEX", "RACE", "HISPANIC")),
    list(tbl = "PRESCRIBING", cols = c("RXNORM_CUI", "RX_ORDER_DATE",
                                       "RX_QUANTITY", "RX_DAYS_SUPPLY")),
    list(tbl = "DIAGNOSIS",   cols = c("DX", "DX_TYPE", "ADMIT_DATE", "DX_DATE")),
    list(tbl = "LAB_RESULT_CM", cols = c("LAB_LOINC", "RESULT_NUM", "RESULT_QUAL",
                                         "RESULT_DATE"))
  )
  for (ct in completeness_tables) {
    total <- safe_query(sprintf("SELECT COUNT(*) AS n FROM %s.%s",
                                schema_name, ct$tbl), paste0(ct$tbl, " total"))
    if (nrow(total) == 0 || total$n[1] == 0) next
    ln("### ", ct$tbl, " (", format(total$n[1], big.mark = ","), " rows)")
    ln("")
    ln("| Column | % Non-NULL |")
    ln("|--------|-----------|")
    for (col in ct$cols) {
      r <- safe_query(sprintf("
        SELECT COUNT(*) AS n FROM %s.%s WHERE %s IS NOT NULL
      ", schema_name, ct$tbl, col), paste0(ct$tbl, ".", col))
      if (nrow(r) > 0) {
        pct <- round(100 * r$n[1] / total$n[1], 1)
        ln("| ", col, " | ", pct, "% |")
      }
    }
    ln("")
  }

  # ---------------------------------------------------------------
  # 12. Top 30 ICD-10 codes
  # ---------------------------------------------------------------
  ln("---")
  ln("## 12. Top 30 ICD-10 Diagnosis Codes")
  ln("")
  top_dx <- safe_query(sprintf("
    SELECT TOP 30 DX, COUNT(DISTINCT PATID) AS n_patients
    FROM %s.DIAGNOSIS
    WHERE DX_TYPE = '10' AND DX IS NOT NULL
    GROUP BY DX
    ORDER BY n_patients DESC
  ", schema_name), "top ICD-10")
  if (nrow(top_dx) > 0) {
    ln("| ICD-10 Code | Patients |")
    ln("|-------------|----------|")
    for (i in seq_len(nrow(top_dx))) {
      ln("| ", top_dx$DX[i], " | ", sup(top_dx$n_patients[i]), " |")
    }
    ln("")
  }

  writeLines(out, output_file)
  message("Data profile written to: ", output_file)
  invisible(out)
}
```

## Run everything

```{r run-schema}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  extract_schema_pcornet(con, schema_name = schema, output_file = paths$schema)
})
```

```{r run-profile}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  profile_pcornet(con, schema_name = schema, output_file = paths$data_profile,
                  suppress_threshold = suppress_threshold)
})
```

```{r done-banner}
#| echo: false
done_banner(`schema  ` = paths$schema,
            `profile ` = paths$data_profile)
```
