---
title: "Protocol 03: OAC vs No OAC at CHA₂DS₂-VASc = 1 (Men) / 2 (Women) in AF"
subtitle: "Target Trial Emulation --- PCORnet CDW"
author: "Auto-Protocol Designer"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
    theme: cosmo
execute:
  eval: true
  warning: false
  message: true
---

## Study Overview

This protocol emulates a target trial comparing **oral anticoagulation (OAC)
initiation vs no OAC** for ischemic stroke/systemic embolism risk in adults
with non-valvular AF at the guideline treatment threshold: CHA₂DS₂-VASc = 1
(men) or 2 (women).

- **Design:** New-user, treated vs untreated cohort
- **Estimand:** ATE (Average Treatment Effect)
- **Method:** Inverse Probability Weighting (IPW) + Weighted Cox PH
- **Time zero:** First qualifying encounter with AF + target CHA₂DS₂-VASc + no prior OAC
- **Grace period:** ±7 days for OAC prescription
- **Follow-up:** 365 days (1 year)
- **Study period:** 2016-01-01 to 2025-12-31


<!-- ============================================================= -->
<!-- PART 1: Function definitions (Sections 0-8)                    -->
<!-- These chunks register functions but produce no visible output  -->
<!-- ============================================================= -->


## 0. Setup and Configuration

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(DBI)
library(odbc)
library(WeightIt)
library(cobalt)
library(survival)
library(sandwich)
library(lmtest)
library(EValue)
library(survminer)
```

```{r}
#| label: config

config <- list(
  question        = "OAC vs No OAC at CHA2DS2-VASc = 1 (Men) / 2 (Women) in AF",
  exposure_var    = "treatment",
  outcome_var     = "stroke_se_event",
  time_var        = "time_to_stroke_se",
  estimand        = "ATE",
  analysis_method = "ipw",
  study_start     = "2016-01-01",
  study_end       = "2025-12-31",
  followup_days   = 365,
  grace_period    = 7,         # +/- days for OAC prescription
  lookback_dx     = 730,       # 2 years for comorbidity lookback (CHA2DS2-VASc)
  lookback_lab    = 180,       # days for baseline lab lookback
  lookback_vital  = 365,       # days for baseline vital lookback
  lookback_rx     = 90         # days for concurrent medication lookback
)
```


## 1. Database Connection

```{r}
#| label: db-connect

connect_cdw <- function() {
  con <- DBI::dbConnect(odbc::odbc(), "SQLODBCD17CDM")
  return(con)
}
```


## 2. Cohort SQL

The cohort is built in sequential steps using temp tables:
`#af_patients` -> `#chadsvasc_components` -> `#chadsvasc_eligible` ->
`#eligible` -> `#treatment` -> `#outcomes` -> `#analytic_cohort`.

All tables are fully qualified as `CDW.dbo.TABLE_NAME`. Medications are
identified by RXNORM_CUI (SCD + SBD), diagnoses by ICD-10 with `DX_TYPE = '10'`.
Legacy encounters are excluded throughout.

The critical step is the CHA₂DS₂-VASc score computation in SQL, which builds
each component from DIAGNOSIS and DEMOGRAPHIC, then filters to the exact
target score (= 1 for men, = 2 for women).

```{r}
#| label: cohort-sql

build_cohort_sql <- function(config) {

  # ── Step 1: Eligible population ──
  # AF + CHA2DS2-VASc = 1 (men) / 2 (women), no prior OAC
  eligibility_sql <- paste0("
    -- =========================================================
    -- STEP 1: Eligible Population
    -- NVAF + CHA2DS2-VASc = 1 (men) / 2 (women), new to OAC
    -- =========================================================

    -- 1a. Find all patients with non-valvular AF
    IF OBJECT_ID('tempdb..#af_patients') IS NOT NULL DROP TABLE #af_patients;

    SELECT DISTINCT dx.PATID
    INTO #af_patients
    FROM CDW.dbo.DIAGNOSIS dx
    INNER JOIN CDW.dbo.ENCOUNTER e ON dx.ENCOUNTERID = e.ENCOUNTERID
      AND e.RAW_ENC_TYPE <> 'Legacy Encounter'
    WHERE dx.DX LIKE 'I48%'
      AND dx.DX NOT IN ('I48.3','I48.4','I48.92')  -- exclude flutter
      AND dx.DX_TYPE = '10'
      AND dx.ADMIT_DATE BETWEEN '", config$study_start, "' AND '", config$study_end, "'
    -- Exclude valvular AF
    EXCEPT
    SELECT DISTINCT vdx.PATID
    FROM CDW.dbo.DIAGNOSIS vdx
    WHERE (vdx.DX LIKE 'I05%' OR vdx.DX LIKE 'I06%'
           OR vdx.DX LIKE 'I07%' OR vdx.DX LIKE 'I08%'
           OR vdx.DX IN ('Z95.2','Z95.3','Z95.4'))
      AND vdx.DX_TYPE = '10'
      AND vdx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
    ;

    -- 1b. For each AF patient, find the first encounter where AF + target
    --     CHA2DS2-VASc score is met. Compute score from components.
    --     Time zero = first encounter date where score = 1 (men) / 2 (women).

    -- Build CHA2DS2-VASc components per patient, per encounter date
    -- We need the score AS OF each encounter to find the first date it hits threshold
    -- Strategy: find first AF encounter, then compute score at that point using
    -- all prior diagnoses. If score matches, that's time zero.

    IF OBJECT_ID('tempdb..#first_af_enc') IS NOT NULL DROP TABLE #first_af_enc;

    -- First AF encounter per patient (the earliest encounter with an AF Dx)
    SELECT
      dx.PATID,
      MIN(dx.ADMIT_DATE) AS first_af_date
    INTO #first_af_enc
    FROM CDW.dbo.DIAGNOSIS dx
    INNER JOIN CDW.dbo.ENCOUNTER e ON dx.ENCOUNTERID = e.ENCOUNTERID
      AND e.RAW_ENC_TYPE <> 'Legacy Encounter'
    INNER JOIN #af_patients af ON dx.PATID = af.PATID
    WHERE dx.DX LIKE 'I48%'
      AND dx.DX NOT IN ('I48.3','I48.4','I48.92')
      AND dx.DX_TYPE = '10'
      AND dx.ADMIT_DATE BETWEEN '", config$study_start, "' AND '", config$study_end, "'
    GROUP BY dx.PATID
    ;

    -- 1c. Compute CHA2DS2-VASc components at the first AF encounter date
    IF OBJECT_ID('tempdb..#chadsvasc_components') IS NOT NULL DROP TABLE #chadsvasc_components;

    SELECT
      fa.PATID,
      fa.first_af_date AS index_date,
      d.BIRTH_DATE,
      d.SEX,
      d.RACE,
      d.HISPANIC,
      DATEDIFF(year, d.BIRTH_DATE, fa.first_af_date) AS age_at_index,

      -- CHA2DS2-VASc components (lookback: all prior diagnoses)
      ISNULL(comps.chf_flag, 0) AS chf_flag,
      ISNULL(comps.htn_flag, 0) AS htn_flag,
      ISNULL(comps.dm_flag, 0) AS dm_flag,
      ISNULL(comps.stroke_tia_flag, 0) AS stroke_tia_flag,
      ISNULL(comps.vasc_flag, 0) AS vasc_flag,

      -- Age score
      CASE WHEN DATEDIFF(year, d.BIRTH_DATE, fa.first_af_date) >= 75 THEN 2
           WHEN DATEDIFF(year, d.BIRTH_DATE, fa.first_af_date) BETWEEN 65 AND 74 THEN 1
           ELSE 0 END AS age_score,

      -- Sex score
      CASE WHEN d.SEX = 'F' THEN 1 ELSE 0 END AS sex_score,

      -- Total CHA2DS2-VASc score
      ISNULL(comps.chf_flag, 0)
      + ISNULL(comps.htn_flag, 0)
      + CASE WHEN DATEDIFF(year, d.BIRTH_DATE, fa.first_af_date) >= 75 THEN 2
             WHEN DATEDIFF(year, d.BIRTH_DATE, fa.first_af_date) BETWEEN 65 AND 74 THEN 1
             ELSE 0 END
      + ISNULL(comps.dm_flag, 0)
      + ISNULL(comps.stroke_tia_flag, 0) * 2
      + ISNULL(comps.vasc_flag, 0)
      + CASE WHEN d.SEX = 'F' THEN 1 ELSE 0 END
      AS chadsvasc_total

    INTO #chadsvasc_components
    FROM #first_af_enc fa
    INNER JOIN CDW.dbo.DEMOGRAPHIC d ON fa.PATID = d.PATID
    LEFT JOIN (
      SELECT
        dx.PATID,
        MAX(CASE WHEN dx.DX LIKE 'I50%' THEN 1 ELSE 0 END) AS chf_flag,
        MAX(CASE WHEN dx.DX LIKE 'I10%' OR dx.DX LIKE 'I11%'
                   OR dx.DX LIKE 'I12%' OR dx.DX LIKE 'I13%'
                   OR dx.DX LIKE 'I15%' OR dx.DX LIKE 'I16%' THEN 1 ELSE 0 END) AS htn_flag,
        MAX(CASE WHEN dx.DX LIKE 'E10%' OR dx.DX LIKE 'E11%'
                   OR dx.DX LIKE 'E13%' THEN 1 ELSE 0 END) AS dm_flag,
        MAX(CASE WHEN dx.DX LIKE 'I63%' OR dx.DX LIKE 'G45%'
                   OR dx.DX LIKE 'I74%' THEN 1 ELSE 0 END) AS stroke_tia_flag,
        MAX(CASE WHEN dx.DX LIKE 'I21%' OR dx.DX LIKE 'I25%'
                   OR dx.DX LIKE 'I70%' OR dx.DX LIKE 'I71%'
                   OR dx.DX = 'I73.9' THEN 1 ELSE 0 END) AS vasc_flag
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN #first_af_enc fa2 ON dx.PATID = fa2.PATID
      WHERE dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= fa2.first_af_date
        AND dx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      GROUP BY dx.PATID
    ) comps ON fa.PATID = comps.PATID
    ;

    -- 1d. Filter to target score: CHA2DS2-VASc = 1 (men) or = 2 (women)
    --     Also exclude age >= 75 (2 pts -> score >= 3), prior stroke/TIA (2 pts),
    --     active cancer, and patients under 18.
    IF OBJECT_ID('tempdb..#eligible') IS NOT NULL DROP TABLE #eligible;

    SELECT
      cv.PATID,
      cv.index_date,
      cv.BIRTH_DATE,
      cv.SEX,
      cv.RACE,
      cv.HISPANIC,
      cv.age_at_index,
      cv.chadsvasc_total,
      cv.chf_flag,
      cv.htn_flag,
      cv.dm_flag,
      cv.vasc_flag,
      cv.age_score,
      cv.sex_score,
      -- Which component drives the score (the single non-sex, non-zero factor)
      CASE
        WHEN cv.chf_flag = 1 THEN 'CHF'
        WHEN cv.htn_flag = 1 THEN 'HTN'
        WHEN cv.dm_flag = 1 THEN 'DM'
        WHEN cv.vasc_flag = 1 THEN 'Vascular'
        WHEN cv.age_score = 1 THEN 'Age65-74'
        ELSE 'Unknown'
      END AS driving_component
    INTO #eligible
    FROM #chadsvasc_components cv
    WHERE (
        (cv.SEX = 'M' AND cv.chadsvasc_total = 1)
        OR
        (cv.SEX = 'F' AND cv.chadsvasc_total = 2)
      )
      -- Explicit exclusions (redundant with score but safety checks)
      AND cv.stroke_tia_flag = 0        -- no prior stroke/TIA (would be +2)
      AND cv.age_at_index < 75          -- age >= 75 would be +2
      AND cv.age_at_index >= 18         -- adults only
      -- Exclude active cancer
      AND NOT EXISTS (
        SELECT 1 FROM CDW.dbo.DIAGNOSIS cdx
        WHERE cdx.PATID = cv.PATID
          AND cdx.DX >= 'C00' AND cdx.DX < 'C98'
          AND cdx.DX_TYPE = '10'
          AND cdx.ADMIT_DATE >= DATEADD(year, -1, cv.index_date)
          AND cdx.ADMIT_DATE <= cv.index_date
          AND cdx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      )
      -- No prior OAC at any time (new-user design)
      AND NOT EXISTS (
        SELECT 1 FROM CDW.dbo.PRESCRIBING prx
        WHERE prx.PATID = cv.PATID
          AND prx.RXNORM_CUI IN (
            -- All OAC RXCUIs
            '1364435','1364445','1364441','1364447',
            '1114198','1232082','1232086','2059015',
            '1114202','1232084','1232088','2059017',
            '1037179','1723476','1037045','1037181','1723478','1037049',
            '1599543','1599551','1599555','1599549','1599553','1599557',
            '855350','855288','855302','855312','855318','855324','855332',
            '855338','855344','855296',
            '855290','855304','855314','855320','855326','855334','855340',
            '855346','855298',
            '855292','855306','855316','855322','855328','855336','855342',
            '855348','855300'
          )
          AND COALESCE(prx.RX_START_DATE, prx.RX_ORDER_DATE)
              < DATEADD(day, -", config$grace_period, ", cv.index_date)
          AND COALESCE(prx.RX_START_DATE, prx.RX_ORDER_DATE)
              BETWEEN '2005-01-01' AND GETDATE()
      )
      -- Must have at least one non-legacy encounter in prior 365 days
      AND EXISTS (
        SELECT 1 FROM CDW.dbo.ENCOUNTER enc
        WHERE enc.PATID = cv.PATID
          AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
          AND enc.ADMIT_DATE >= DATEADD(day, -365, cv.index_date)
          AND enc.ADMIT_DATE <= cv.index_date
          AND enc.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      )
    ;

    -- ── CONSORT: capture sub-step counts ──
    IF OBJECT_ID('tempdb..#consort_counts') IS NOT NULL DROP TABLE #consort_counts;
    SELECT
      (SELECT COUNT(DISTINCT PATID) FROM #af_patients)            AS n_af_patients,
      (SELECT COUNT(DISTINCT PATID) FROM #first_af_enc)           AS n_first_af,
      (SELECT COUNT(DISTINCT PATID) FROM #chadsvasc_components)   AS n_scored,
      (SELECT COUNT(DISTINCT PATID) FROM #eligible)               AS n_eligible
    INTO #consort_counts;

    DROP TABLE IF EXISTS #af_patients;
    DROP TABLE IF EXISTS #first_af_enc;
    DROP TABLE IF EXISTS #chadsvasc_components;
  ")

  # ── Step 2: Treatment assignment ──
  # OAC within +/-7 days = treated; no OAC = untreated
  treatment_sql <- paste0("
    -- =========================================================
    -- STEP 2: Treatment Assignment
    -- OAC (treatment=1) vs No OAC (treatment=0)
    -- Grace period: +/-", config$grace_period, " days from index_date
    -- =========================================================

    IF OBJECT_ID('tempdb..#treatment') IS NOT NULL DROP TABLE #treatment;

    SELECT
      e.*,
      CASE WHEN oac.PATID IS NOT NULL THEN 1 ELSE 0 END AS treatment,
      oac.oac_drug AS oac_type,
      YEAR(e.index_date) AS index_year
    INTO #treatment
    FROM #eligible e
    LEFT JOIN (
      SELECT
        p.PATID,
        CASE
          WHEN p.RXNORM_CUI IN ('1364435','1364445','1364441','1364447')
            THEN 'apixaban'
          WHEN p.RXNORM_CUI IN ('1114198','1232082','1232086','2059015',
                                 '1114202','1232084','1232088','2059017')
            THEN 'rivaroxaban'
          WHEN p.RXNORM_CUI IN ('1037179','1723476','1037045',
                                 '1037181','1723478','1037049')
            THEN 'dabigatran'
          WHEN p.RXNORM_CUI IN ('1599543','1599551','1599555',
                                 '1599549','1599553','1599557')
            THEN 'edoxaban'
          ELSE 'warfarin'
        END AS oac_drug,
        ROW_NUMBER() OVER (PARTITION BY p.PATID
                           ORDER BY ABS(DATEDIFF(day,
                             COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE),
                             e2.index_date))) AS rn
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #eligible e2 ON p.PATID = e2.PATID
      WHERE p.RXNORM_CUI IN (
        -- All OAC RXCUIs
        '1364435','1364445','1364441','1364447',
        '1114198','1232082','1232086','2059015',
        '1114202','1232084','1232088','2059017',
        '1037179','1723476','1037045','1037181','1723478','1037049',
        '1599543','1599551','1599555','1599549','1599553','1599557',
        '855350','855288','855302','855312','855318','855324','855332',
        '855338','855344','855296',
        '855290','855304','855314','855320','855326','855334','855340',
        '855346','855298',
        '855292','855306','855316','855322','855328','855336','855342',
        '855348','855300'
      )
      AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
          BETWEEN DATEADD(day, -", config$grace_period, ", e2.index_date)
              AND DATEADD(day,  ", config$grace_period, ", e2.index_date)
    ) oac ON e.PATID = oac.PATID AND oac.rn = 1
    ;
  ")

  # ── Step 3: Outcome ascertainment ──
  outcome_sql <- paste0("
    -- =========================================================
    -- STEP 3: Outcome Ascertainment
    -- Primary: Ischemic stroke / systemic embolism
    -- Secondary: Major bleeding, ICH, all-cause mortality
    -- Follow-up: ", config$followup_days, " days from index_date
    -- =========================================================

    IF OBJECT_ID('tempdb..#outcomes') IS NOT NULL DROP TABLE #outcomes;

    SELECT
      t.PATID,
      t.index_date,
      t.treatment,
      t.oac_type,
      t.index_year,
      t.BIRTH_DATE, t.SEX, t.RACE, t.HISPANIC, t.age_at_index,
      t.chadsvasc_total, t.chf_flag, t.htn_flag, t.dm_flag,
      t.vasc_flag, t.age_score, t.sex_score, t.driving_component,

      -- Primary: Ischemic stroke / systemic embolism
      stroke_se.first_stroke_se_date,
      CASE WHEN stroke_se.first_stroke_se_date IS NOT NULL THEN 1 ELSE 0 END AS stroke_se_event,

      -- Secondary: Major bleeding (all)
      bleed.first_bleed_date,
      CASE WHEN bleed.first_bleed_date IS NOT NULL THEN 1 ELSE 0 END AS bleed_event,

      -- Secondary: Intracranial hemorrhage (subset)
      ich.first_ich_date,
      CASE WHEN ich.first_ich_date IS NOT NULL THEN 1 ELSE 0 END AS ich_event,

      -- Secondary: All-cause mortality
      death.DEATH_DATE AS death_date,
      CASE WHEN death.DEATH_DATE IS NOT NULL
           AND death.DEATH_DATE > t.index_date
           AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
           THEN 1 ELSE 0 END AS death_event,

      -- Time to stroke/SE or censoring
      CASE
        WHEN stroke_se.first_stroke_se_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, stroke_se.first_stroke_se_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_stroke_se,

      -- Time to major bleeding or censoring
      CASE
        WHEN bleed.first_bleed_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, bleed.first_bleed_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_bleed,

      -- Time to ICH or censoring
      CASE
        WHEN ich.first_ich_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, ich.first_ich_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_ich,

      -- Time to death or censoring
      CASE
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_death

    INTO #outcomes
    FROM #treatment t

    -- Stroke / systemic embolism (I63.x, I74.x) on IP/EI/ED
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_stroke_se_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t2 ON dx.PATID = t2.PATID
      WHERE (dx.DX LIKE 'I63%' OR dx.DX LIKE 'I74%')
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t2.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t2.index_date)
      GROUP BY dx.PATID
    ) stroke_se ON t.PATID = stroke_se.PATID

    -- Major bleeding: ICH (I60-I62) + GI (K92.0-K92.2, K62.5)
    --   + ulcer/gastritis hemorrhage
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_bleed_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t3 ON dx.PATID = t3.PATID
      WHERE (dx.DX LIKE 'I60%'    -- subarachnoid hemorrhage
          OR dx.DX LIKE 'I61%'    -- intracerebral hemorrhage
          OR dx.DX LIKE 'I62%'    -- other intracranial hemorrhage
          OR dx.DX IN ('K92.0','K92.1','K92.2','K62.5')  -- GI bleed
          OR dx.DX IN ('K25.0','K25.4','K26.0','K26.4',
                        'K27.0','K27.4','K29.01'))  -- ulcer/gastritis with bleed
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t3.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t3.index_date)
      GROUP BY dx.PATID
    ) bleed ON t.PATID = bleed.PATID

    -- Intracranial hemorrhage (I60-I62 only, for NCB calculation)
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_ich_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t4 ON dx.PATID = t4.PATID
      WHERE (dx.DX LIKE 'I60%' OR dx.DX LIKE 'I61%' OR dx.DX LIKE 'I62%')
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t4.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t4.index_date)
      GROUP BY dx.PATID
    ) ich ON t.PATID = ich.PATID

    -- All-cause mortality (deduplicated; date guard)
    LEFT JOIN (
      SELECT d.PATID, d.DEATH_DATE,
             ROW_NUMBER() OVER (PARTITION BY d.PATID ORDER BY d.DEATH_DATE) AS rn
      FROM CDW.dbo.DEATH d
      WHERE d.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
    ) death ON t.PATID = death.PATID AND death.rn = 1
    ;
  ")

  # ── Step 4: Confounder extraction ──
  confounders_sql <- paste0("
    -- =========================================================
    -- STEP 4: Confounder Extraction
    -- Lookback: ", config$lookback_dx, " days for Dx,
    --           ", config$lookback_lab, " days for labs,
    --           ", config$lookback_vital, " days for vitals,
    --           ", config$lookback_rx, " days for concurrent meds
    -- =========================================================

    IF OBJECT_ID('tempdb..#analytic_cohort') IS NOT NULL DROP TABLE #analytic_cohort;

    SELECT
      o.*,

      -- ===== Comorbidities (lookback ", config$lookback_dx, " days) =====
      -- Note: CHA2DS2-VASc components already captured; these add
      -- non-score comorbidities for confounder adjustment
      ISNULL(comorbid.ckd, 0) AS ckd,
      ISNULL(comorbid.prior_bleed, 0) AS prior_bleed,
      ISNULL(comorbid.liver_disease, 0) AS liver_disease,
      ISNULL(comorbid.anemia, 0) AS anemia,
      ISNULL(comorbid.obesity_dx, 0) AS obesity_dx,
      ISNULL(comorbid.copd, 0) AS copd,
      ISNULL(comorbid.osa, 0) AS osa,
      ISNULL(comorbid.alcohol_use, 0) AS alcohol_use,

      -- ===== Vitals (most recent within ", config$lookback_vital, " days) =====
      vitals.ORIGINAL_BMI AS bmi,
      vitals.SYSTOLIC AS bp_systolic,
      vitals.DIASTOLIC AS bp_diastolic,

      -- ===== Labs (most recent within ", config$lookback_lab, " days) =====
      lab_egfr.RESULT_NUM AS egfr_value,
      lab_cr.RESULT_NUM AS creatinine,
      lab_hba1c.RESULT_NUM AS hba1c,
      lab_hgb.RESULT_NUM AS hemoglobin,

      -- ===== Concurrent medications (within ", config$lookback_rx, " days) =====
      CASE WHEN antiplatelet.PATID IS NOT NULL THEN 1 ELSE 0 END AS antiplatelet_use,
      CASE WHEN statin.PATID IS NOT NULL THEN 1 ELSE 0 END AS statin_use,
      CASE WHEN acei_arb.PATID IS NOT NULL THEN 1 ELSE 0 END AS acei_arb_use,
      CASE WHEN beta_blocker.PATID IS NOT NULL THEN 1 ELSE 0 END AS beta_blocker_use,
      CASE WHEN ppi.PATID IS NOT NULL THEN 1 ELSE 0 END AS ppi_use,
      CASE WHEN nsaid.PATID IS NOT NULL THEN 1 ELSE 0 END AS nsaid_use,

      -- ===== Enrollment =====
      enr.ENR_BASIS

    INTO #analytic_cohort
    FROM #outcomes o

    -- ----- Non-score comorbidities (GROUP BY -> 1 row per PATID) -----
    LEFT JOIN (
      SELECT
        dx.PATID,
        MAX(CASE WHEN dx.DX LIKE 'N18%' THEN 1 ELSE 0 END) AS ckd,
        MAX(CASE WHEN dx.DX LIKE 'I60%' OR dx.DX LIKE 'I61%'
                   OR dx.DX LIKE 'I62%' OR dx.DX LIKE 'K92%'
                   OR dx.DX IN ('K62.5','K25.0','K25.4','K26.0','K26.4',
                                'K27.0','K27.4','K29.01') THEN 1 ELSE 0 END) AS prior_bleed,
        MAX(CASE WHEN dx.DX LIKE 'K70%' OR dx.DX LIKE 'K71%'
                   OR dx.DX LIKE 'K72%' OR dx.DX LIKE 'K73%'
                   OR dx.DX LIKE 'K74%' THEN 1 ELSE 0 END) AS liver_disease,
        MAX(CASE WHEN dx.DX LIKE 'D50%' OR dx.DX LIKE 'D51%'
                   OR dx.DX LIKE 'D52%' OR dx.DX LIKE 'D53%'
                   OR dx.DX LIKE 'D63%' OR dx.DX LIKE 'D64%' THEN 1 ELSE 0 END) AS anemia,
        MAX(CASE WHEN dx.DX LIKE 'E66%' THEN 1 ELSE 0 END) AS obesity_dx,
        MAX(CASE WHEN dx.DX LIKE 'J44%' THEN 1 ELSE 0 END) AS copd,
        MAX(CASE WHEN dx.DX LIKE 'G47.3%' THEN 1 ELSE 0 END) AS osa,
        MAX(CASE WHEN dx.DX LIKE 'F10%' THEN 1 ELSE 0 END) AS alcohol_use
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN #outcomes o2 ON dx.PATID = o2.PATID
      WHERE dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= o2.index_date
        AND dx.ADMIT_DATE >= DATEADD(day, -", config$lookback_dx, ", o2.index_date)
        AND dx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      GROUP BY dx.PATID
    ) comorbid ON o.PATID = comorbid.PATID

    -- ----- Vitals (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ORIGINAL_BMI, SYSTOLIC, DIASTOLIC FROM (
        SELECT v.PATID, v.ORIGINAL_BMI, v.SYSTOLIC, v.DIASTOLIC,
               ROW_NUMBER() OVER (PARTITION BY v.PATID ORDER BY v.MEASURE_DATE DESC) AS rn
        FROM CDW.dbo.VITAL v
        INNER JOIN #outcomes o3 ON v.PATID = o3.PATID
        WHERE v.MEASURE_DATE <= o3.index_date
          AND v.MEASURE_DATE >= DATEADD(day, -", config$lookback_vital, ", o3.index_date)
          AND v.MEASURE_DATE BETWEEN '2005-01-01' AND GETDATE()
      ) v_ranked WHERE rn = 1
    ) vitals ON o.PATID = vitals.PATID

    -- ----- Lab: eGFR (LOINC 48642-3, 62238-1) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o4 ON l.PATID = o4.PATID
        WHERE l.LAB_LOINC IN ('48642-3','62238-1')
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o4.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o4.index_date)
      ) ranked WHERE rn = 1
    ) lab_egfr ON o.PATID = lab_egfr.PATID

    -- ----- Lab: Serum creatinine (LOINC 2160-0) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o5 ON l.PATID = o5.PATID
        WHERE l.LAB_LOINC = '2160-0'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o5.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o5.index_date)
      ) ranked WHERE rn = 1
    ) lab_cr ON o.PATID = lab_cr.PATID

    -- ----- Lab: HbA1c (LOINC 4548-4) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o6 ON l.PATID = o6.PATID
        WHERE l.LAB_LOINC = '4548-4'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o6.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o6.index_date)
      ) ranked WHERE rn = 1
    ) lab_hba1c ON o.PATID = lab_hba1c.PATID

    -- ----- Lab: Hemoglobin (LOINC 718-7) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o7 ON l.PATID = o7.PATID
        WHERE l.LAB_LOINC = '718-7'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o7.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o7.index_date)
      ) ranked WHERE rn = 1
    ) lab_hgb ON o.PATID = lab_hgb.PATID

    -- ----- Concurrent antiplatelet (aspirin, clopidogrel, prasugrel, ticagrelor) -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o9 ON p.PATID = o9.PATID
      WHERE p.RXNORM_CUI IN (
              '243670','308416','212033','198467',
              '309362','213169',
              '855812','855818','855814','855820',
              '1116632','1116635','1116634','1116637'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o9.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o9.index_date
    ) antiplatelet ON o.PATID = antiplatelet.PATID

    -- ----- Concurrent statin -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o10 ON p.PATID = o10.PATID
      WHERE p.RXNORM_CUI IN (
              '259255','617311','617314','617318',
              '859747','859751','859755','859749',
              '198211','314231','152923','312961',
              '312962','312963','433849','312964'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o10.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o10.index_date
    ) statin ON o.PATID = statin.PATID

    -- ----- Concurrent ACEi / ARB -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o11 ON p.PATID = o11.PATID
      WHERE p.RXNORM_CUI IN (
              '314076','314077','197884','314078',
              '979467','979471','979480',
              '349483','349199','349201','349485'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o11.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o11.index_date
    ) acei_arb ON o.PATID = acei_arb.PATID

    -- ----- Concurrent beta-blocker -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o12 ON p.PATID = o12.PATID
      WHERE p.RXNORM_CUI IN (
              '866924','866514','866511',
              '866427','866436','866412','866419',
              '200031','200032','200033','200034'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o12.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o12.index_date
    ) beta_blocker ON o.PATID = beta_blocker.PATID

    -- ----- Concurrent PPI -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o13 ON p.PATID = o13.PATID
      WHERE p.RXNORM_CUI IN (
              '198053','198054',
              '261296','261297',
              '606725','606730'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o13.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o13.index_date
    ) ppi ON o.PATID = ppi.PATID

    -- ----- Concurrent NSAID -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o14 ON p.PATID = o14.PATID
      WHERE p.RXNORM_CUI IN (
              '197805','197806','197807','197808',
              '198013','198014',
              '311461','311462'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o14.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o14.index_date
    ) nsaid ON o.PATID = nsaid.PATID

    -- ----- Enrollment (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ENR_BASIS FROM (
        SELECT enr.PATID, enr.ENR_BASIS,
               ROW_NUMBER() OVER (PARTITION BY enr.PATID ORDER BY enr.ENR_START_DATE DESC) AS rn
        FROM CDW.dbo.ENROLLMENT enr
        INNER JOIN #outcomes o15 ON enr.PATID = o15.PATID
        WHERE o15.index_date BETWEEN enr.ENR_START_DATE
          AND COALESCE(enr.ENR_END_DATE, '9999-12-31')
      ) ranked WHERE rn = 1
    ) enr ON o.PATID = enr.PATID
    ;
  ")

  return(list(
    eligibility  = eligibility_sql,
    treatment    = treatment_sql,
    outcome      = outcome_sql,
    confounders  = confounders_sql
  ))
}
```


## 3. Execute SQL and Pull Data

```{r}
#| label: helper-count

count_temp <- function(con, tbl) {
  res <- dbGetQuery(con, paste("SELECT COUNT(DISTINCT PATID) AS n FROM", tbl))
  res$n[1]
}
```

```{r}
#| label: pull-cohort

pull_analytic_cohort <- function(con, config) {
  sql_parts <- build_cohort_sql(config)
  consort <- list()

  message("=== Executing Cohort SQL ===")

  message("  Step 1: Building eligible population (NVAF + CHA2DS2-VASc target + new to OAC)...")
  dbExecute(con, sql_parts$eligibility)

  consort_raw <- dbGetQuery(con, "SELECT * FROM #consort_counts")
  names(consort_raw) <- tolower(names(consort_raw))
  consort$af_patients   <- consort_raw$n_af_patients[1]
  consort$first_af      <- consort_raw$n_first_af[1]
  consort$scored        <- consort_raw$n_scored[1]
  consort$eligible      <- consort_raw$n_eligible[1]
  try(dbExecute(con, "DROP TABLE IF EXISTS #consort_counts"), silent = TRUE)

  message(sprintf("    -> NVAF patients:          %d", consort$af_patients))
  message(sprintf("    -> With first AF encounter: %d", consort$first_af))
  message(sprintf("    -> Score computed:          %d", consort$scored))
  message(sprintf("    -> Eligible (target score): %d", consort$eligible))

  message("  Step 2: Assigning treatment (OAC vs No OAC)...")
  dbExecute(con, sql_parts$treatment)
  consort$treatment <- count_temp(con, "#treatment")
  message(sprintf("    -> #treatment: %d patients", consort$treatment))

  message("  Step 3: Ascertaining outcomes...")
  dbExecute(con, sql_parts$outcome)
  consort$outcomes <- count_temp(con, "#outcomes")
  message(sprintf("    -> #outcomes: %d patients", consort$outcomes))

  message("  Step 4: Extracting confounders and building analytic cohort...")
  dbExecute(con, sql_parts$confounders)
  cohort <- dbGetQuery(con, "SELECT * FROM #analytic_cohort")

  names(cohort) <- tolower(names(cohort))

  consort$analytic  <- nrow(cohort)
  consort$n_oac     <- sum(cohort$treatment == 1, na.rm = TRUE)
  consort$n_no_oac  <- sum(cohort$treatment == 0, na.rm = TRUE)

  message(sprintf(
    "  Analytic cohort: %d patients (%d OAC, %d No OAC)",
    consort$analytic, consort$n_oac, consort$n_no_oac
  ))

  for (tbl in c("#eligible", "#treatment", "#outcomes", "#analytic_cohort")) {
    try(dbExecute(con, paste("DROP TABLE IF EXISTS", tbl)), silent = TRUE)
  }

  attr(cohort, "consort") <- consort
  return(cohort)
}
```


## 3a. CONSORT Flow Diagram

```{r}
#| label: consort-text

print_consort_table <- function(consort) {
  steps <- tibble::tibble(
    Step = c("1a. NVAF patients (I48.x, non-valvular)",
             "1b. With first AF encounter in study period",
             "1c. CHA2DS2-VASc score computed",
             "1d. Eligible (target score, no prior OAC, no exclusions)",
             "2.  Treatment assigned (OAC vs No OAC)",
             "3.  Outcomes ascertained",
             "4.  Analytic cohort (with confounders)"),
    N    = c(consort$af_patients,
             consort$first_af,
             consort$scored,
             consort$eligible,
             consort$treatment,
             consort$outcomes,
             consort$analytic),
    Excluded = c(NA_integer_,
                 consort$af_patients - consort$first_af,
                 consort$first_af    - consort$scored,
                 consort$scored      - consort$eligible,
                 consort$eligible    - consort$treatment,
                 consort$treatment   - consort$outcomes,
                 consort$outcomes    - consort$analytic)
  )
  message("\n=== CONSORT Flow (text) ===")
  print(steps, n = Inf)
  invisible(steps)
}
```

```{r}
#| label: consort-diagram

render_consort_diagram <- function(consort) {

  grid::grid.newpage()
  fmt <- function(x) format(x, big.mark = ",")

  box_w  <- grid::unit(0.50, "npc")
  box_h  <- grid::unit(0.055, "npc")
  excl_w <- grid::unit(0.28, "npc")
  excl_h <- grid::unit(0.040, "npc")
  x_main <- 0.38
  x_excl <- 0.82

  y_pos <- seq(0.92, 0.12, length.out = 8)

  draw_box <- function(x, y, w, h, label, fill = "white") {
    grid::grid.rect(x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    width = w, height = h,
                    gp = grid::gpar(fill = fill, col = "grey30", lwd = 1.5))
    grid::grid.text(label, x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    gp = grid::gpar(fontsize = 9, fontface = "bold"))
  }

  draw_arrow <- function(x1, y1, x2, y2) {
    grid::grid.lines(x = grid::unit(c(x1, x2), "npc"),
                     y = grid::unit(c(y1, y2), "npc"),
                     arrow = grid::arrow(length = grid::unit(0.015, "npc"),
                                         type = "closed"),
                     gp = grid::gpar(fill = "grey30", col = "grey30"))
  }

  draw_excl <- function(y_from, y_to, n, reason) {
    if (n <= 0) return(invisible(NULL))
    mid_y <- mean(c(y_from - 0.028, y_to + 0.028))
    draw_arrow(x_main + 0.25, mid_y, x_excl - 0.14, mid_y)
    draw_box(x_excl, mid_y, excl_w, excl_h,
             sprintf("Excluded: %s\n(%s)", fmt(n), reason),
             fill = "#FFF3F3")
  }

  grid::grid.text("CONSORT Flow: OAC vs No OAC at Low CHA2DS2-VASc",
                  x = 0.5, y = 0.97,
                  gp = grid::gpar(fontsize = 13, fontface = "bold"))

  labels <- c(
    sprintf("1a. NVAF Patients\nn = %s", fmt(consort$af_patients)),
    sprintf("1b. First AF Encounter\nn = %s", fmt(consort$first_af)),
    sprintf("1c. Score Computed\nn = %s", fmt(consort$scored)),
    sprintf("1d. Eligible (target score)\nn = %s", fmt(consort$eligible)),
    sprintf("2. Treatment Assigned\nn = %s", fmt(consort$treatment)),
    sprintf("3. Outcomes Ascertained\nn = %s", fmt(consort$outcomes)),
    sprintf("4. Analytic Cohort\nn = %s", fmt(consort$analytic))
  )
  fills <- c(rep("white", 6), "#F0FFF0")

  for (i in seq_along(labels)) {
    draw_box(x_main, y_pos[i], box_w, box_h, labels[i], fill = fills[i])
    if (i < length(labels)) {
      draw_arrow(x_main, y_pos[i] - 0.028, x_main, y_pos[i + 1] + 0.028)
    }
  }

  excl_counts  <- c(consort$af_patients - consort$first_af,
                    consort$first_af    - consort$scored,
                    consort$scored      - consort$eligible,
                    consort$eligible    - consort$treatment,
                    consort$treatment   - consort$outcomes,
                    consort$outcomes    - consort$analytic)
  excl_reasons <- c("no AF encounter in period",
                    "demographics/score not computable",
                    "wrong score / exclusions",
                    "all assigned (OAC or not)",
                    "outcome ascertainment",
                    "confounder join")
  for (i in seq_along(excl_counts)) {
    draw_excl(y_pos[i], y_pos[i + 1], excl_counts[i], excl_reasons[i])
  }

  draw_arrow(x_main - 0.10, y_pos[7] - 0.028, x_main - 0.18, y_pos[8] + 0.020)
  draw_arrow(x_main + 0.10, y_pos[7] - 0.028, x_main + 0.18, y_pos[8] + 0.020)

  draw_box(x_main - 0.20, y_pos[8], grid::unit(0.26, "npc"), box_h,
           sprintf("OAC\nn = %s", fmt(consort$n_oac)),
           fill = "#F0F0FF")
  draw_box(x_main + 0.20, y_pos[8], grid::unit(0.26, "npc"), box_h,
           sprintf("No OAC\nn = %s", fmt(consort$n_no_oac)),
           fill = "#F0F0FF")

  invisible(NULL)
}
```


## 4. Data Preparation

```{r}
#| label: prepare-cohort

prepare_cohort <- function(cohort) {
  cohort <- cohort |>
    mutate(
      sex_cat = factor(sex, levels = c("F", "M"),
                       labels = c("Female", "Male")),
      race_cat = case_when(
        race == "03" ~ "Black",
        race == "05" ~ "White",
        race == "02" ~ "Asian",
        race %in% c("01", "04", "06", "07") ~ "Other",
        TRUE ~ "Unknown"
      ) |> factor(),
      hispanic_cat = factor(
        if_else(hispanic == "Y", "Hispanic", "Non-Hispanic")
      ),
      driving_component = factor(driving_component,
                                  levels = c("CHF", "HTN", "DM",
                                             "Vascular", "Age65-74", "Unknown")),
      oac_type_cat = factor(
        case_when(
          is.na(oac_type) ~ "None",
          TRUE ~ oac_type
        )
      ),
      enr_basis_cat = factor(
        case_when(
          enr_basis %in% c("I", "A") ~ enr_basis,
          TRUE ~ "Other/Unknown"
        )
      ),
      treatment = as.integer(treatment),
      treatment_label = factor(
        if_else(treatment == 1, "OAC", "No OAC"),
        levels = c("No OAC", "OAC")
      ),
      # Ensure minimum follow-up time of 1 day
      time_to_stroke_se = pmax(time_to_stroke_se, 1),
      time_to_bleed     = pmax(time_to_bleed, 1),
      time_to_ich       = pmax(time_to_ich, 1),
      time_to_death     = pmax(time_to_death, 1)
    ) |>
    filter(!is.na(treatment))

  # Impute missing continuous covariates with median
  cohort <- cohort |>
    mutate(
      bmi          = if_else(is.na(bmi), median(bmi, na.rm = TRUE), bmi),
      bp_systolic  = if_else(is.na(bp_systolic),
                             median(bp_systolic, na.rm = TRUE), bp_systolic),
      bp_diastolic = if_else(is.na(bp_diastolic),
                             median(bp_diastolic, na.rm = TRUE), bp_diastolic),
      egfr_value   = if_else(is.na(egfr_value),
                             median(egfr_value, na.rm = TRUE), egfr_value),
      creatinine   = if_else(is.na(creatinine),
                             median(creatinine, na.rm = TRUE), creatinine),
      hba1c        = if_else(is.na(hba1c),
                             median(hba1c, na.rm = TRUE), hba1c),
      hemoglobin   = if_else(is.na(hemoglobin),
                             median(hemoglobin, na.rm = TRUE), hemoglobin)
    )

  return(cohort)
}
```


## 5. Causal Analysis Functions (IPW)

```{r}
#| label: confounders

confounders <- c(
  "age_at_index", "sex_cat", "race_cat", "hispanic_cat",
  "bmi", "bp_systolic", "bp_diastolic",
  "egfr_value", "creatinine", "hba1c", "hemoglobin",
  "chf_flag", "htn_flag", "dm_flag", "vasc_flag",
  "ckd", "prior_bleed", "liver_disease", "anemia",
  "obesity_dx", "copd", "osa", "alcohol_use",
  "antiplatelet_use", "statin_use", "acei_arb_use",
  "beta_blocker_use", "ppi_use", "nsaid_use",
  "enr_basis_cat",
  "index_year"
)

build_ps_formula <- function(confounders, cohort) {
  dropped <- character()
  keep    <- character()

  for (v in confounders) {
    col <- cohort[[v]]
    if (is.null(col)) {
      dropped <- c(dropped, v)
      next
    }
    if (is.factor(col) || is.character(col)) {
      n_levels <- length(unique(na.omit(col)))
      if (n_levels < 2) {
        dropped <- c(dropped, v)
        next
      }
    } else {
      if (sd(col, na.rm = TRUE) == 0 || all(is.na(col))) {
        dropped <- c(dropped, v)
        next
      }
    }
    keep <- c(keep, v)
  }

  if (length(dropped) > 0) {
    message(sprintf("  PS model: dropped %d constant/single-level variable(s): %s",
                    length(dropped), paste(dropped, collapse = ", ")))
  }
  message(sprintf("  PS model: using %d confounder(s)", length(keep)))

  as.formula(paste("treatment ~", paste(keep, collapse = " + ")))
}
```

```{r}
#| label: ipw-analysis

run_ipw_analysis <- function(cohort, confounders, config) {
  message("\n=== IPW Analysis ===")

  # --- Guard: treatment must have at least 2 arms ---
  n_arms <- length(unique(cohort$treatment))
  if (n_arms < 2) {
    stop(sprintf(
      "Cannot run IPW: treatment variable has only %d unique value(s) (%s). ",
      n_arms, paste(unique(cohort$treatment), collapse = ", ")),
      "Check the CONSORT diagram -- one treatment arm may have been ",
      "completely filtered out during eligibility.", call. = FALSE)
  }
  message(sprintf("Treatment arms: OAC (n=%d), No OAC (n=%d)",
                  sum(cohort$treatment == 1), sum(cohort$treatment == 0)))

  # --- Build PS formula, dropping single-level factors ---
  ps_formula <- build_ps_formula(confounders, cohort)

  # --- Fit propensity score model and compute weights ---
  message("Fitting propensity score model (logistic regression)...")
  weights <- weightit(
    ps_formula,
    data     = cohort,
    method   = "glm",
    estimand = config$estimand
  )

  # --- Balance diagnostics ---
  message("Checking covariate balance...")
  bal <- bal.tab(weights, stats = c("m", "v"), thresholds = c(m = 0.1))
  print(bal)

  p_love <- love.plot(weights, threshold = 0.1, abs = TRUE,
                      title = "Covariate Balance: OAC vs No OAC (IPW)")

  cohort$ps <- weights$ps
  p_ps <- ggplot(cohort, aes(x = ps, fill = treatment_label)) +
    geom_density(alpha = 0.5) +
    labs(x = "Propensity Score (P[OAC])", y = "Density",
         title = "Propensity Score Distribution by Treatment Group",
         fill = "Treatment") +
    theme_minimal()

  ess_fn <- function(w) sum(w)^2 / sum(w^2)
  message(sprintf("Effective sample size: OAC=%.1f, No OAC=%.1f",
                  ess_fn(weights$weights[cohort$treatment == 1]),
                  ess_fn(weights$weights[cohort$treatment == 0])))

  cohort$ipw <- weights$weights

  # --- Primary: Ischemic stroke / SE ---
  message("\n--- Primary Outcome: Ischemic Stroke / SE ---")
  fit_stroke_se <- coxph(
    Surv(time_to_stroke_se, stroke_se_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_stroke_se))

  # --- Secondary: Major bleeding ---
  message("\n--- Secondary Outcome: Major Bleeding ---")
  fit_bleed <- coxph(
    Surv(time_to_bleed, bleed_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_bleed))

  # --- Secondary: ICH ---
  message("\n--- Secondary Outcome: Intracranial Hemorrhage ---")
  fit_ich <- coxph(
    Surv(time_to_ich, ich_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_ich))

  # --- Secondary: All-cause mortality ---
  message("\n--- Secondary Outcome: All-Cause Mortality ---")
  fit_death <- coxph(
    Surv(time_to_death, death_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_death))

  # --- Weighted KM curves ---
  km_stroke_se <- survfit(
    Surv(time_to_stroke_se, stroke_se_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_stroke_se <- ggsurvplot(
    km_stroke_se, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "Ischemic Stroke/SE: OAC vs No OAC (IPW)",
    xlab = "Days from Time Zero",
    ylab = "Stroke-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  km_bleed <- survfit(
    Surv(time_to_bleed, bleed_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_bleed <- ggsurvplot(
    km_bleed, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "Major Bleeding: OAC vs No OAC (IPW)",
    xlab = "Days from Time Zero",
    ylab = "Bleeding-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  km_death <- survfit(
    Surv(time_to_death, death_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_death <- ggsurvplot(
    km_death, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "All-Cause Mortality: OAC vs No OAC (IPW)",
    xlab = "Days from Time Zero",
    ylab = "Survival Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  return(list(
    weights = weights,
    balance = bal,
    fits = list(
      stroke_se = fit_stroke_se,
      bleed     = fit_bleed,
      ich       = fit_ich,
      death     = fit_death
    ),
    plots = list(
      love       = p_love,
      ps_dist    = p_ps,
      km_stroke  = p_km_stroke_se,
      km_bleed   = p_km_bleed,
      km_death   = p_km_death
    ),
    cohort = cohort
  ))
}
```


## 6. Net Clinical Benefit

```{r}
#| label: net-clinical-benefit

compute_ncb <- function(cohort) {
  message("\n=== Net Clinical Benefit ===")

  # Weighted event rates per 100 person-years
  compute_rate <- function(data, event_col, time_col) {
    events <- sum(data[[event_col]] * data$ipw)
    py     <- sum(data[[time_col]] * data$ipw) / 365.25
    rate   <- events / py * 100
    return(rate)
  }

  oac_data    <- cohort |> filter(treatment == 1)
  no_oac_data <- cohort |> filter(treatment == 0)

  # Stroke/SE rates
  stroke_rate_oac    <- compute_rate(oac_data, "stroke_se_event", "time_to_stroke_se")
  stroke_rate_no_oac <- compute_rate(no_oac_data, "stroke_se_event", "time_to_stroke_se")

  # ICH rates (for NCB calculation with weight = 1.5)
  ich_rate_oac    <- compute_rate(oac_data, "ich_event", "time_to_ich")
  ich_rate_no_oac <- compute_rate(no_oac_data, "ich_event", "time_to_ich")

  # NCB = (stroke reduction) - 1.5 * (ICH excess)
  stroke_reduction <- stroke_rate_no_oac - stroke_rate_oac
  ich_excess       <- ich_rate_oac - ich_rate_no_oac
  ncb              <- stroke_reduction - 1.5 * ich_excess

  results <- tibble::tibble(
    Metric = c(
      "Stroke/SE rate (OAC), per 100 PY",
      "Stroke/SE rate (No OAC), per 100 PY",
      "Stroke reduction (No OAC - OAC)",
      "ICH rate (OAC), per 100 PY",
      "ICH rate (No OAC), per 100 PY",
      "ICH excess (OAC - No OAC)",
      "Net Clinical Benefit (per 100 PY)"
    ),
    Value = round(c(
      stroke_rate_oac,
      stroke_rate_no_oac,
      stroke_reduction,
      ich_rate_oac,
      ich_rate_no_oac,
      ich_excess,
      ncb
    ), 3)
  )

  message(sprintf("  Stroke rate OAC:    %.3f per 100 PY", stroke_rate_oac))
  message(sprintf("  Stroke rate No OAC: %.3f per 100 PY", stroke_rate_no_oac))
  message(sprintf("  Stroke reduction:   %.3f per 100 PY", stroke_reduction))
  message(sprintf("  ICH rate OAC:       %.3f per 100 PY", ich_rate_oac))
  message(sprintf("  ICH rate No OAC:    %.3f per 100 PY", ich_rate_no_oac))
  message(sprintf("  ICH excess:         %.3f per 100 PY", ich_excess))
  message(sprintf("  NCB:                %.3f per 100 PY", ncb))
  message(sprintf("  Interpretation: %s",
                  if_else(ncb > 0,
                          "POSITIVE NCB -- stroke reduction outweighs ICH risk",
                          "NEGATIVE/ZERO NCB -- ICH risk offsets stroke reduction")))

  return(results)
}
```


## 7. Subgroup Analyses

```{r}
#| label: subgroup-analysis

run_subgroup_analyses <- function(cohort, confounders, config) {
  message("\n=== Subgroup Analyses ===")

  subgroups <- list(
    "Driving: CHF"       = cohort$driving_component == "CHF",
    "Driving: HTN"       = cohort$driving_component == "HTN",
    "Driving: DM"        = cohort$driving_component == "DM",
    "Driving: Vascular"  = cohort$driving_component == "Vascular",
    "Driving: Age 65-74" = cohort$driving_component == "Age65-74",
    "Male"               = cohort$sex_cat == "Male",
    "Female"             = cohort$sex_cat == "Female",
    "Age < 70"           = cohort$age_at_index < 70,
    "Age >= 70"          = cohort$age_at_index >= 70,
    "Prior bleeding"     = cohort$prior_bleed == 1,
    "No prior bleeding"  = cohort$prior_bleed == 0
  )

  results <- tibble::tibble(
    Subgroup = character(),
    N = integer(),
    N_OAC = integer(),
    N_NoOAC = integer(),
    N_events = integer(),
    HR = numeric(),
    CI_lo = numeric(),
    CI_hi = numeric(),
    P_value = numeric()
  )

  for (sg_name in names(subgroups)) {
    sg_data <- cohort[subgroups[[sg_name]], ]

    # Guard: need 2 arms and at least 10 per arm
    if (length(unique(sg_data$treatment)) < 2 ||
        sum(sg_data$treatment == 1) < 10 ||
        sum(sg_data$treatment == 0) < 10) {
      message(sprintf("  Skipping '%s': insufficient patients in one arm", sg_name))
      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_OAC = sum(sg_data$treatment == 1),
        N_NoOAC = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$stroke_se_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
      next
    }

    tryCatch({
      ps_formula <- build_ps_formula(confounders, sg_data)
      w <- weightit(ps_formula, data = sg_data, method = "glm",
                    estimand = config$estimand)
      sg_data$ipw <- w$weights

      fit <- coxph(Surv(time_to_stroke_se, stroke_se_event) ~ treatment,
                   data = sg_data, weights = ipw, robust = TRUE)
      s <- summary(fit)

      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name,
        N = nrow(sg_data),
        N_OAC = sum(sg_data$treatment == 1),
        N_NoOAC = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$stroke_se_event),
        HR = s$conf.int[1, 1],
        CI_lo = s$conf.int[1, 3],
        CI_hi = s$conf.int[1, 4],
        P_value = s$coefficients[1, "Pr(>|z|)"]
      ))
      message(sprintf("  %s: HR=%.2f (%.2f-%.2f), N=%d",
                      sg_name, s$conf.int[1,1], s$conf.int[1,3],
                      s$conf.int[1,4], nrow(sg_data)))
    }, error = function(e) {
      message(sprintf("  Error in '%s': %s", sg_name, e$message))
      results <<- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_OAC = sum(sg_data$treatment == 1),
        N_NoOAC = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$stroke_se_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
    })
  }

  return(results)
}
```


## 8. Sensitivity Analyses

```{r}
#| label: sensitivity-analysis

run_sensitivity <- function(ipw_results, cohort, confounders, config) {
  message("\n=== Sensitivity Analyses ===")

  sensitivity <- list()

  # --- 8a. E-value for stroke/SE ---
  message("\n--- E-value Analysis (Stroke/SE) ---")
  fit_stroke <- ipw_results$fits$stroke_se
  s <- summary(fit_stroke)
  hr  <- s$conf.int[1, 1]
  lo  <- s$conf.int[1, 3]
  hi  <- s$conf.int[1, 4]

  evals <- evalues.HR(hr, lo = lo, hi = hi, rare = TRUE)
  sensitivity$evalue_stroke <- evals
  message("E-value for stroke/SE:")
  print(evals)

  # --- 8b. E-value for major bleeding ---
  message("\n--- E-value Analysis (Major Bleeding) ---")
  fit_bleed <- ipw_results$fits$bleed
  s_bleed <- summary(fit_bleed)
  hr_b  <- s_bleed$conf.int[1, 1]
  lo_b  <- s_bleed$conf.int[1, 3]
  hi_b  <- s_bleed$conf.int[1, 4]

  evals_bleed <- evalues.HR(hr_b, lo = lo_b, hi = hi_b, rare = TRUE)
  sensitivity$evalue_bleed <- evals_bleed
  message("E-value for major bleeding:")
  print(evals_bleed)

  # --- 8c. PS trimming (5th-95th percentile) ---
  message("\n--- PS Trimming (5th-95th Percentile) ---")
  ps_lo <- quantile(cohort$ps, 0.05, na.rm = TRUE)
  ps_hi <- quantile(cohort$ps, 0.95, na.rm = TRUE)
  cohort_trimmed <- cohort |>
    filter(ps >= ps_lo & ps <= ps_hi)

  message(sprintf("  Trimmed from %d to %d patients (PS range: %.3f-%.3f)",
                  nrow(cohort), nrow(cohort_trimmed), ps_lo, ps_hi))

  if (length(unique(cohort_trimmed$treatment)) < 2) {
    message("WARNING: After PS trimming, only 1 arm remains. Skipping.")
    sensitivity$trimmed_fit <- NULL
  } else {
    ps_formula_trim <- build_ps_formula(confounders, cohort_trimmed)
    w_trim <- weightit(ps_formula_trim, data = cohort_trimmed,
                       method = "glm", estimand = config$estimand)
    cohort_trimmed$ipw <- w_trim$weights

    fit_trim <- coxph(
      Surv(time_to_stroke_se, stroke_se_event) ~ treatment,
      data = cohort_trimmed, weights = ipw, robust = TRUE
    )
    sensitivity$trimmed_fit <- fit_trim
    message("Trimmed analysis (stroke/SE):")
    print(summary(fit_trim))
  }

  # --- 8d. Post-Epic era only (2021+) ---
  message("\n--- Post-Epic Era (2021+) ---")
  cohort_epic <- cohort |> filter(index_year >= 2021)

  if (length(unique(cohort_epic$treatment)) < 2 ||
      min(table(cohort_epic$treatment)) < 10) {
    message("WARNING: Insufficient patients in post-Epic era. Skipping.")
    sensitivity$epic_fit <- NULL
  } else {
    ps_formula_epic <- build_ps_formula(confounders, cohort_epic)
    w_epic <- weightit(ps_formula_epic, data = cohort_epic,
                       method = "glm", estimand = config$estimand)
    cohort_epic$ipw <- w_epic$weights

    fit_epic <- coxph(
      Surv(time_to_stroke_se, stroke_se_event) ~ treatment,
      data = cohort_epic, weights = ipw, robust = TRUE
    )
    sensitivity$epic_fit <- fit_epic
    message(sprintf("Post-Epic analysis (N=%d):", nrow(cohort_epic)))
    print(summary(fit_epic))
  }

  # --- 8e. DOAC-only OAC arm ---
  message("\n--- DOAC-Only OAC Arm ---")
  cohort_doac <- cohort |>
    filter(treatment == 0 | oac_type_cat %in% c("apixaban", "rivaroxaban",
                                                  "dabigatran", "edoxaban"))

  if (length(unique(cohort_doac$treatment)) < 2 ||
      min(table(cohort_doac$treatment)) < 10) {
    message("WARNING: Insufficient DOAC-only patients. Skipping.")
    sensitivity$doac_fit <- NULL
  } else {
    ps_formula_doac <- build_ps_formula(confounders, cohort_doac)
    w_doac <- weightit(ps_formula_doac, data = cohort_doac,
                       method = "glm", estimand = config$estimand)
    cohort_doac$ipw <- w_doac$weights

    fit_doac <- coxph(
      Surv(time_to_stroke_se, stroke_se_event) ~ treatment,
      data = cohort_doac, weights = ipw, robust = TRUE
    )
    sensitivity$doac_fit <- fit_doac
    message(sprintf("DOAC-only analysis (N=%d):", nrow(cohort_doac)))
    print(summary(fit_doac))
  }

  return(sensitivity)
}
```


## 8a. Summary Helper

```{r}
#| label: summary-helper

format_hr <- function(fit) {
  s <- summary(fit)
  sprintf("HR = %.2f (95%% CI: %.2f-%.2f), p = %.4f",
          s$conf.int[1, 1], s$conf.int[1, 3], s$conf.int[1, 4],
          s$coefficients[1, "Pr(>|z|)"])
}
```


<!-- ============================================================= -->
<!-- PART 2: Execution & Results (Sections 9+)                      -->
<!-- Each section calls its function and displays results inline    -->
<!-- ============================================================= -->


## 9. Connect and Pull Cohort

```{r}
#| label: exec-connect

con <- connect_cdw()
cohort_raw <- pull_analytic_cohort(con, config)
consort <- attr(cohort_raw, "consort")
dbDisconnect(con)
```

### CONSORT Flow (Text)

```{r}
#| label: exec-consort-text

consort_table <- print_consort_table(consort)
knitr::kable(consort_table, caption = "CONSORT Patient Flow")
```

### CONSORT Flow Diagram

```{r}
#| label: fig-consort
#| fig-cap: "CONSORT flow diagram showing patient attrition at each cohort-building step."
#| fig-width: 10
#| fig-height: 12

render_consort_diagram(consort)
```

### Empty Cohort Guard

```{r}
#| label: exec-empty-guard

if (nrow(cohort_raw) == 0) {
  message("*** STOPPING: Analytic cohort has 0 patients. ***")
  message("Review the CONSORT diagram above to identify where patients were lost.")
  knitr::knit_exit()
}
```


## 10. Prepare Data and Table 1

```{r}
#| label: exec-prepare

cohort <- prepare_cohort(cohort_raw)

message(sprintf("Prepared cohort: %d patients (%d OAC, %d No OAC)",
                nrow(cohort),
                sum(cohort$treatment == 1),
                sum(cohort$treatment == 0)))
```

### Table 1: Baseline Characteristics

```{r}
#| label: exec-table1

table1_data <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = n(),
    `Age (mean +/- SD)` = sprintf("%.1f +/- %.1f", mean(age_at_index), sd(age_at_index)),
    `Female (%)` = sprintf("%.1f", 100 * mean(sex_cat == "Female")),
    `White (%)` = sprintf("%.1f", 100 * mean(race_cat == "White")),
    `Black (%)` = sprintf("%.1f", 100 * mean(race_cat == "Black")),
    `Hispanic (%)` = sprintf("%.1f", 100 * mean(hispanic_cat == "Hispanic")),
    `CHA2DS2-VASc (mean)` = sprintf("%.1f", mean(chadsvasc_total)),
    `Driving: CHF (%)` = sprintf("%.1f", 100 * mean(driving_component == "CHF")),
    `Driving: HTN (%)` = sprintf("%.1f", 100 * mean(driving_component == "HTN")),
    `Driving: DM (%)` = sprintf("%.1f", 100 * mean(driving_component == "DM")),
    `Driving: Vascular (%)` = sprintf("%.1f", 100 * mean(driving_component == "Vascular")),
    `Driving: Age 65-74 (%)` = sprintf("%.1f", 100 * mean(driving_component == "Age65-74")),
    `BMI (mean)` = sprintf("%.1f", mean(bmi, na.rm = TRUE)),
    `SBP (mean)` = sprintf("%.0f", mean(bp_systolic, na.rm = TRUE)),
    `eGFR (mean)` = sprintf("%.1f", mean(egfr_value, na.rm = TRUE)),
    `Creatinine (mean)` = sprintf("%.1f", mean(creatinine, na.rm = TRUE)),
    `Hemoglobin (mean)` = sprintf("%.1f", mean(hemoglobin, na.rm = TRUE)),
    `CKD (%)` = sprintf("%.1f", 100 * mean(ckd)),
    `Prior bleeding (%)` = sprintf("%.1f", 100 * mean(prior_bleed)),
    `Liver disease (%)` = sprintf("%.1f", 100 * mean(liver_disease)),
    `Anemia (%)` = sprintf("%.1f", 100 * mean(anemia)),
    `Antiplatelet (%)` = sprintf("%.1f", 100 * mean(antiplatelet_use)),
    `Statin (%)` = sprintf("%.1f", 100 * mean(statin_use)),
    `ACEi/ARB (%)` = sprintf("%.1f", 100 * mean(acei_arb_use)),
    `Beta-blocker (%)` = sprintf("%.1f", 100 * mean(beta_blocker_use)),
    `PPI (%)` = sprintf("%.1f", 100 * mean(ppi_use)),
    `NSAID (%)` = sprintf("%.1f", 100 * mean(nsaid_use)),
    .groups = "drop"
  ) |>
  pivot_longer(-treatment_label, names_to = "Characteristic", values_to = "Value") |>
  pivot_wider(names_from = treatment_label, values_from = Value)

knitr::kable(table1_data, caption = "Table 1: Baseline Characteristics by Treatment Group")
```

### OAC Type Distribution

```{r}
#| label: exec-oac-types

oac_patients <- cohort |> filter(treatment == 1)
oac_dist <- oac_patients |>
  count(oac_type_cat) |>
  mutate(pct = round(100 * n / sum(n), 1))
knitr::kable(oac_dist, caption = "OAC Type Distribution Among Treated Patients")
```

### Event Rates

```{r}
#| label: exec-events

events <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = n(),
    `Stroke/SE events` = sum(stroke_se_event),
    `Major bleeding events` = sum(bleed_event),
    `ICH events` = sum(ich_event),
    `Deaths` = sum(death_event),
    .groups = "drop"
  )
knitr::kable(events, caption = "Outcome Event Counts by Treatment Group")
```


## 11. IPW Analysis

```{r}
#| label: exec-ipw

ipw_results <- run_ipw_analysis(cohort, confounders, config)
cohort <- ipw_results$cohort
```

### Love Plot (Covariate Balance)

```{r}
#| label: fig-love-plot
#| fig-cap: "Absolute standardized mean differences before and after IPW."
#| fig-width: 10
#| fig-height: 8

print(ipw_results$plots$love)
```

### Propensity Score Distribution

```{r}
#| label: fig-ps-dist
#| fig-cap: "Propensity score distributions by treatment group."
#| fig-width: 8
#| fig-height: 5

print(ipw_results$plots$ps_dist)
```


## 12. Cox Model Results and Kaplan-Meier Curves

### Summary of All Outcomes

```{r}
#| label: exec-results-summary

outcome_summary <- tibble::tibble(
  Outcome = c("Ischemic stroke / SE (primary)",
              "Major bleeding",
              "Intracranial hemorrhage",
              "All-cause mortality"),
  Result = c(
    format_hr(ipw_results$fits$stroke_se),
    format_hr(ipw_results$fits$bleed),
    format_hr(ipw_results$fits$ich),
    format_hr(ipw_results$fits$death)
  )
)
knitr::kable(outcome_summary,
             caption = "Weighted Cox PH Results: OAC (treatment=1) vs No OAC (ref)")
```

### Kaplan-Meier: Stroke / SE

```{r}
#| label: fig-km-stroke
#| fig-cap: "Weighted Kaplan-Meier curves for ischemic stroke/systemic embolism."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_stroke)
```

### Kaplan-Meier: Major Bleeding

```{r}
#| label: fig-km-bleed
#| fig-cap: "Weighted Kaplan-Meier curves for major bleeding."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_bleed)
```

### Kaplan-Meier: All-Cause Mortality

```{r}
#| label: fig-km-death
#| fig-cap: "Weighted Kaplan-Meier curves for all-cause mortality."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_death)
```


## 13. Net Clinical Benefit

```{r}
#| label: exec-ncb

ncb_results <- compute_ncb(cohort)
knitr::kable(ncb_results,
             caption = "Net Clinical Benefit: OAC vs No OAC (per 100 person-years)")
```


## 14. Subgroup Analyses

```{r}
#| label: exec-subgroups

subgroup_results <- run_subgroup_analyses(cohort, confounders, config)
knitr::kable(subgroup_results |>
               mutate(across(where(is.numeric), ~round(., 3))),
             caption = "Subgroup Analysis: Stroke/SE Outcome (OAC vs No OAC)")
```

### Forest Plot

```{r}
#| label: fig-forest
#| fig-cap: "Forest plot of subgroup analyses for ischemic stroke/SE."
#| fig-width: 10
#| fig-height: 8

sg_plot_data <- subgroup_results |> filter(!is.na(HR))

if (nrow(sg_plot_data) > 0) {
  ggplot(sg_plot_data, aes(x = HR, y = reorder(Subgroup, HR))) +
    geom_point(size = 3) +
    geom_errorbarh(aes(xmin = CI_lo, xmax = CI_hi), height = 0.2) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
    scale_x_log10() +
    labs(x = "Hazard Ratio (OAC vs No OAC)",
         y = NULL,
         title = "Subgroup Analyses: Ischemic Stroke / SE",
         subtitle = "HR < 1 favors OAC") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 10))
} else {
  message("No subgroup analyses had sufficient data for plotting.")
}
```


## 15. Sensitivity Analyses

```{r}
#| label: exec-sensitivity

sensitivity_results <- run_sensitivity(ipw_results, cohort, confounders, config)
```

### E-value Results

```{r}
#| label: exec-evalue

message("=== E-value: Stroke/SE ===")
print(sensitivity_results$evalue_stroke)

message("\n=== E-value: Major Bleeding ===")
print(sensitivity_results$evalue_bleed)
```

### PS Trimming Results

```{r}
#| label: exec-trimmed

if (!is.null(sensitivity_results$trimmed_fit)) {
  message("Trimmed stroke/SE result:")
  message(format_hr(sensitivity_results$trimmed_fit))
} else {
  message("PS trimming analysis was skipped (insufficient arms).")
}
```

### Post-Epic Era Results

```{r}
#| label: exec-epic

if (!is.null(sensitivity_results$epic_fit)) {
  message("Post-Epic (2021+) stroke/SE result:")
  message(format_hr(sensitivity_results$epic_fit))
} else {
  message("Post-Epic analysis was skipped (insufficient patients).")
}
```

### DOAC-Only OAC Arm Results

```{r}
#| label: exec-doac-only

if (!is.null(sensitivity_results$doac_fit)) {
  message("DOAC-only OAC arm stroke/SE result:")
  message(format_hr(sensitivity_results$doac_fit))
} else {
  message("DOAC-only analysis was skipped (insufficient patients).")
}
```


## 16. Summary

```{r}
#| label: exec-summary

message("\n", strrep("=", 70))
message("PROTOCOL 03 SUMMARY: OAC vs No OAC at Low CHA2DS2-VASc in AF")
message(strrep("=", 70))
message(sprintf("Analytic cohort: %d patients", nrow(cohort)))
message(sprintf("  OAC:    %d", sum(cohort$treatment == 1)))
message(sprintf("  No OAC: %d", sum(cohort$treatment == 0)))
message(sprintf("Follow-up:     %d days", config$followup_days))
message(sprintf("Study period:  %s to %s", config$study_start, config$study_end))
message("")
message("PRIMARY OUTCOME (Ischemic Stroke / SE):")
message(sprintf("  %s", format_hr(ipw_results$fits$stroke_se)))
message("")
message("SECONDARY OUTCOMES:")
message(sprintf("  Major bleeding: %s", format_hr(ipw_results$fits$bleed)))
message(sprintf("  ICH:            %s", format_hr(ipw_results$fits$ich)))
message(sprintf("  All-cause death:%s", format_hr(ipw_results$fits$death)))
message("")
message("NET CLINICAL BENEFIT:")
knitr::kable(ncb_results)
message("")
message("Interpretation: HR < 1 favors OAC; HR > 1 favors No OAC.")
message("Treatment = 1 (OAC) is the 'treated' group in the Cox model.")
message("")
message("CRITICAL CAVEAT: This is a treated-vs-untreated design with")
message("strong confounding by indication. E-values should be reviewed")
message("carefully to assess the plausibility of the causal estimate.")
message(strrep("=", 70))
```
