---
title: "Protocol 01: Apixaban vs Rivaroxaban in AF with CKD Stage 3b-5"
subtitle: "Target Trial Emulation — PCORnet CDW"
author: "Auto-Protocol Designer"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
    theme: cosmo
execute:
  eval: true
  warning: false
  message: true
---

## Study Overview

This protocol emulates a target trial comparing **apixaban vs rivaroxaban** for
stroke prevention in adults with non-valvular atrial fibrillation (AF) and
chronic kidney disease (CKD) stage 3b-5 (eGFR < 45 mL/min/1.73m2).

- **Design:** Active-comparator, new-user cohort
- **Estimand:** ATE (Average Treatment Effect)
- **Method:** Inverse Probability Weighting (IPW) + Weighted Cox PH
- **Time zero:** First DOAC prescription date (index date)
- **Follow-up:** 365 days
- **Study period:** 2016-01-01 to 2025-12-31


<!-- ============================================================= -->
<!-- PART 1: Function definitions (Sections 0-8)                    -->
<!-- These chunks register functions but produce no visible output  -->
<!-- ============================================================= -->


## 0. Setup and Configuration

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(DBI)
library(odbc)
library(WeightIt)
library(cobalt)
library(survival)
library(sandwich)
library(lmtest)
library(EValue)
library(survminer)
```

```{r}
#| label: config

config <- list(
  question        = "Apixaban vs Rivaroxaban in AF with CKD Stage 3b-5",
  exposure_var    = "treatment",
  outcome_var     = "composite_event",
  time_var        = "time_to_composite",
  estimand        = "ATE",
  analysis_method = "ipw",
  study_start     = "2016-01-01",
  study_end       = "2025-12-31",
  followup_days   = 365,
  washout_days    = 365,    # new-user washout for any prior DOAC
  lookback_dx     = 365,    # days for baseline comorbidity lookback
  lookback_lab    = 180,    # days for baseline lab lookback
  lookback_vital  = 365,    # days for baseline vital lookback
  lookback_rx     = 90      # days for concurrent medication lookback
)
```


## 1. Database Connection

```{r}
#| label: db-connect

connect_cdw <- function() {
  con <- DBI::dbConnect(odbc::odbc(), "SQLODBCD17CDM")
  return(con)
}
```


## 2. Cohort SQL

The cohort is built in sequential steps using temp tables:
`#first_doac` -> `#af_patients` -> `#ckd_patients` -> `#eligible` ->
`#treatment` -> `#outcomes` -> `#analytic_cohort`.

All tables are fully qualified as `CDW.dbo.TABLE_NAME`. Medications are
identified by RXNORM_CUI (SCD + SBD), diagnoses by ICD-10 with `DX_TYPE = '10'`.
Legacy encounters are excluded throughout.

```{r}
#| label: cohort-sql

build_cohort_sql <- function(config) {

  # ── Step 1: Eligible population ──
  # AF + CKD 3b-5 patients with new DOAC use
  eligibility_sql <- paste0("
    -- =========================================================
    -- STEP 1: Eligible Population
    -- AF + CKD 3b-5, new apixaban or rivaroxaban users
    -- =========================================================

    -- 1a. Find first apixaban or rivaroxaban prescription per patient
    --     New-user design: no prior DOAC in ", config$washout_days, " days
    IF OBJECT_ID('tempdb..#first_doac_raw') IS NOT NULL DROP TABLE #first_doac_raw;

    -- Step 1: Find earliest apixaban/rivaroxaban Rx per patient
    SELECT
      p.PATID,
      MIN(COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)) AS index_date
    INTO #first_doac_raw
    FROM CDW.dbo.PRESCRIBING p
    WHERE p.RXNORM_CUI IN (
        -- Apixaban (SCD + SBD)
        '1364435','1364445','1364441','1364447',
        -- Rivaroxaban (SCD + SBD)
        '1114198','1232082','1232086','2059015',
        '1114202','1232084','1232088','2059017'
      )
      AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
          BETWEEN '", config$study_start, "' AND '", config$study_end, "'
    GROUP BY p.PATID
    ;

    -- Step 2: Apply new-user washout (no prior DOAC in ", config$washout_days, " days)
    IF OBJECT_ID('tempdb..#first_doac') IS NOT NULL DROP TABLE #first_doac;

    SELECT f.PATID, f.index_date
    INTO #first_doac
    FROM #first_doac_raw f
    WHERE NOT EXISTS (
      SELECT 1 FROM CDW.dbo.PRESCRIBING p2
      WHERE p2.PATID = f.PATID
        AND p2.RXNORM_CUI IN (
          -- All DOACs: apixaban, rivaroxaban, dabigatran, edoxaban
          '1364435','1364445','1364441','1364447',
          '1114198','1232082','1232086','2059015',
          '1114202','1232084','1232088','2059017',
          '1037179','1723476','1037045','1037181','1723478','1037049',
          '1599543','1599551','1599555','1599549','1599553','1599557'
        )
        AND COALESCE(p2.RX_START_DATE, p2.RX_ORDER_DATE)
            >= DATEADD(day, -", config$washout_days, ", f.index_date)
        AND COALESCE(p2.RX_START_DATE, p2.RX_ORDER_DATE)
            < f.index_date
    )
    ;
    DROP TABLE IF EXISTS #first_doac_raw;

    -- 1b. Require AF diagnosis on or before index date
    IF OBJECT_ID('tempdb..#af_patients') IS NOT NULL DROP TABLE #af_patients;

    SELECT fd.PATID, fd.index_date
    INTO #af_patients
    FROM #first_doac fd
    WHERE EXISTS (
      SELECT 1 FROM CDW.dbo.DIAGNOSIS dx
      WHERE dx.PATID = fd.PATID
        AND dx.DX LIKE 'I48%'
        AND dx.DX NOT IN ('I48.3','I48.4','I48.92')  -- exclude flutter
        AND dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= fd.index_date
        AND dx.ADMIT_DATE BETWEEN '", config$study_start, "' AND GETDATE()
    )
    -- Exclude valvular AF
    AND NOT EXISTS (
      SELECT 1 FROM CDW.dbo.DIAGNOSIS vdx
      WHERE vdx.PATID = fd.PATID
        AND (vdx.DX LIKE 'I05%' OR vdx.DX LIKE 'I06%'
             OR vdx.DX LIKE 'I07%' OR vdx.DX LIKE 'I08%'
             OR vdx.DX IN ('Z95.2','Z95.3','Z95.4'))
        AND vdx.DX_TYPE = '10'
        AND vdx.ADMIT_DATE <= fd.index_date
        AND vdx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
    )
    ;

    -- 1c. Require CKD 3b-5 (dual ascertainment: Dx OR eGFR < 45)
    IF OBJECT_ID('tempdb..#ckd_patients') IS NOT NULL DROP TABLE #ckd_patients;

    SELECT af.PATID, af.index_date
    INTO #ckd_patients
    FROM #af_patients af
    WHERE (
      -- Strategy 1: Diagnosis-based CKD 3b-5
      EXISTS (
        SELECT 1 FROM CDW.dbo.DIAGNOSIS d
        WHERE d.PATID = af.PATID
          AND d.DX IN ('N18.32','N18.4','N18.5')
          AND d.DX_TYPE = '10'
          AND d.ADMIT_DATE <= af.index_date
          AND d.ADMIT_DATE BETWEEN '", config$study_start, "' AND GETDATE()
      )
      OR
      -- Strategy 2: Lab-based eGFR < 45
      EXISTS (
        SELECT 1 FROM CDW.dbo.LAB_RESULT_CM l
        WHERE l.PATID = af.PATID
          AND l.LAB_LOINC IN ('48642-3','62238-1','33914-3','98979-8')
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_NUM > 0 AND l.RESULT_NUM < 45
          AND l.RESULT_DATE <= af.index_date
          AND l.RESULT_DATE >= DATEADD(day, -180, af.index_date)
          AND l.RESULT_DATE BETWEEN '", config$study_start, "' AND GETDATE()
      )
    )
    -- Exclude ESRD / dialysis
    AND NOT EXISTS (
      SELECT 1 FROM CDW.dbo.DIAGNOSIS ed
      WHERE ed.PATID = af.PATID
        AND ed.DX IN ('N18.6','Z99.2')
        AND ed.DX_TYPE = '10'
        AND ed.ADMIT_DATE <= af.index_date
        AND ed.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
    )
    AND NOT EXISTS (
      SELECT 1 FROM CDW.dbo.PROCEDURES pr
      WHERE pr.PATID = af.PATID
        AND pr.PX IN ('90935','90937','90940','90945','90947')
        AND pr.PX_TYPE = 'CH'
        AND pr.PX_DATE <= af.index_date
        AND pr.PX_DATE BETWEEN '2005-01-01' AND GETDATE()
    )
    ;

    -- 1d. Apply demographics and age filter
    IF OBJECT_ID('tempdb..#eligible') IS NOT NULL DROP TABLE #eligible;

    SELECT
      ck.PATID,
      ck.index_date,
      d.BIRTH_DATE,
      d.SEX,
      d.RACE,
      d.HISPANIC,
      DATEDIFF(year, d.BIRTH_DATE, ck.index_date) AS age_at_index
    INTO #eligible
    FROM #ckd_patients ck
    INNER JOIN CDW.dbo.DEMOGRAPHIC d ON ck.PATID = d.PATID
    WHERE DATEDIFF(year, d.BIRTH_DATE, ck.index_date) >= 18
      -- Must have at least one non-legacy encounter in prior 365 days
      AND EXISTS (
        SELECT 1 FROM CDW.dbo.ENCOUNTER e
        WHERE e.PATID = ck.PATID
          AND e.RAW_ENC_TYPE <> 'Legacy Encounter'
          AND e.ADMIT_DATE >= DATEADD(day, -365, ck.index_date)
          AND e.ADMIT_DATE <= ck.index_date
          AND e.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      )
    ;

    -- ── CONSORT: capture sub-step counts ──
    IF OBJECT_ID('tempdb..#consort_counts') IS NOT NULL DROP TABLE #consort_counts;
    SELECT
      (SELECT COUNT(DISTINCT PATID) FROM #first_doac)    AS n_first_doac,
      (SELECT COUNT(DISTINCT PATID) FROM #af_patients)   AS n_af_patients,
      (SELECT COUNT(DISTINCT PATID) FROM #ckd_patients)  AS n_ckd_patients,
      (SELECT COUNT(DISTINCT PATID) FROM #eligible)      AS n_eligible
    INTO #consort_counts;

    DROP TABLE IF EXISTS #first_doac;
    DROP TABLE IF EXISTS #af_patients;
    DROP TABLE IF EXISTS #ckd_patients;
  ")

  # ── Step 2: Treatment assignment ──
  treatment_sql <- paste0("
    -- =========================================================
    -- STEP 2: Treatment Assignment
    -- Apixaban (treatment=1) vs Rivaroxaban (treatment=0)
    -- =========================================================

    IF OBJECT_ID('tempdb..#treatment') IS NOT NULL DROP TABLE #treatment;

    SELECT
      e.PATID,
      e.index_date,
      e.BIRTH_DATE, e.SEX, e.RACE, e.HISPANIC, e.age_at_index,
      -- Assign treatment based on which DOAC was prescribed at index
      CASE
        WHEN apix.PATID IS NOT NULL THEN 1  -- Apixaban
        WHEN riva.PATID IS NOT NULL THEN 0  -- Rivaroxaban
        ELSE NULL  -- should not happen (both or neither)
      END AS treatment,
      -- Dose detail for subgroup analyses
      CASE
        WHEN apix.RXNORM_CUI IN ('1364445','1364447') THEN 'apixaban_5mg'
        WHEN apix.RXNORM_CUI IN ('1364435','1364441') THEN 'apixaban_2.5mg'
        WHEN riva.RXNORM_CUI IN ('1232086','1232088') THEN 'rivaroxaban_20mg'
        WHEN riva.RXNORM_CUI IN ('1232082','1232084') THEN 'rivaroxaban_15mg'
        WHEN riva.RXNORM_CUI IN ('1114198','1114202') THEN 'rivaroxaban_10mg'
        WHEN riva.RXNORM_CUI IN ('2059015','2059017') THEN 'rivaroxaban_2.5mg'
        ELSE 'unknown'
      END AS dose_group,
      -- Calendar year
      YEAR(e.index_date) AS index_year
    INTO #treatment
    FROM #eligible e
    -- Apixaban at index (first Rx on index date)
    LEFT JOIN (
      SELECT p.PATID, p.RXNORM_CUI,
             ROW_NUMBER() OVER (PARTITION BY p.PATID
                                ORDER BY COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)) AS rn
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #eligible e2 ON p.PATID = e2.PATID
      WHERE p.RXNORM_CUI IN ('1364435','1364445','1364441','1364447')
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) = e2.index_date
    ) apix ON e.PATID = apix.PATID AND apix.rn = 1
    -- Rivaroxaban at index (first Rx on index date)
    LEFT JOIN (
      SELECT p.PATID, p.RXNORM_CUI,
             ROW_NUMBER() OVER (PARTITION BY p.PATID
                                ORDER BY COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)) AS rn
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #eligible e3 ON p.PATID = e3.PATID
      WHERE p.RXNORM_CUI IN ('1114198','1232082','1232086','2059015',
                              '1114202','1232084','1232088','2059017')
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) = e3.index_date
    ) riva ON e.PATID = riva.PATID AND riva.rn = 1
    WHERE COALESCE(apix.PATID, riva.PATID) IS NOT NULL
    ;

    -- Remove ambiguous (both prescribed on same day) or missing
    DELETE FROM #treatment WHERE treatment IS NULL;
  ")

  # ── Step 3: Outcome ascertainment ──
  outcome_sql <- paste0("
    -- =========================================================
    -- STEP 3: Outcome Ascertainment
    -- Primary: Composite stroke/SE + major bleeding
    -- Secondary: Stroke/SE, major bleeding, GI bleeding, mortality
    -- Follow-up: ", config$followup_days, " days from index_date
    -- =========================================================

    IF OBJECT_ID('tempdb..#outcomes') IS NOT NULL DROP TABLE #outcomes;

    SELECT
      t.PATID,
      t.index_date,
      t.treatment,
      t.dose_group,
      t.index_year,
      t.BIRTH_DATE, t.SEX, t.RACE, t.HISPANIC, t.age_at_index,

      -- Secondary: Stroke / systemic embolism
      stroke.first_stroke_date,
      CASE WHEN stroke.first_stroke_date IS NOT NULL THEN 1 ELSE 0 END AS stroke_event,

      -- Secondary: Major bleeding
      bleed.first_bleed_date,
      CASE WHEN bleed.first_bleed_date IS NOT NULL THEN 1 ELSE 0 END AS bleed_event,

      -- Secondary: GI bleeding (subset)
      gi_bleed.first_gi_date,
      CASE WHEN gi_bleed.first_gi_date IS NOT NULL THEN 1 ELSE 0 END AS gi_bleed_event,

      -- Secondary: All-cause mortality
      death.DEATH_DATE AS death_date,
      CASE WHEN death.DEATH_DATE IS NOT NULL
           AND death.DEATH_DATE > t.index_date
           AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
           THEN 1 ELSE 0 END AS death_event,

      -- Primary composite: first of stroke/SE or major bleeding
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL OR bleed.first_bleed_date IS NOT NULL
        THEN 1 ELSE 0
      END AS composite_event,

      -- Time to stroke/SE or censoring
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, stroke.first_stroke_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_stroke,

      -- Time to major bleeding or censoring
      CASE
        WHEN bleed.first_bleed_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, bleed.first_bleed_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_bleed,

      -- Time to GI bleeding or censoring
      CASE
        WHEN gi_bleed.first_gi_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, gi_bleed.first_gi_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_gi_bleed,

      -- Time to death or censoring
      CASE
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_death,

      -- Time to composite (first of stroke/SE or major bleeding)
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL AND bleed.first_bleed_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date,
               CASE WHEN stroke.first_stroke_date <= bleed.first_bleed_date
                    THEN stroke.first_stroke_date ELSE bleed.first_bleed_date END)
        WHEN stroke.first_stroke_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, stroke.first_stroke_date)
        WHEN bleed.first_bleed_date IS NOT NULL
          THEN DATEDIFF(day, t.index_date, bleed.first_bleed_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > t.index_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", t.index_date)
          THEN DATEDIFF(day, t.index_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_composite

    INTO #outcomes
    FROM #treatment t

    -- Stroke / systemic embolism (I63.x, I74.x) in IP/ED/EI
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_stroke_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t2 ON dx.PATID = t2.PATID
      WHERE (dx.DX LIKE 'I63%' OR dx.DX LIKE 'I74%')
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t2.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t2.index_date)
      GROUP BY dx.PATID
    ) stroke ON t.PATID = stroke.PATID

    -- Major bleeding: ICH (I60-I62) + GI (K92.0, K92.1, K92.2, K62.5)
    --   + ulcer hemorrhage + gastritis with bleeding
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_bleed_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t3 ON dx.PATID = t3.PATID
      WHERE (dx.DX LIKE 'I60%'    -- subarachnoid hemorrhage
          OR dx.DX LIKE 'I61%'    -- intracerebral hemorrhage
          OR dx.DX LIKE 'I62%'    -- other intracranial hemorrhage
          OR dx.DX IN ('K92.0','K92.1','K92.2','K62.5')  -- GI bleed
          OR dx.DX IN ('K25.0','K25.4','K26.0','K26.4',
                        'K27.0','K27.4','K29.01'))  -- ulcer/gastritis with bleed
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t3.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t3.index_date)
      GROUP BY dx.PATID
    ) bleed ON t.PATID = bleed.PATID

    -- GI bleeding (subset)
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_gi_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #treatment t4 ON dx.PATID = t4.PATID
      WHERE dx.DX IN ('K92.0','K92.1','K92.2','K62.5',
                       'K25.0','K25.4','K26.0','K26.4',
                       'K27.0','K27.4','K29.01')
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI','ED')
        AND dx.ADMIT_DATE > t4.index_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", t4.index_date)
      GROUP BY dx.PATID
    ) gi_bleed ON t.PATID = gi_bleed.PATID

    -- All-cause mortality (deduplicated; date guard for junk dates)
    LEFT JOIN (
      SELECT d.PATID, d.DEATH_DATE,
             ROW_NUMBER() OVER (PARTITION BY d.PATID ORDER BY d.DEATH_DATE) AS rn
      FROM CDW.dbo.DEATH d
      WHERE d.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
    ) death ON t.PATID = death.PATID AND death.rn = 1
    ;
  ")

  # ── Step 4: Confounder extraction ──
  confounders_sql <- paste0("
    -- =========================================================
    -- STEP 4: Confounder Extraction
    -- Lookback: ", config$lookback_dx, " days for Dx,
    --           ", config$lookback_lab, " days for labs,
    --           ", config$lookback_vital, " days for vitals,
    --           ", config$lookback_rx, " days for concurrent meds
    -- =========================================================

    IF OBJECT_ID('tempdb..#analytic_cohort') IS NOT NULL DROP TABLE #analytic_cohort;

    SELECT
      o.*,

      -- ===== Comorbidities (lookback ", config$lookback_dx, " days) =====
      ISNULL(comorbid.chf, 0) AS chf,
      ISNULL(comorbid.hypertension, 0) AS hypertension,
      ISNULL(comorbid.diabetes, 0) AS diabetes,
      ISNULL(comorbid.prior_stroke, 0) AS prior_stroke,
      ISNULL(comorbid.prior_embolism, 0) AS prior_embolism,
      ISNULL(comorbid.vascular_disease, 0) AS vascular_disease,
      ISNULL(comorbid.prior_bleed, 0) AS prior_bleed,
      ISNULL(comorbid.liver_disease, 0) AS liver_disease,
      ISNULL(comorbid.anemia, 0) AS anemia,
      ISNULL(comorbid.obesity_dx, 0) AS obesity_dx,
      ISNULL(comorbid.cancer, 0) AS cancer,
      -- CKD stage (for subgroup analyses)
      CASE
        WHEN ckd_stage.ckd_5 = 1 THEN 'CKD5'
        WHEN ckd_stage.ckd_4 = 1 THEN 'CKD4'
        ELSE 'CKD3b'
      END AS ckd_stage_cat,

      -- ===== Vitals (most recent within ", config$lookback_vital, " days) =====
      vitals.ORIGINAL_BMI AS bmi,
      vitals.SYSTOLIC AS bp_systolic,
      vitals.DIASTOLIC AS bp_diastolic,

      -- ===== Labs (most recent within ", config$lookback_lab, " days) =====
      lab_egfr.RESULT_NUM AS egfr_value,
      lab_cr.RESULT_NUM AS creatinine,
      lab_hba1c.RESULT_NUM AS hba1c,
      lab_hgb.RESULT_NUM AS hemoglobin,
      lab_plt.RESULT_NUM AS platelets,

      -- ===== Concurrent medications (within ", config$lookback_rx, " days) =====
      CASE WHEN antiplatelet.PATID IS NOT NULL THEN 1 ELSE 0 END AS antiplatelet_use,
      CASE WHEN statin.PATID IS NOT NULL THEN 1 ELSE 0 END AS statin_use,
      CASE WHEN acei_arb.PATID IS NOT NULL THEN 1 ELSE 0 END AS acei_arb_use,
      CASE WHEN beta_blocker.PATID IS NOT NULL THEN 1 ELSE 0 END AS beta_blocker_use,
      CASE WHEN ppi.PATID IS NOT NULL THEN 1 ELSE 0 END AS ppi_use,
      CASE WHEN nsaid.PATID IS NOT NULL THEN 1 ELSE 0 END AS nsaid_use,

      -- ===== Enrollment =====
      enr.ENR_BASIS

    INTO #analytic_cohort
    FROM #outcomes o

    -- ----- Comorbidities (GROUP BY -> safe, 1 row per PATID) -----
    LEFT JOIN (
      SELECT
        dx.PATID,
        MAX(CASE WHEN dx.DX LIKE 'I50%' THEN 1 ELSE 0 END) AS chf,
        MAX(CASE WHEN dx.DX LIKE 'I10%' OR dx.DX LIKE 'I11%'
                   OR dx.DX LIKE 'I12%' OR dx.DX LIKE 'I13%'
                   OR dx.DX LIKE 'I15%' OR dx.DX LIKE 'I16%' THEN 1 ELSE 0 END) AS hypertension,
        MAX(CASE WHEN dx.DX LIKE 'E10%' OR dx.DX LIKE 'E11%'
                   OR dx.DX LIKE 'E13%' THEN 1 ELSE 0 END) AS diabetes,
        MAX(CASE WHEN dx.DX LIKE 'I63%' OR dx.DX LIKE 'G45%' THEN 1 ELSE 0 END) AS prior_stroke,
        MAX(CASE WHEN dx.DX LIKE 'I74%' THEN 1 ELSE 0 END) AS prior_embolism,
        MAX(CASE WHEN dx.DX LIKE 'I21%' OR dx.DX LIKE 'I25%'
                   OR dx.DX LIKE 'I70%' THEN 1 ELSE 0 END) AS vascular_disease,
        MAX(CASE WHEN dx.DX LIKE 'I60%' OR dx.DX LIKE 'I61%'
                   OR dx.DX LIKE 'I62%' OR dx.DX LIKE 'K92%'
                   OR dx.DX IN ('K62.5','K25.0','K25.4','K26.0','K26.4',
                                'K27.0','K27.4','K29.01') THEN 1 ELSE 0 END) AS prior_bleed,
        MAX(CASE WHEN dx.DX LIKE 'K70%' OR dx.DX LIKE 'K71%'
                   OR dx.DX LIKE 'K72%' OR dx.DX LIKE 'K73%'
                   OR dx.DX LIKE 'K74%' THEN 1 ELSE 0 END) AS liver_disease,
        MAX(CASE WHEN dx.DX LIKE 'D50%' OR dx.DX LIKE 'D51%'
                   OR dx.DX LIKE 'D52%' OR dx.DX LIKE 'D53%'
                   OR dx.DX LIKE 'D63%' OR dx.DX LIKE 'D64%' THEN 1 ELSE 0 END) AS anemia,
        MAX(CASE WHEN dx.DX LIKE 'E66%' THEN 1 ELSE 0 END) AS obesity_dx,
        MAX(CASE WHEN dx.DX >= 'C00' AND dx.DX < 'C98' THEN 1 ELSE 0 END) AS cancer
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN #outcomes o2 ON dx.PATID = o2.PATID
      WHERE dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= o2.index_date
        AND dx.ADMIT_DATE >= DATEADD(day, -", config$lookback_dx, ", o2.index_date)
        AND dx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      GROUP BY dx.PATID
    ) comorbid ON o.PATID = comorbid.PATID

    -- ----- CKD stage detail -----
    LEFT JOIN (
      SELECT
        dx.PATID,
        MAX(CASE WHEN dx.DX = 'N18.5' THEN 1 ELSE 0 END) AS ckd_5,
        MAX(CASE WHEN dx.DX = 'N18.4' THEN 1 ELSE 0 END) AS ckd_4
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN #outcomes o_ckd ON dx.PATID = o_ckd.PATID
      WHERE dx.DX IN ('N18.32','N18.4','N18.5')
        AND dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= o_ckd.index_date
        AND dx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      GROUP BY dx.PATID
    ) ckd_stage ON o.PATID = ckd_stage.PATID

    -- ----- Vitals (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ORIGINAL_BMI, SYSTOLIC, DIASTOLIC FROM (
        SELECT v.PATID, v.ORIGINAL_BMI, v.SYSTOLIC, v.DIASTOLIC,
               ROW_NUMBER() OVER (PARTITION BY v.PATID ORDER BY v.MEASURE_DATE DESC) AS rn
        FROM CDW.dbo.VITAL v
        INNER JOIN #outcomes o3 ON v.PATID = o3.PATID
        WHERE v.MEASURE_DATE <= o3.index_date
          AND v.MEASURE_DATE >= DATEADD(day, -", config$lookback_vital, ", o3.index_date)
          AND v.MEASURE_DATE BETWEEN '2005-01-01' AND GETDATE()
      ) v_ranked WHERE rn = 1
    ) vitals ON o.PATID = vitals.PATID

    -- ----- Lab: eGFR (LOINC 48642-3, 62238-1) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o4 ON l.PATID = o4.PATID
        WHERE l.LAB_LOINC IN ('48642-3','62238-1')
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o4.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o4.index_date)
      ) ranked WHERE rn = 1
    ) lab_egfr ON o.PATID = lab_egfr.PATID

    -- ----- Lab: Serum creatinine (LOINC 2160-0) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o5 ON l.PATID = o5.PATID
        WHERE l.LAB_LOINC = '2160-0'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o5.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o5.index_date)
      ) ranked WHERE rn = 1
    ) lab_cr ON o.PATID = lab_cr.PATID

    -- ----- Lab: HbA1c (LOINC 4548-4) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o6 ON l.PATID = o6.PATID
        WHERE l.LAB_LOINC = '4548-4'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o6.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o6.index_date)
      ) ranked WHERE rn = 1
    ) lab_hba1c ON o.PATID = lab_hba1c.PATID

    -- ----- Lab: Hemoglobin (LOINC 718-7) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o7 ON l.PATID = o7.PATID
        WHERE l.LAB_LOINC = '718-7'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o7.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o7.index_date)
      ) ranked WHERE rn = 1
    ) lab_hgb ON o.PATID = lab_hgb.PATID

    -- ----- Lab: Platelets (LOINC 777-3) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o8 ON l.PATID = o8.PATID
        WHERE l.LAB_LOINC = '777-3'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o8.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o8.index_date)
      ) ranked WHERE rn = 1
    ) lab_plt ON o.PATID = lab_plt.PATID

    -- ----- Concurrent antiplatelet (aspirin, clopidogrel, prasugrel, ticagrelor) -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o9 ON p.PATID = o9.PATID
      WHERE p.RXNORM_CUI IN (
              -- aspirin: representative oral SCD CUIs
              '243670','308416','212033','198467',
              -- clopidogrel 75mg (SCD + SBD Plavix)
              '309362','213169',
              -- prasugrel 5mg, 10mg (SCD + SBD Effient)
              '855812','855818','855814','855820',
              -- ticagrelor 60mg, 90mg (SCD + SBD Brilinta)
              '1116632','1116635','1116634','1116637'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o9.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o9.index_date
    ) antiplatelet ON o.PATID = antiplatelet.PATID

    -- ----- Concurrent statin -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o10 ON p.PATID = o10.PATID
      WHERE p.RXNORM_CUI IN (
              -- atorvastatin 10/20/40/80 (SCD)
              '259255','617311','617314','617318',
              -- rosuvastatin 5/10/20/40 (SCD)
              '859747','859751','859755','859749',
              -- simvastatin 10/20/40/80 (SCD)
              '198211','314231','152923','312961',
              -- pravastatin 10/20/40/80 (SCD)
              '312962','312963','433849','312964'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o10.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o10.index_date
    ) statin ON o.PATID = statin.PATID

    -- ----- Concurrent ACEi / ARB -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o11 ON p.PATID = o11.PATID
      WHERE p.RXNORM_CUI IN (
              -- lisinopril 5/10/20/40 (SCD)
              '314076','314077','197884','314078',
              -- losartan 25/50/100 (SCD)
              '979467','979471','979480',
              -- valsartan 40/80/160/320 (SCD)
              '349483','349199','349201','349485'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o11.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o11.index_date
    ) acei_arb ON o.PATID = acei_arb.PATID

    -- ----- Concurrent beta-blocker -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o12 ON p.PATID = o12.PATID
      WHERE p.RXNORM_CUI IN (
              -- metoprolol tartrate 25/50/100 (SCD)
              '866924','866514','866511',
              -- metoprolol succinate ER 25/50/100/200 (SCD)
              '866427','866436','866412','866419',
              -- carvedilol 3.125/6.25/12.5/25 (SCD)
              '200031','200032','200033','200034'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o12.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o12.index_date
    ) beta_blocker ON o.PATID = beta_blocker.PATID

    -- ----- Concurrent PPI -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o13 ON p.PATID = o13.PATID
      WHERE p.RXNORM_CUI IN (
              -- omeprazole 20/40 (SCD)
              '198053','198054',
              -- pantoprazole 20/40 (SCD)
              '261296','261297',
              -- esomeprazole 20/40 (SCD)
              '606725','606730'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o13.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o13.index_date
    ) ppi ON o.PATID = ppi.PATID

    -- ----- Concurrent NSAID -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o14 ON p.PATID = o14.PATID
      WHERE p.RXNORM_CUI IN (
              -- ibuprofen 200/400/600/800 (SCD)
              '197805','197806','197807','197808',
              -- naproxen 250/500 (SCD)
              '198013','198014',
              -- meloxicam 7.5/15 (SCD)
              '311461','311462'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o14.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o14.index_date
    ) nsaid ON o.PATID = nsaid.PATID

    -- ----- Enrollment (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ENR_BASIS FROM (
        SELECT enr.PATID, enr.ENR_BASIS,
               ROW_NUMBER() OVER (PARTITION BY enr.PATID ORDER BY enr.ENR_START_DATE DESC) AS rn
        FROM CDW.dbo.ENROLLMENT enr
        INNER JOIN #outcomes o15 ON enr.PATID = o15.PATID
        WHERE o15.index_date BETWEEN enr.ENR_START_DATE
          AND COALESCE(enr.ENR_END_DATE, '9999-12-31')
      ) ranked WHERE rn = 1
    ) enr ON o.PATID = enr.PATID
    ;
  ")

  return(list(
    eligibility  = eligibility_sql,
    treatment    = treatment_sql,
    outcome      = outcome_sql,
    confounders  = confounders_sql
  ))
}
```


## 3. Execute SQL and Pull Data

```{r}
#| label: helper-count

count_temp <- function(con, tbl) {
  res <- dbGetQuery(con, paste("SELECT COUNT(DISTINCT PATID) AS n FROM", tbl))
  res$n[1]
}
```

```{r}
#| label: pull-cohort

pull_analytic_cohort <- function(con, config) {
  sql_parts <- build_cohort_sql(config)
  consort <- list()

  message("=== Executing Cohort SQL ===")

  message("  Step 1: Building eligible population (AF + CKD 3b-5 + new DOAC)...")
  dbExecute(con, sql_parts$eligibility)

  consort_raw <- dbGetQuery(con, "SELECT * FROM #consort_counts")
  names(consort_raw) <- tolower(names(consort_raw))
  consort$first_doac    <- consort_raw$n_first_doac[1]
  consort$af_patients   <- consort_raw$n_af_patients[1]
  consort$ckd_patients  <- consort_raw$n_ckd_patients[1]
  consort$eligible      <- consort_raw$n_eligible[1]
  try(dbExecute(con, "DROP TABLE IF EXISTS #consort_counts"), silent = TRUE)

  message(sprintf("    -> New DOAC users:    %d", consort$first_doac))
  message(sprintf("    -> With AF:           %d", consort$af_patients))
  message(sprintf("    -> With CKD 3b-5:     %d", consort$ckd_patients))
  message(sprintf("    -> Eligible:          %d", consort$eligible))

  message("  Step 2: Assigning treatment (apixaban vs rivaroxaban)...")
  dbExecute(con, sql_parts$treatment)
  consort$treatment <- count_temp(con, "#treatment")
  message(sprintf("    -> #treatment: %d patients", consort$treatment))

  message("  Step 3: Ascertaining outcomes...")
  dbExecute(con, sql_parts$outcome)
  consort$outcomes <- count_temp(con, "#outcomes")
  message(sprintf("    -> #outcomes: %d patients", consort$outcomes))

  message("  Step 4: Extracting confounders and building analytic cohort...")
  dbExecute(con, sql_parts$confounders)
  cohort <- dbGetQuery(con, "SELECT * FROM #analytic_cohort")

  names(cohort) <- tolower(names(cohort))

  consort$analytic    <- nrow(cohort)
  consort$n_apixaban  <- sum(cohort$treatment == 1, na.rm = TRUE)
  consort$n_rivaroxaban <- sum(cohort$treatment == 0, na.rm = TRUE)

  message(sprintf(
    "  Analytic cohort: %d patients (%d apixaban, %d rivaroxaban)",
    consort$analytic, consort$n_apixaban, consort$n_rivaroxaban
  ))

  for (tbl in c("#eligible", "#treatment", "#outcomes", "#analytic_cohort")) {
    try(dbExecute(con, paste("DROP TABLE IF EXISTS", tbl)), silent = TRUE)
  }

  attr(cohort, "consort") <- consort
  return(cohort)
}
```


## 3a. CONSORT Flow Diagram

```{r}
#| label: consort-text

print_consort_table <- function(consort) {
  steps <- tibble::tibble(
    Step = c("1a. New DOAC users (apixaban/rivaroxaban, 365-day washout)",
             "1b. With AF diagnosis (I48.x, non-valvular)",
             "1c. With CKD 3b-5 (Dx or eGFR < 45)",
             "1d. Eligible (age >= 18, active in CDW)",
             "2.  Treatment assigned (apixaban vs rivaroxaban)",
             "3.  Outcomes ascertained",
             "4.  Analytic cohort (with confounders)"),
    N    = c(consort$first_doac,
             consort$af_patients,
             consort$ckd_patients,
             consort$eligible,
             consort$treatment,
             consort$outcomes,
             consort$analytic),
    Excluded = c(NA_integer_,
                 consort$first_doac   - consort$af_patients,
                 consort$af_patients  - consort$ckd_patients,
                 consort$ckd_patients - consort$eligible,
                 consort$eligible     - consort$treatment,
                 consort$treatment    - consort$outcomes,
                 consort$outcomes     - consort$analytic)
  )
  message("\n=== CONSORT Flow (text) ===")
  print(steps, n = Inf)
  invisible(steps)
}
```

```{r}
#| label: consort-diagram

render_consort_diagram <- function(consort) {

  grid::grid.newpage()
  fmt <- function(x) format(x, big.mark = ",")

  box_w  <- grid::unit(0.50, "npc")
  box_h  <- grid::unit(0.055, "npc")
  excl_w <- grid::unit(0.28, "npc")
  excl_h <- grid::unit(0.040, "npc")
  x_main <- 0.38
  x_excl <- 0.82

  y_pos <- seq(0.92, 0.12, length.out = 8)

  draw_box <- function(x, y, w, h, label, fill = "white") {
    grid::grid.rect(x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    width = w, height = h,
                    gp = grid::gpar(fill = fill, col = "grey30", lwd = 1.5))
    grid::grid.text(label, x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    gp = grid::gpar(fontsize = 9, fontface = "bold"))
  }

  draw_arrow <- function(x1, y1, x2, y2) {
    grid::grid.lines(x = grid::unit(c(x1, x2), "npc"),
                     y = grid::unit(c(y1, y2), "npc"),
                     arrow = grid::arrow(length = grid::unit(0.015, "npc"),
                                         type = "closed"),
                     gp = grid::gpar(fill = "grey30", col = "grey30"))
  }

  draw_excl <- function(y_from, y_to, n, reason) {
    if (n <= 0) return(invisible(NULL))
    mid_y <- mean(c(y_from - 0.028, y_to + 0.028))
    draw_arrow(x_main + 0.25, mid_y, x_excl - 0.14, mid_y)
    draw_box(x_excl, mid_y, excl_w, excl_h,
             sprintf("Excluded: %s\n(%s)", fmt(n), reason),
             fill = "#FFF3F3")
  }

  grid::grid.text("CONSORT Flow Diagram — Apixaban vs Rivaroxaban in AF+CKD",
                  x = 0.5, y = 0.97,
                  gp = grid::gpar(fontsize = 13, fontface = "bold"))

  labels <- c(
    sprintf("1a. New DOAC Users\nn = %s", fmt(consort$first_doac)),
    sprintf("1b. With AF (I48.x, non-valvular)\nn = %s", fmt(consort$af_patients)),
    sprintf("1c. With CKD 3b-5\nn = %s", fmt(consort$ckd_patients)),
    sprintf("1d. Eligible\nn = %s", fmt(consort$eligible)),
    sprintf("2. Treatment Assigned\nn = %s", fmt(consort$treatment)),
    sprintf("3. Outcomes Ascertained\nn = %s", fmt(consort$outcomes)),
    sprintf("4. Analytic Cohort\nn = %s", fmt(consort$analytic))
  )
  fills <- c(rep("white", 6), "#F0FFF0")

  for (i in seq_along(labels)) {
    draw_box(x_main, y_pos[i], box_w, box_h, labels[i], fill = fills[i])
    if (i < length(labels)) {
      draw_arrow(x_main, y_pos[i] - 0.028, x_main, y_pos[i + 1] + 0.028)
    }
  }

  excl_counts  <- c(consort$first_doac   - consort$af_patients,
                    consort$af_patients  - consort$ckd_patients,
                    consort$ckd_patients - consort$eligible,
                    consort$eligible     - consort$treatment,
                    consort$treatment    - consort$outcomes,
                    consort$outcomes     - consort$analytic)
  excl_reasons <- c("no AF diagnosis",
                    "no CKD 3b-5 / ESRD excluded",
                    "age < 18 / no active encounter",
                    "ambiguous or missing assignment",
                    "outcome ascertainment",
                    "confounder join")
  for (i in seq_along(excl_counts)) {
    draw_excl(y_pos[i], y_pos[i + 1], excl_counts[i], excl_reasons[i])
  }

  draw_arrow(x_main - 0.10, y_pos[7] - 0.028, x_main - 0.18, y_pos[8] + 0.020)
  draw_arrow(x_main + 0.10, y_pos[7] - 0.028, x_main + 0.18, y_pos[8] + 0.020)

  draw_box(x_main - 0.20, y_pos[8], grid::unit(0.26, "npc"), box_h,
           sprintf("Apixaban\nn = %s", fmt(consort$n_apixaban)),
           fill = "#F0F0FF")
  draw_box(x_main + 0.20, y_pos[8], grid::unit(0.26, "npc"), box_h,
           sprintf("Rivaroxaban\nn = %s", fmt(consort$n_rivaroxaban)),
           fill = "#F0F0FF")

  invisible(NULL)
}
```


## 4. Data Preparation

```{r}
#| label: prepare-cohort

prepare_cohort <- function(cohort) {
  cohort <- cohort |>
    mutate(
      sex_cat = factor(sex, levels = c("F", "M"),
                       labels = c("Female", "Male")),
      race_cat = case_when(
        race == "03" ~ "Black",
        race == "05" ~ "White",
        race == "02" ~ "Asian",
        race %in% c("01", "04", "06", "07") ~ "Other",
        TRUE ~ "Unknown"
      ) |> factor(),
      hispanic_cat = factor(
        if_else(hispanic == "Y", "Hispanic", "Non-Hispanic")
      ),
      ckd_stage_cat = factor(ckd_stage_cat,
                             levels = c("CKD3b", "CKD4", "CKD5")),
      dose_cat = case_when(
        dose_group %in% c("apixaban_5mg", "rivaroxaban_20mg") ~ "Standard",
        dose_group %in% c("apixaban_2.5mg", "rivaroxaban_15mg",
                          "rivaroxaban_10mg", "rivaroxaban_2.5mg") ~ "Reduced",
        TRUE ~ "Unknown"
      ) |> factor(),
      enr_basis_cat = factor(
        case_when(
          enr_basis %in% c("I", "A") ~ enr_basis,
          TRUE ~ "Other/Unknown"
        )
      ),
      treatment = as.integer(treatment),
      treatment_label = factor(
        if_else(treatment == 1, "Apixaban", "Rivaroxaban"),
        levels = c("Rivaroxaban", "Apixaban")
      ),
      # Ensure minimum follow-up time of 1 day
      time_to_composite = pmax(time_to_composite, 1),
      time_to_stroke    = pmax(time_to_stroke, 1),
      time_to_bleed     = pmax(time_to_bleed, 1),
      time_to_gi_bleed  = pmax(time_to_gi_bleed, 1),
      time_to_death     = pmax(time_to_death, 1)
    ) |>
    filter(!is.na(treatment))

  # Impute missing continuous covariates with median
  cohort <- cohort |>
    mutate(
      bmi         = if_else(is.na(bmi), median(bmi, na.rm = TRUE), bmi),
      bp_systolic = if_else(is.na(bp_systolic),
                            median(bp_systolic, na.rm = TRUE), bp_systolic),
      bp_diastolic = if_else(is.na(bp_diastolic),
                             median(bp_diastolic, na.rm = TRUE), bp_diastolic),
      egfr_value  = if_else(is.na(egfr_value),
                            median(egfr_value, na.rm = TRUE), egfr_value),
      creatinine  = if_else(is.na(creatinine),
                            median(creatinine, na.rm = TRUE), creatinine),
      hba1c       = if_else(is.na(hba1c),
                            median(hba1c, na.rm = TRUE), hba1c),
      hemoglobin  = if_else(is.na(hemoglobin),
                            median(hemoglobin, na.rm = TRUE), hemoglobin),
      platelets   = if_else(is.na(platelets),
                            median(platelets, na.rm = TRUE), platelets)
    )

  return(cohort)
}
```


## 5. Causal Analysis Functions (IPW)

```{r}
#| label: confounders

confounders <- c(
  "age_at_index", "sex_cat", "race_cat", "hispanic_cat",
  "bmi", "bp_systolic", "bp_diastolic",
  "egfr_value", "creatinine", "hba1c", "hemoglobin", "platelets",
  "chf", "hypertension", "diabetes",
  "prior_stroke", "prior_embolism", "vascular_disease",
  "prior_bleed", "liver_disease", "anemia", "obesity_dx", "cancer",
  "ckd_stage_cat",
  "antiplatelet_use", "statin_use", "acei_arb_use",
  "beta_blocker_use", "ppi_use", "nsaid_use",
  "enr_basis_cat",
  "index_year"
)

build_ps_formula <- function(confounders, cohort) {
  dropped <- character()
  keep    <- character()

  for (v in confounders) {
    col <- cohort[[v]]
    if (is.null(col)) {
      dropped <- c(dropped, v)
      next
    }
    if (is.factor(col) || is.character(col)) {
      n_levels <- length(unique(na.omit(col)))
      if (n_levels < 2) {
        dropped <- c(dropped, v)
        next
      }
    } else {
      if (sd(col, na.rm = TRUE) == 0 || all(is.na(col))) {
        dropped <- c(dropped, v)
        next
      }
    }
    keep <- c(keep, v)
  }

  if (length(dropped) > 0) {
    message(sprintf("  PS model: dropped %d constant/single-level variable(s): %s",
                    length(dropped), paste(dropped, collapse = ", ")))
  }
  message(sprintf("  PS model: using %d confounder(s)", length(keep)))

  as.formula(paste("treatment ~", paste(keep, collapse = " + ")))
}
```

```{r}
#| label: ipw-analysis

run_ipw_analysis <- function(cohort, confounders, config) {
  message("\n=== IPW Analysis ===")

  # --- Guard: treatment must have at least 2 arms ---
  n_arms <- length(unique(cohort$treatment))
  if (n_arms < 2) {
    stop(sprintf(
      "Cannot run IPW: treatment variable has only %d unique value(s) (%s). ",
      n_arms, paste(unique(cohort$treatment), collapse = ", ")),
      "Check the CONSORT diagram — one treatment arm may have been ",
      "completely filtered out during eligibility.", call. = FALSE)
  }
  message(sprintf("Treatment arms: Apixaban (n=%d), Rivaroxaban (n=%d)",
                  sum(cohort$treatment == 1), sum(cohort$treatment == 0)))

  # --- Build PS formula, dropping single-level factors ---
  ps_formula <- build_ps_formula(confounders, cohort)

  # --- Fit propensity score model and compute weights ---
  message("Fitting propensity score model (logistic regression)...")
  weights <- weightit(
    ps_formula,
    data     = cohort,
    method   = "glm",
    estimand = config$estimand
  )

  # --- Balance diagnostics ---
  message("Checking covariate balance...")
  bal <- bal.tab(weights, stats = c("m", "v"), thresholds = c(m = 0.1))
  print(bal)

  p_love <- love.plot(weights, threshold = 0.1, abs = TRUE,
                      title = "Covariate Balance: Apixaban vs Rivaroxaban (IPW)")

  cohort$ps <- weights$ps
  p_ps <- ggplot(cohort, aes(x = ps, fill = treatment_label)) +
    geom_density(alpha = 0.5) +
    labs(x = "Propensity Score (P[Apixaban])", y = "Density",
         title = "Propensity Score Distribution by Treatment Group",
         fill = "Treatment") +
    theme_minimal()

  ess_fn <- function(w) sum(w)^2 / sum(w^2)
  message(sprintf("Effective sample size: Apixaban=%.1f, Rivaroxaban=%.1f",
                  ess_fn(weights$weights[cohort$treatment == 1]),
                  ess_fn(weights$weights[cohort$treatment == 0])))

  cohort$ipw <- weights$weights

  # --- Primary: Composite (stroke/SE + major bleeding) ---
  message("\n--- Primary Outcome: Composite (Stroke/SE + Major Bleeding) ---")
  fit_composite <- coxph(
    Surv(time_to_composite, composite_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_composite))

  # --- Secondary: Stroke / SE ---
  message("\n--- Secondary Outcome: Stroke / Systemic Embolism ---")
  fit_stroke <- coxph(
    Surv(time_to_stroke, stroke_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_stroke))

  # --- Secondary: Major bleeding ---
  message("\n--- Secondary Outcome: Major Bleeding ---")
  fit_bleed <- coxph(
    Surv(time_to_bleed, bleed_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_bleed))

  # --- Secondary: GI bleeding ---
  message("\n--- Secondary Outcome: GI Bleeding ---")
  fit_gi_bleed <- coxph(
    Surv(time_to_gi_bleed, gi_bleed_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_gi_bleed))

  # --- Secondary: All-cause mortality ---
  message("\n--- Secondary Outcome: All-Cause Mortality ---")
  fit_death <- coxph(
    Surv(time_to_death, death_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_death))

  # --- Weighted KM curves ---
  km_composite <- survfit(
    Surv(time_to_composite, composite_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_composite <- ggsurvplot(
    km_composite, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "Composite (Stroke/SE + Bleeding): Apixaban vs Rivaroxaban (IPW)",
    xlab = "Days from DOAC Initiation",
    ylab = "Event-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  km_stroke <- survfit(
    Surv(time_to_stroke, stroke_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_stroke <- ggsurvplot(
    km_stroke, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "Stroke/SE: Apixaban vs Rivaroxaban (IPW)",
    xlab = "Days from DOAC Initiation",
    ylab = "Stroke-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  km_bleed <- survfit(
    Surv(time_to_bleed, bleed_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_bleed <- ggsurvplot(
    km_bleed, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "Major Bleeding: Apixaban vs Rivaroxaban (IPW)",
    xlab = "Days from DOAC Initiation",
    ylab = "Bleeding-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  return(list(
    weights = weights,
    balance = bal,
    fits = list(
      composite = fit_composite,
      stroke    = fit_stroke,
      bleed     = fit_bleed,
      gi_bleed  = fit_gi_bleed,
      death     = fit_death
    ),
    plots = list(
      love         = p_love,
      ps_dist      = p_ps,
      km_composite = p_km_composite,
      km_stroke    = p_km_stroke,
      km_bleed     = p_km_bleed
    ),
    cohort = cohort
  ))
}
```


## 6. Subgroup Analyses

```{r}
#| label: subgroup-analysis

run_subgroup_analyses <- function(cohort, confounders, config) {
  message("\n=== Subgroup Analyses ===")

  subgroups <- list(
    "CKD Stage 3b"      = cohort$ckd_stage_cat == "CKD3b",
    "CKD Stage 4-5"     = cohort$ckd_stage_cat %in% c("CKD4", "CKD5"),
    "Age < 75"          = cohort$age_at_index < 75,
    "Age >= 75"         = cohort$age_at_index >= 75,
    "Female"            = cohort$sex_cat == "Female",
    "Male"              = cohort$sex_cat == "Male",
    "Standard dose"     = cohort$dose_cat == "Standard",
    "Reduced dose"      = cohort$dose_cat == "Reduced",
    "Prior bleeding"    = cohort$prior_bleed == 1,
    "No prior bleeding" = cohort$prior_bleed == 0,
    "Diabetes"          = cohort$diabetes == 1,
    "No diabetes"       = cohort$diabetes == 0
  )

  results <- tibble::tibble(
    Subgroup = character(),
    N = integer(),
    N_apixaban = integer(),
    N_rivaroxaban = integer(),
    N_events = integer(),
    HR = numeric(),
    CI_lo = numeric(),
    CI_hi = numeric(),
    P_value = numeric()
  )

  for (sg_name in names(subgroups)) {
    sg_data <- cohort[subgroups[[sg_name]], ]

    # Guard: need 2 arms and at least 10 patients per arm
    if (length(unique(sg_data$treatment)) < 2 ||
        sum(sg_data$treatment == 1) < 10 ||
        sum(sg_data$treatment == 0) < 10) {
      message(sprintf("  Skipping '%s': insufficient patients in one arm", sg_name))
      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_apixaban = sum(sg_data$treatment == 1),
        N_rivaroxaban = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
      next
    }

    # Re-fit PS within subgroup
    tryCatch({
      ps_formula <- build_ps_formula(confounders, sg_data)
      w <- weightit(ps_formula, data = sg_data, method = "glm",
                    estimand = config$estimand)
      sg_data$ipw <- w$weights

      fit <- coxph(Surv(time_to_composite, composite_event) ~ treatment,
                   data = sg_data, weights = ipw, robust = TRUE)
      s <- summary(fit)

      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name,
        N = nrow(sg_data),
        N_apixaban = sum(sg_data$treatment == 1),
        N_rivaroxaban = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = s$conf.int[1, 1],
        CI_lo = s$conf.int[1, 3],
        CI_hi = s$conf.int[1, 4],
        P_value = s$coefficients[1, "Pr(>|z|)"]
      ))
      message(sprintf("  %s: HR=%.2f (%.2f-%.2f), N=%d",
                      sg_name, s$conf.int[1,1], s$conf.int[1,3],
                      s$conf.int[1,4], nrow(sg_data)))
    }, error = function(e) {
      message(sprintf("  Error in '%s': %s", sg_name, e$message))
      results <<- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_apixaban = sum(sg_data$treatment == 1),
        N_rivaroxaban = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
    })
  }

  return(results)
}
```


## 7. Sensitivity Analyses

```{r}
#| label: sensitivity-analysis

run_sensitivity <- function(ipw_results, cohort, confounders, config) {
  message("\n=== Sensitivity Analyses ===")

  sensitivity <- list()

  # --- 7a. E-value for composite outcome ---
  message("\n--- E-value Analysis (Composite Outcome) ---")
  fit_composite <- ipw_results$fits$composite
  s <- summary(fit_composite)
  hr  <- s$conf.int[1, 1]
  lo  <- s$conf.int[1, 3]
  hi  <- s$conf.int[1, 4]

  evals <- evalues.HR(hr, lo = lo, hi = hi, rare = TRUE)
  sensitivity$evalue_composite <- evals
  message("E-value for composite outcome:")
  print(evals)

  # --- 7b. E-value for major bleeding ---
  message("\n--- E-value Analysis (Major Bleeding) ---")
  fit_bleed <- ipw_results$fits$bleed
  s_bleed <- summary(fit_bleed)
  hr_b  <- s_bleed$conf.int[1, 1]
  lo_b  <- s_bleed$conf.int[1, 3]
  hi_b  <- s_bleed$conf.int[1, 4]

  evals_bleed <- evalues.HR(hr_b, lo = lo_b, hi = hi_b, rare = TRUE)
  sensitivity$evalue_bleed <- evals_bleed
  message("E-value for major bleeding:")
  print(evals_bleed)

  # --- 7c. PS trimming (5th-95th percentile) ---
  message("\n--- PS Trimming (5th-95th Percentile) ---")
  ps_lo <- quantile(cohort$ps, 0.05, na.rm = TRUE)
  ps_hi <- quantile(cohort$ps, 0.95, na.rm = TRUE)
  cohort_trimmed <- cohort |>
    filter(ps >= ps_lo & ps <= ps_hi)

  message(sprintf("  Trimmed from %d to %d patients (PS range: %.3f-%.3f)",
                  nrow(cohort), nrow(cohort_trimmed), ps_lo, ps_hi))

  if (length(unique(cohort_trimmed$treatment)) < 2) {
    message("WARNING: After PS trimming, only 1 arm remains. Skipping.")
    sensitivity$trimmed_fit <- NULL
  } else {
    ps_formula_trim <- build_ps_formula(confounders, cohort_trimmed)
    w_trim <- weightit(ps_formula_trim, data = cohort_trimmed,
                       method = "glm", estimand = config$estimand)
    cohort_trimmed$ipw <- w_trim$weights

    fit_trim <- coxph(
      Surv(time_to_composite, composite_event) ~ treatment,
      data = cohort_trimmed, weights = ipw, robust = TRUE
    )
    sensitivity$trimmed_fit <- fit_trim
    message("Trimmed analysis (composite):")
    print(summary(fit_trim))
  }

  # --- 7d. Post-Epic era only (2021+) ---
  message("\n--- Post-Epic Era (2021+) ---")
  cohort_epic <- cohort |> filter(index_year >= 2021)

  if (length(unique(cohort_epic$treatment)) < 2 ||
      min(table(cohort_epic$treatment)) < 10) {
    message("WARNING: Insufficient patients in post-Epic era. Skipping.")
    sensitivity$epic_fit <- NULL
  } else {
    ps_formula_epic <- build_ps_formula(confounders, cohort_epic)
    w_epic <- weightit(ps_formula_epic, data = cohort_epic,
                       method = "glm", estimand = config$estimand)
    cohort_epic$ipw <- w_epic$weights

    fit_epic <- coxph(
      Surv(time_to_composite, composite_event) ~ treatment,
      data = cohort_epic, weights = ipw, robust = TRUE
    )
    sensitivity$epic_fit <- fit_epic
    message(sprintf("Post-Epic analysis (N=%d):", nrow(cohort_epic)))
    print(summary(fit_epic))
  }

  return(sensitivity)
}
```


## 8. Summary Helper

```{r}
#| label: summary-helper

format_hr <- function(fit) {
  s <- summary(fit)
  sprintf("HR = %.2f (95%% CI: %.2f-%.2f), p = %.4f",
          s$conf.int[1, 1], s$conf.int[1, 3], s$conf.int[1, 4],
          s$coefficients[1, "Pr(>|z|)"])
}
```


<!-- ============================================================= -->
<!-- PART 2: Execution & Results (Sections 9+)                      -->
<!-- Each section calls its function and displays results inline    -->
<!-- ============================================================= -->


## 9. Connect and Pull Cohort

```{r}
#| label: exec-connect

con <- connect_cdw()
cohort_raw <- pull_analytic_cohort(con, config)
consort <- attr(cohort_raw, "consort")
dbDisconnect(con)
```

### CONSORT Flow (Text)

```{r}
#| label: exec-consort-text

consort_table <- print_consort_table(consort)
knitr::kable(consort_table, caption = "CONSORT Patient Flow")
```

### CONSORT Flow Diagram

```{r}
#| label: fig-consort
#| fig-cap: "CONSORT flow diagram showing patient attrition at each cohort-building step."
#| fig-width: 10
#| fig-height: 12

render_consort_diagram(consort)
```

### Empty Cohort Guard

```{r}
#| label: exec-empty-guard

if (nrow(cohort_raw) == 0) {
  message("*** STOPPING: Analytic cohort has 0 patients. ***")
  message("Review the CONSORT diagram above to identify where patients were lost.")
  knitr::knit_exit()
}
```


## 10. Prepare Data and Table 1

```{r}
#| label: exec-prepare

cohort <- prepare_cohort(cohort_raw)

message(sprintf("Prepared cohort: %d patients (%d apixaban, %d rivaroxaban)",
                nrow(cohort),
                sum(cohort$treatment == 1),
                sum(cohort$treatment == 0)))
```

### Table 1: Baseline Characteristics

```{r}
#| label: exec-table1

table1_data <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = as.character(n()),
    `Age (mean +/- SD)` = sprintf("%.1f +/- %.1f", mean(age_at_index), sd(age_at_index)),
    `Female (%)` = sprintf("%.1f", 100 * mean(sex_cat == "Female")),
    `White (%)` = sprintf("%.1f", 100 * mean(race_cat == "White")),
    `Black (%)` = sprintf("%.1f", 100 * mean(race_cat == "Black")),
    `Hispanic (%)` = sprintf("%.1f", 100 * mean(hispanic_cat == "Hispanic")),
    `eGFR (mean)` = sprintf("%.1f", mean(egfr_value, na.rm = TRUE)),
    `Creatinine (mean)` = sprintf("%.1f", mean(creatinine, na.rm = TRUE)),
    `BMI (mean)` = sprintf("%.1f", mean(bmi, na.rm = TRUE)),
    `SBP (mean)` = sprintf("%.0f", mean(bp_systolic, na.rm = TRUE)),
    `HbA1c (mean)` = sprintf("%.1f", mean(hba1c, na.rm = TRUE)),
    `Hemoglobin (mean)` = sprintf("%.1f", mean(hemoglobin, na.rm = TRUE)),
    `CKD 3b (%)` = sprintf("%.1f", 100 * mean(ckd_stage_cat == "CKD3b")),
    `CKD 4 (%)` = sprintf("%.1f", 100 * mean(ckd_stage_cat == "CKD4")),
    `CKD 5 (%)` = sprintf("%.1f", 100 * mean(ckd_stage_cat == "CKD5")),
    `CHF (%)` = sprintf("%.1f", 100 * mean(chf)),
    `Hypertension (%)` = sprintf("%.1f", 100 * mean(hypertension)),
    `Diabetes (%)` = sprintf("%.1f", 100 * mean(diabetes)),
    `Prior stroke/TIA (%)` = sprintf("%.1f", 100 * mean(prior_stroke)),
    `Prior bleeding (%)` = sprintf("%.1f", 100 * mean(prior_bleed)),
    `Liver disease (%)` = sprintf("%.1f", 100 * mean(liver_disease)),
    `Antiplatelet (%)` = sprintf("%.1f", 100 * mean(antiplatelet_use)),
    `Statin (%)` = sprintf("%.1f", 100 * mean(statin_use)),
    `ACEi/ARB (%)` = sprintf("%.1f", 100 * mean(acei_arb_use)),
    `Beta-blocker (%)` = sprintf("%.1f", 100 * mean(beta_blocker_use)),
    `PPI (%)` = sprintf("%.1f", 100 * mean(ppi_use)),
    `NSAID (%)` = sprintf("%.1f", 100 * mean(nsaid_use)),
    `Reduced dose (%)` = sprintf("%.1f", 100 * mean(dose_cat == "Reduced")),
    .groups = "drop"
  ) |>
  pivot_longer(-treatment_label, names_to = "Characteristic", values_to = "Value") |>
  pivot_wider(names_from = treatment_label, values_from = Value)

knitr::kable(table1_data, caption = "Table 1: Baseline Characteristics by Treatment Group")
```

### Event Rates

```{r}
#| label: exec-events

events <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = n(),
    `Composite events` = sum(composite_event),
    `Stroke/SE events` = sum(stroke_event),
    `Major bleeding events` = sum(bleed_event),
    `GI bleeding events` = sum(gi_bleed_event),
    `Deaths` = sum(death_event),
    .groups = "drop"
  )
knitr::kable(events, caption = "Outcome Event Counts by Treatment Group")
```


## 11. IPW Analysis

```{r}
#| label: exec-ipw

ipw_results <- run_ipw_analysis(cohort, confounders, config)
cohort <- ipw_results$cohort
```

### Love Plot (Covariate Balance)

```{r}
#| label: fig-love-plot
#| fig-cap: "Absolute standardized mean differences before and after IPW."
#| fig-width: 10
#| fig-height: 8

print(ipw_results$plots$love)
```

### Propensity Score Distribution

```{r}
#| label: fig-ps-dist
#| fig-cap: "Propensity score distributions by treatment group."
#| fig-width: 8
#| fig-height: 5

print(ipw_results$plots$ps_dist)
```


## 12. Cox Model Results and Kaplan-Meier Curves

### Summary of All Outcomes

```{r}
#| label: exec-results-summary

outcome_summary <- tibble::tibble(
  Outcome = c("Composite (stroke/SE + bleeding)",
              "Stroke / systemic embolism",
              "Major bleeding",
              "GI bleeding",
              "All-cause mortality"),
  Result = c(
    format_hr(ipw_results$fits$composite),
    format_hr(ipw_results$fits$stroke),
    format_hr(ipw_results$fits$bleed),
    format_hr(ipw_results$fits$gi_bleed),
    format_hr(ipw_results$fits$death)
  )
)
knitr::kable(outcome_summary,
             caption = "Weighted Cox PH Results: Apixaban (treatment=1) vs Rivaroxaban (ref)")
```

### Kaplan-Meier: Composite Outcome

```{r}
#| label: fig-km-composite
#| fig-cap: "Weighted Kaplan-Meier curves for the composite outcome (stroke/SE + major bleeding)."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_composite)
```

### Kaplan-Meier: Stroke / SE

```{r}
#| label: fig-km-stroke
#| fig-cap: "Weighted Kaplan-Meier curves for stroke/systemic embolism."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_stroke)
```

### Kaplan-Meier: Major Bleeding

```{r}
#| label: fig-km-bleed
#| fig-cap: "Weighted Kaplan-Meier curves for major bleeding."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_bleed)
```


## 13. Subgroup Analyses

```{r}
#| label: exec-subgroups

subgroup_results <- run_subgroup_analyses(cohort, confounders, config)
knitr::kable(subgroup_results |>
               mutate(across(where(is.numeric), ~round(., 3))),
             caption = "Subgroup Analysis: Composite Outcome (Apixaban vs Rivaroxaban)")
```

### Forest Plot

```{r}
#| label: fig-forest
#| fig-cap: "Forest plot of subgroup analyses for the composite outcome."
#| fig-width: 10
#| fig-height: 8

sg_plot_data <- subgroup_results |> filter(!is.na(HR))

if (nrow(sg_plot_data) > 0) {
  ggplot(sg_plot_data, aes(x = HR, y = reorder(Subgroup, HR))) +
    geom_point(size = 3) +
    geom_errorbarh(aes(xmin = CI_lo, xmax = CI_hi), height = 0.2) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
    scale_x_log10() +
    labs(x = "Hazard Ratio (Apixaban vs Rivaroxaban)",
         y = NULL,
         title = "Subgroup Analyses: Composite Outcome",
         subtitle = "HR < 1 favors apixaban") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 10))
} else {
  message("No subgroup analyses had sufficient data for plotting.")
}
```


## 14. Sensitivity Analyses

```{r}
#| label: exec-sensitivity

sensitivity_results <- run_sensitivity(ipw_results, cohort, confounders, config)
```

### E-value Results

```{r}
#| label: exec-evalue

message("=== E-value: Composite Outcome ===")
print(sensitivity_results$evalue_composite)

message("\n=== E-value: Major Bleeding ===")
print(sensitivity_results$evalue_bleed)
```

### PS Trimming Results

```{r}
#| label: exec-trimmed

if (!is.null(sensitivity_results$trimmed_fit)) {
  message("Trimmed composite result:")
  message(format_hr(sensitivity_results$trimmed_fit))
} else {
  message("PS trimming analysis was skipped (insufficient arms).")
}
```

### Post-Epic Era Results

```{r}
#| label: exec-epic

if (!is.null(sensitivity_results$epic_fit)) {
  message("Post-Epic (2021+) composite result:")
  message(format_hr(sensitivity_results$epic_fit))
} else {
  message("Post-Epic analysis was skipped (insufficient patients).")
}
```


## 15. Summary

```{r}
#| label: exec-summary

message("\n", strrep("=", 70))
message("PROTOCOL 01 SUMMARY: Apixaban vs Rivaroxaban in AF + CKD 3b-5")
message(strrep("=", 70))
message(sprintf("Analytic cohort: %d patients", nrow(cohort)))
message(sprintf("  Apixaban:    %d", sum(cohort$treatment == 1)))
message(sprintf("  Rivaroxaban: %d", sum(cohort$treatment == 0)))
message(sprintf("Follow-up:     %d days", config$followup_days))
message(sprintf("Study period:  %s to %s", config$study_start, config$study_end))
message("")
message("PRIMARY OUTCOME (Composite: Stroke/SE + Major Bleeding):")
message(sprintf("  %s", format_hr(ipw_results$fits$composite)))
message("")
message("SECONDARY OUTCOMES:")
message(sprintf("  Stroke/SE:       %s", format_hr(ipw_results$fits$stroke)))
message(sprintf("  Major bleeding:  %s", format_hr(ipw_results$fits$bleed)))
message(sprintf("  GI bleeding:     %s", format_hr(ipw_results$fits$gi_bleed)))
message(sprintf("  All-cause death: %s", format_hr(ipw_results$fits$death)))
message("")
message("Interpretation: HR < 1 favors apixaban; HR > 1 favors rivaroxaban.")
message("Treatment = 1 (apixaban) is the 'treated' group in the Cox model.")
message(strrep("=", 70))
```
