---
title: "Protocol 02: Early Rhythm Control vs Rate Control in Elderly AF (Age >= 80)"
subtitle: "Target Trial Emulation --- PCORnet CDW"
author: "Auto-Protocol Designer"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    self-contained: true
    theme: cosmo
execute:
  eval: true
  warning: false
  message: true
---

## Study Overview

This protocol emulates a target trial comparing **early rhythm control vs rate
control** for a composite cardiovascular outcome in adults aged >= 80 with newly
diagnosed atrial fibrillation (AF).

- **Design:** Landmark cohort with 12-month treatment classification window
- **Estimand:** ATE (Average Treatment Effect)
- **Method:** Inverse Probability Weighting (IPW) + Weighted Cox PH
- **Time zero (enrollment):** First AF diagnosis date (first I48.x at age >= 80)
- **Landmark:** 12 months after AF diagnosis (day 366)
- **Follow-up:** 1,095 days (3 years) from the landmark
- **Study period:** 2016-01-01 to 2024-12-31


<!-- ============================================================= -->
<!-- PART 1: Function definitions (Sections 0-8)                    -->
<!-- These chunks register functions but produce no visible output  -->
<!-- ============================================================= -->


## 0. Setup and Configuration

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(DBI)
library(odbc)
library(WeightIt)
library(cobalt)
library(survival)
library(sandwich)
library(lmtest)
library(EValue)
library(survminer)
```

```{r}
#| label: config

config <- list(
  question        = "Early Rhythm Control vs Rate Control in Elderly AF (Age >= 80)",
  exposure_var    = "treatment",
  outcome_var     = "composite_event",
  time_var        = "time_to_composite",
  estimand        = "ATE",
  analysis_method = "ipw",
  study_start     = "2016-01-01",
  study_end       = "2024-12-31",
  classification_days = 365,   # 12-month treatment classification window
  followup_days   = 1095,      # 3-year follow-up from landmark
  washout_af_days = 365,       # no prior AF in 365 days = incident AF
  lookback_dx     = 365,       # days for baseline comorbidity lookback
  lookback_lab    = 180,       # days for baseline lab lookback
  lookback_vital  = 365,       # days for baseline vital lookback
  lookback_rx     = 90         # days for concurrent medication lookback
)
```


## 1. Database Connection

```{r}
#| label: db-connect

connect_cdw <- function() {
  con <- DBI::dbConnect(odbc::odbc(), "SQLODBCD17CDM")
  return(con)
}
```


## 2. Cohort SQL

The cohort is built in sequential steps using temp tables:
`#first_af` -> `#eligible` -> `#treatment` -> `#landmark` -> `#outcomes` ->
`#analytic_cohort`.

All tables are fully qualified as `CDW.dbo.TABLE_NAME`. Medications are
identified by RXNORM_CUI (SCD + SBD), diagnoses by ICD-10 with `DX_TYPE = '10'`.
Legacy encounters are excluded throughout. A 12-month landmark design is used:
patients must survive 12 months post-AF diagnosis to be included, and outcomes
are assessed from the landmark forward.

```{r}
#| label: cohort-sql

build_cohort_sql <- function(config) {

  # ── Step 1: Eligible population ──
  # Newly diagnosed AF at age >= 80
  eligibility_sql <- paste0("
    -- =========================================================
    -- STEP 1: Eligible Population
    -- Newly diagnosed AF (first I48.x) at age >= 80
    -- =========================================================

    -- 1a. Find first AF diagnosis per patient (incident AF)
    IF OBJECT_ID('tempdb..#first_af_raw') IS NOT NULL DROP TABLE #first_af_raw;

    SELECT
      dx.PATID,
      MIN(dx.ADMIT_DATE) AS af_dx_date
    INTO #first_af_raw
    FROM CDW.dbo.DIAGNOSIS dx
    INNER JOIN CDW.dbo.ENCOUNTER e ON dx.ENCOUNTERID = e.ENCOUNTERID
      AND e.RAW_ENC_TYPE <> 'Legacy Encounter'
    WHERE dx.DX LIKE 'I48%'
      AND dx.DX_TYPE = '10'
      AND dx.ADMIT_DATE BETWEEN '", config$study_start, "' AND '", config$study_end, "'
    GROUP BY dx.PATID
    ;

    -- 1b. Apply incident AF requirement (no prior I48.x in ", config$washout_af_days, " days)
    IF OBJECT_ID('tempdb..#first_af') IS NOT NULL DROP TABLE #first_af;

    SELECT f.PATID, f.af_dx_date AS index_date
    INTO #first_af
    FROM #first_af_raw f
    WHERE NOT EXISTS (
      SELECT 1 FROM CDW.dbo.DIAGNOSIS dx2
      WHERE dx2.PATID = f.PATID
        AND dx2.DX LIKE 'I48%'
        AND dx2.DX_TYPE = '10'
        AND dx2.ADMIT_DATE >= DATEADD(day, -", config$washout_af_days, ", f.af_dx_date)
        AND dx2.ADMIT_DATE < f.af_dx_date
        AND dx2.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
    )
    ;
    DROP TABLE IF EXISTS #first_af_raw;

    -- 1c. Apply age >= 80 and demographic filters
    IF OBJECT_ID('tempdb..#eligible') IS NOT NULL DROP TABLE #eligible;

    SELECT
      fa.PATID,
      fa.index_date,
      d.BIRTH_DATE,
      d.SEX,
      d.RACE,
      d.HISPANIC,
      DATEDIFF(year, d.BIRTH_DATE, fa.index_date) AS age_at_index
    INTO #eligible
    FROM #first_af fa
    INNER JOIN CDW.dbo.DEMOGRAPHIC d ON fa.PATID = d.PATID
    WHERE DATEDIFF(year, d.BIRTH_DATE, fa.index_date) >= 80
      -- Must have at least one non-legacy encounter in prior 365 days
      AND EXISTS (
        SELECT 1 FROM CDW.dbo.ENCOUNTER e
        WHERE e.PATID = fa.PATID
          AND e.RAW_ENC_TYPE <> 'Legacy Encounter'
          AND e.ADMIT_DATE >= DATEADD(day, -365, fa.index_date)
          AND e.ADMIT_DATE <= fa.index_date
          AND e.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      )
    ;

    -- ── CONSORT: capture sub-step counts ──
    IF OBJECT_ID('tempdb..#consort_counts') IS NOT NULL DROP TABLE #consort_counts;
    SELECT
      (SELECT COUNT(DISTINCT PATID) FROM #first_af)   AS n_incident_af,
      (SELECT COUNT(DISTINCT PATID) FROM #eligible)    AS n_eligible
    INTO #consort_counts;

    DROP TABLE IF EXISTS #first_af;
  ")

  # ── Step 2: Treatment assignment ──
  # Classify by strategy received within 12 months of AF diagnosis
  treatment_sql <- paste0("
    -- =========================================================
    -- STEP 2: Treatment Assignment
    -- Rhythm control (treatment=1) vs Rate control (treatment=0)
    -- Classification window: ", config$classification_days, " days from AF diagnosis
    -- ITT: any rhythm control intervention -> rhythm control arm
    -- =========================================================

    IF OBJECT_ID('tempdb..#treatment') IS NOT NULL DROP TABLE #treatment;

    SELECT
      e.PATID,
      e.index_date,
      e.BIRTH_DATE, e.SEX, e.RACE, e.HISPANIC, e.age_at_index,
      -- Treatment assignment: rhythm control if ANY AAD/cardioversion/ablation
      CASE
        WHEN rhythm.PATID IS NOT NULL THEN 1  -- Rhythm control
        WHEN rate.PATID IS NOT NULL   THEN 0  -- Rate control only
        ELSE NULL  -- No treatment identified
      END AS treatment,
      -- Calendar year
      YEAR(e.index_date) AS index_year
    INTO #treatment
    FROM #eligible e

    -- Rhythm control: AAD prescription OR cardioversion OR ablation
    LEFT JOIN (
      SELECT DISTINCT sub.PATID FROM (
        -- AAD prescriptions within classification window
        SELECT p.PATID
        FROM CDW.dbo.PRESCRIBING p
        INNER JOIN #eligible e2 ON p.PATID = e2.PATID
        WHERE p.RXNORM_CUI IN (
          -- Amiodarone oral (SCD + SBD)
          '835956','833528','835960','834348','833530','835958','834346','834350',
          -- Flecainide
          '886662','886666','886671',
          -- Sotalol (all oral formulations including AF-labeled)
          '904634','1923426','1923422','1923424','904632','904589',
          '1922765','1922720','1922763',
          '904605','904571','904583','904593',
          '1923427','1923423','1923425','904591',
          '1922766','1922721','1922764',
          -- Dronedarone
          '854856','854859',
          -- Dofetilide
          '310003','310004','310005','284404','284405','285016',
          -- Propafenone (IR + ER)
          '861424','861427','861430','861156','861164','861171',
          '861159','861167','861173'
        )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) > e2.index_date
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            <= DATEADD(day, ", config$classification_days, ", e2.index_date)

        UNION

        -- Cardioversion or catheter ablation within classification window
        SELECT pr.PATID
        FROM CDW.dbo.PROCEDURES pr
        INNER JOIN CDW.dbo.ENCOUNTER enc ON pr.ENCOUNTERID = enc.ENCOUNTERID
          AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
        INNER JOIN #eligible e3 ON pr.PATID = e3.PATID
        WHERE pr.PX IN ('92960','92961','93656','93657')
          AND pr.PX_TYPE = 'CH'
          AND pr.PX_DATE > e3.index_date
          AND pr.PX_DATE <= DATEADD(day, ", config$classification_days, ", e3.index_date)
      ) sub
    ) rhythm ON e.PATID = rhythm.PATID

    -- Rate control: beta-blocker, CCB, or digoxin within classification window
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #eligible e4 ON p.PATID = e4.PATID
      WHERE p.RXNORM_CUI IN (
        -- Metoprolol tartrate IR
        '866924','866514','866511','2723025','1606347','1606349',
        -- Metoprolol succinate ER
        '866427','866436','866412','866419',
        '866429','866438','866414','866421',
        -- Diltiazem IR + ER
        '833217','831103','831102','831054',
        '830861','830845','830837','830801','830795','831359',
        '830874','830877','830879','830882','830897','830900',
        -- Verapamil IR + ER
        '897722','897683','897666','901438','901446',
        '897584','897612','897618','897590','897624','897596','897630',
        '897659','897640','897649',
        -- Digoxin
        '245273','197604','197606','393245','309888','309889',
        '260350','260351','1245443','1245373'
      )
      AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) > e4.index_date
      AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
          <= DATEADD(day, ", config$classification_days, ", e4.index_date)
    ) rate ON e.PATID = rate.PATID

    -- Must have received at least one treatment
    WHERE COALESCE(rhythm.PATID, rate.PATID) IS NOT NULL
    ;

    -- Remove patients with no classifiable treatment
    DELETE FROM #treatment WHERE treatment IS NULL;
  ")

  # ── Step 3: Landmark filter ──
  # Patients must survive 12 months post-AF diagnosis
  landmark_sql <- paste0("
    -- =========================================================
    -- STEP 3: Landmark Filter
    -- Exclude patients who die within ", config$classification_days, " days
    -- of AF diagnosis (landmark survival requirement)
    -- =========================================================

    IF OBJECT_ID('tempdb..#landmark') IS NOT NULL DROP TABLE #landmark;

    SELECT
      t.*,
      DATEADD(day, ", config$classification_days, ", t.index_date) AS landmark_date
    INTO #landmark
    FROM #treatment t
    -- Exclude patients who died before the landmark
    WHERE NOT EXISTS (
      SELECT 1 FROM (
        SELECT d.PATID, d.DEATH_DATE,
               ROW_NUMBER() OVER (PARTITION BY d.PATID ORDER BY d.DEATH_DATE) AS rn
        FROM CDW.dbo.DEATH d
        WHERE d.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
      ) death
      WHERE death.PATID = t.PATID
        AND death.rn = 1
        AND death.DEATH_DATE <= DATEADD(day, ", config$classification_days, ", t.index_date)
    )
    ;
  ")

  # ── Step 4: Outcome ascertainment ──
  outcome_sql <- paste0("
    -- =========================================================
    -- STEP 4: Outcome Ascertainment (from landmark forward)
    -- Primary: Composite (CV death + stroke + HF hospitalization)
    -- Follow-up: ", config$followup_days, " days from landmark
    -- =========================================================

    -- First check DEATH_CAUSE completeness
    IF OBJECT_ID('tempdb..#death_cause_check') IS NOT NULL DROP TABLE #death_cause_check;
    SELECT
      (SELECT COUNT(DISTINCT d.PATID)
       FROM CDW.dbo.DEATH d
       INNER JOIN #landmark lm ON d.PATID = lm.PATID
       WHERE d.DEATH_DATE > lm.landmark_date
         AND d.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
         AND d.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
      ) AS n_deaths_in_followup,
      (SELECT COUNT(DISTINCT dc.PATID)
       FROM CDW.dbo.DEATH_CAUSE dc
       INNER JOIN #landmark lm2 ON dc.PATID = lm2.PATID
       INNER JOIN CDW.dbo.DEATH d2 ON dc.PATID = d2.PATID
       WHERE d2.DEATH_DATE > lm2.landmark_date
         AND d2.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm2.landmark_date)
         AND d2.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
      ) AS n_deaths_with_cause
    INTO #death_cause_check;

    IF OBJECT_ID('tempdb..#outcomes') IS NOT NULL DROP TABLE #outcomes;

    SELECT
      lm.PATID,
      lm.index_date,
      lm.landmark_date,
      lm.treatment,
      lm.index_year,
      lm.BIRTH_DATE, lm.SEX, lm.RACE, lm.HISPANIC, lm.age_at_index,

      -- Component: Stroke (I63.x ischemic + I61.x hemorrhagic) on IP/EI
      stroke.first_stroke_date,
      CASE WHEN stroke.first_stroke_date IS NOT NULL THEN 1 ELSE 0 END AS stroke_event,

      -- Component: HF hospitalization (I50.x on IP/EI)
      hf_hosp.first_hf_date,
      CASE WHEN hf_hosp.first_hf_date IS NOT NULL THEN 1 ELSE 0 END AS hf_hosp_event,

      -- Component: All-cause death (used if DEATH_CAUSE sparse; also secondary)
      death.DEATH_DATE AS death_date,
      CASE WHEN death.DEATH_DATE IS NOT NULL
           AND death.DEATH_DATE > lm.landmark_date
           AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
           THEN 1 ELSE 0 END AS death_event,

      -- Component: CV death (I-codes in DEATH_CAUSE)
      CASE WHEN cv_death.PATID IS NOT NULL
           AND death.DEATH_DATE IS NOT NULL
           AND death.DEATH_DATE > lm.landmark_date
           AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
           THEN 1 ELSE 0 END AS cv_death_event,

      -- Secondary: AF-related hospitalization (I48.x on IP/EI)
      af_hosp.first_af_hosp_date,
      CASE WHEN af_hosp.first_af_hosp_date IS NOT NULL THEN 1 ELSE 0 END AS af_hosp_event,

      -- Primary composite: first of CV death, stroke, or HF hospitalization
      -- (Will be recomputed in R based on DEATH_CAUSE completeness)
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL
          OR hf_hosp.first_hf_date IS NOT NULL
          OR (death.DEATH_DATE IS NOT NULL
              AND death.DEATH_DATE > lm.landmark_date
              AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date))
        THEN 1 ELSE 0
      END AS composite_allcause_event,

      -- Time to stroke (from landmark)
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date, stroke.first_stroke_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > lm.landmark_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_stroke,

      -- Time to HF hospitalization (from landmark)
      CASE
        WHEN hf_hosp.first_hf_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date, hf_hosp.first_hf_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > lm.landmark_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_hf_hosp,

      -- Time to death (from landmark)
      CASE
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > lm.landmark_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_death,

      -- Time to AF hospitalization (from landmark)
      CASE
        WHEN af_hosp.first_af_hosp_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date, af_hosp.first_af_hosp_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > lm.landmark_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_af_hosp,

      -- Time to composite (first of any component -- from landmark)
      -- Uses all-cause death; R will recompute with CV death if available
      CASE
        WHEN stroke.first_stroke_date IS NOT NULL
          AND hf_hosp.first_hf_date IS NOT NULL
          AND death.DEATH_DATE IS NOT NULL
            AND death.DEATH_DATE > lm.landmark_date
            AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date,
            CASE WHEN stroke.first_stroke_date <= hf_hosp.first_hf_date
                  AND stroke.first_stroke_date <= death.DEATH_DATE
                 THEN stroke.first_stroke_date
                 WHEN hf_hosp.first_hf_date <= stroke.first_stroke_date
                  AND hf_hosp.first_hf_date <= death.DEATH_DATE
                 THEN hf_hosp.first_hf_date
                 ELSE death.DEATH_DATE END)
        WHEN stroke.first_stroke_date IS NOT NULL AND hf_hosp.first_hf_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date,
            CASE WHEN stroke.first_stroke_date <= hf_hosp.first_hf_date
                 THEN stroke.first_stroke_date ELSE hf_hosp.first_hf_date END)
        WHEN stroke.first_stroke_date IS NOT NULL
          AND death.DEATH_DATE IS NOT NULL
            AND death.DEATH_DATE > lm.landmark_date
            AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date,
            CASE WHEN stroke.first_stroke_date <= death.DEATH_DATE
                 THEN stroke.first_stroke_date ELSE death.DEATH_DATE END)
        WHEN hf_hosp.first_hf_date IS NOT NULL
          AND death.DEATH_DATE IS NOT NULL
            AND death.DEATH_DATE > lm.landmark_date
            AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date,
            CASE WHEN hf_hosp.first_hf_date <= death.DEATH_DATE
                 THEN hf_hosp.first_hf_date ELSE death.DEATH_DATE END)
        WHEN stroke.first_stroke_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date, stroke.first_stroke_date)
        WHEN hf_hosp.first_hf_date IS NOT NULL
          THEN DATEDIFF(day, lm.landmark_date, hf_hosp.first_hf_date)
        WHEN death.DEATH_DATE IS NOT NULL
          AND death.DEATH_DATE > lm.landmark_date
          AND death.DEATH_DATE <= DATEADD(day, ", config$followup_days, ", lm.landmark_date)
          THEN DATEDIFF(day, lm.landmark_date, death.DEATH_DATE)
        ELSE ", config$followup_days, "
      END AS time_to_composite

    INTO #outcomes
    FROM #landmark lm

    -- Stroke: I63.x (ischemic) or I61.x (hemorrhagic) on IP/EI after landmark
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_stroke_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #landmark lm2 ON dx.PATID = lm2.PATID
      WHERE (dx.DX LIKE 'I63%' OR dx.DX LIKE 'I61%')
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI')
        AND dx.ADMIT_DATE > lm2.landmark_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", lm2.landmark_date)
      GROUP BY dx.PATID
    ) stroke ON lm.PATID = stroke.PATID

    -- HF hospitalization: I50.x on IP/EI after landmark
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_hf_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #landmark lm3 ON dx.PATID = lm3.PATID
      WHERE dx.DX LIKE 'I50%'
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI')
        AND dx.ADMIT_DATE > lm3.landmark_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", lm3.landmark_date)
      GROUP BY dx.PATID
    ) hf_hosp ON lm.PATID = hf_hosp.PATID

    -- AF-related hospitalization: I48.x on IP/EI after landmark
    LEFT JOIN (
      SELECT dx.PATID, MIN(dx.ADMIT_DATE) AS first_af_hosp_date
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN CDW.dbo.ENCOUNTER enc ON dx.ENCOUNTERID = enc.ENCOUNTERID
        AND enc.RAW_ENC_TYPE <> 'Legacy Encounter'
      INNER JOIN #landmark lm4 ON dx.PATID = lm4.PATID
      WHERE dx.DX LIKE 'I48%'
        AND dx.DX_TYPE = '10'
        AND enc.ENC_TYPE IN ('IP','EI')
        AND dx.ADMIT_DATE > lm4.landmark_date
        AND dx.ADMIT_DATE <= DATEADD(day, ", config$followup_days, ", lm4.landmark_date)
      GROUP BY dx.PATID
    ) af_hosp ON lm.PATID = af_hosp.PATID

    -- All-cause mortality (deduplicated; date guard for junk dates)
    LEFT JOIN (
      SELECT d.PATID, d.DEATH_DATE,
             ROW_NUMBER() OVER (PARTITION BY d.PATID ORDER BY d.DEATH_DATE) AS rn
      FROM CDW.dbo.DEATH d
      WHERE d.DEATH_DATE BETWEEN '2000-01-01' AND GETDATE()
    ) death ON lm.PATID = death.PATID AND death.rn = 1

    -- CV death flag: patient has a cardiovascular DEATH_CAUSE (I-codes)
    LEFT JOIN (
      SELECT DISTINCT dc.PATID
      FROM CDW.dbo.DEATH_CAUSE dc
      WHERE dc.DEATH_CAUSE LIKE 'I%'
        AND dc.DEATH_CAUSE_CODE = '10'
    ) cv_death ON lm.PATID = cv_death.PATID
    ;
  ")

  # ── Step 5: Confounder extraction ──
  confounders_sql <- paste0("
    -- =========================================================
    -- STEP 5: Confounder Extraction
    -- Baseline covariates measured at or before AF diagnosis (index_date)
    -- Lookback: ", config$lookback_dx, " days for Dx,
    --           ", config$lookback_lab, " days for labs,
    --           ", config$lookback_vital, " days for vitals,
    --           ", config$lookback_rx, " days for concurrent meds
    -- =========================================================

    IF OBJECT_ID('tempdb..#analytic_cohort') IS NOT NULL DROP TABLE #analytic_cohort;

    SELECT
      o.*,

      -- ===== Comorbidities (lookback ", config$lookback_dx, " days before index) =====
      ISNULL(comorbid.chf, 0) AS chf,
      ISNULL(comorbid.hypertension, 0) AS hypertension,
      ISNULL(comorbid.diabetes, 0) AS diabetes,
      ISNULL(comorbid.ckd, 0) AS ckd,
      ISNULL(comorbid.prior_stroke, 0) AS prior_stroke,
      ISNULL(comorbid.prior_tia, 0) AS prior_tia,
      ISNULL(comorbid.vascular_disease, 0) AS vascular_disease,
      ISNULL(comorbid.prior_bleed, 0) AS prior_bleed,
      ISNULL(comorbid.copd, 0) AS copd,
      ISNULL(comorbid.dementia, 0) AS dementia,

      -- ===== CHA2DS2-VASc score (computed from components) =====
      -- All patients are >= 80 so age component = 2
      -- CHF(1) + HTN(1) + Age>=75(2) + DM(1) + Stroke/TIA(2) + Vasc(1) + Female(1)
      (ISNULL(comorbid.chf, 0)
       + ISNULL(comorbid.hypertension, 0)
       + 2  -- age >= 75 guaranteed (all >= 80)
       + ISNULL(comorbid.diabetes, 0)
       + 2 * CASE WHEN ISNULL(comorbid.prior_stroke, 0) = 1
                    OR ISNULL(comorbid.prior_tia, 0) = 1 THEN 1 ELSE 0 END
       + ISNULL(comorbid.vascular_disease, 0)
       + CASE WHEN o.SEX = 'F' THEN 1 ELSE 0 END
      ) AS chadsvasc_score,

      -- ===== Vitals (most recent within ", config$lookback_vital, " days) =====
      vitals.ORIGINAL_BMI AS bmi,
      vitals.SYSTOLIC AS bp_systolic,
      vitals.DIASTOLIC AS bp_diastolic,

      -- ===== Labs (most recent within ", config$lookback_lab, " days) =====
      lab_egfr.RESULT_NUM AS egfr_value,
      lab_hba1c.RESULT_NUM AS hba1c,
      lab_hgb.RESULT_NUM AS hemoglobin,

      -- ===== Concurrent medications (within ", config$lookback_rx, " days of index) =====
      CASE WHEN oac.PATID IS NOT NULL THEN 1 ELSE 0 END AS oac_use,
      CASE WHEN statin.PATID IS NOT NULL THEN 1 ELSE 0 END AS statin_use,
      CASE WHEN acei_arb.PATID IS NOT NULL THEN 1 ELSE 0 END AS acei_arb_use,

      -- ===== Enrollment =====
      enr.ENR_BASIS

    INTO #analytic_cohort
    FROM #outcomes o

    -- ----- Comorbidities (GROUP BY -> safe, 1 row per PATID) -----
    LEFT JOIN (
      SELECT
        dx.PATID,
        MAX(CASE WHEN dx.DX LIKE 'I50%' THEN 1 ELSE 0 END) AS chf,
        MAX(CASE WHEN dx.DX LIKE 'I10%' OR dx.DX LIKE 'I11%'
                   OR dx.DX LIKE 'I12%' OR dx.DX LIKE 'I13%'
                   OR dx.DX LIKE 'I15%' OR dx.DX LIKE 'I16%' THEN 1 ELSE 0 END) AS hypertension,
        MAX(CASE WHEN dx.DX LIKE 'E10%' OR dx.DX LIKE 'E11%'
                   OR dx.DX LIKE 'E13%' THEN 1 ELSE 0 END) AS diabetes,
        MAX(CASE WHEN dx.DX LIKE 'N18%' THEN 1 ELSE 0 END) AS ckd,
        MAX(CASE WHEN dx.DX LIKE 'I63%' THEN 1 ELSE 0 END) AS prior_stroke,
        MAX(CASE WHEN dx.DX LIKE 'G45%' THEN 1 ELSE 0 END) AS prior_tia,
        MAX(CASE WHEN dx.DX LIKE 'I21%' OR dx.DX LIKE 'I25%'
                   OR dx.DX LIKE 'I70%' OR dx.DX LIKE 'I71%'
                   OR dx.DX = 'I73.9' THEN 1 ELSE 0 END) AS vascular_disease,
        MAX(CASE WHEN dx.DX LIKE 'I60%' OR dx.DX LIKE 'I61%'
                   OR dx.DX LIKE 'I62%' OR dx.DX LIKE 'K92%'
                   OR dx.DX IN ('K62.5','K25.0','K25.4','K26.0','K26.4',
                                'K27.0','K27.4','K29.01') THEN 1 ELSE 0 END) AS prior_bleed,
        MAX(CASE WHEN dx.DX LIKE 'J44%' OR dx.DX LIKE 'J43%' THEN 1 ELSE 0 END) AS copd,
        MAX(CASE WHEN dx.DX LIKE 'F01%' OR dx.DX LIKE 'F02%'
                   OR dx.DX LIKE 'F03%' OR dx.DX LIKE 'G30%' THEN 1 ELSE 0 END) AS dementia
      FROM CDW.dbo.DIAGNOSIS dx
      INNER JOIN #outcomes o2 ON dx.PATID = o2.PATID
      WHERE dx.DX_TYPE = '10'
        AND dx.ADMIT_DATE <= o2.index_date
        AND dx.ADMIT_DATE >= DATEADD(day, -", config$lookback_dx, ", o2.index_date)
        AND dx.ADMIT_DATE BETWEEN '2005-01-01' AND GETDATE()
      GROUP BY dx.PATID
    ) comorbid ON o.PATID = comorbid.PATID

    -- ----- Vitals (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ORIGINAL_BMI, SYSTOLIC, DIASTOLIC FROM (
        SELECT v.PATID, v.ORIGINAL_BMI, v.SYSTOLIC, v.DIASTOLIC,
               ROW_NUMBER() OVER (PARTITION BY v.PATID ORDER BY v.MEASURE_DATE DESC) AS rn
        FROM CDW.dbo.VITAL v
        INNER JOIN #outcomes o3 ON v.PATID = o3.PATID
        WHERE v.MEASURE_DATE <= o3.index_date
          AND v.MEASURE_DATE >= DATEADD(day, -", config$lookback_vital, ", o3.index_date)
          AND v.MEASURE_DATE BETWEEN '2005-01-01' AND GETDATE()
      ) v_ranked WHERE rn = 1
    ) vitals ON o.PATID = vitals.PATID

    -- ----- Lab: eGFR (LOINC 48642-3, 62238-1) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o4 ON l.PATID = o4.PATID
        WHERE l.LAB_LOINC IN ('48642-3','62238-1')
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o4.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o4.index_date)
      ) ranked WHERE rn = 1
    ) lab_egfr ON o.PATID = lab_egfr.PATID

    -- ----- Lab: HbA1c (LOINC 4548-4) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o5 ON l.PATID = o5.PATID
        WHERE l.LAB_LOINC = '4548-4'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o5.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o5.index_date)
      ) ranked WHERE rn = 1
    ) lab_hba1c ON o.PATID = lab_hba1c.PATID

    -- ----- Lab: Hemoglobin (LOINC 718-7) -----
    LEFT JOIN (
      SELECT PATID, RESULT_NUM FROM (
        SELECT l.PATID, l.RESULT_NUM,
               ROW_NUMBER() OVER (PARTITION BY l.PATID ORDER BY l.RESULT_DATE DESC) AS rn
        FROM CDW.dbo.LAB_RESULT_CM l
        INNER JOIN #outcomes o6 ON l.PATID = o6.PATID
        WHERE l.LAB_LOINC = '718-7'
          AND l.RESULT_NUM IS NOT NULL
          AND l.RESULT_DATE <= o6.index_date
          AND l.RESULT_DATE >= DATEADD(day, -", config$lookback_lab, ", o6.index_date)
      ) ranked WHERE rn = 1
    ) lab_hgb ON o.PATID = lab_hgb.PATID

    -- ----- Concurrent OAC (DOAC or warfarin) -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o7 ON p.PATID = o7.PATID
      WHERE p.RXNORM_CUI IN (
              -- Apixaban
              '1364435','1364445','1364441','1364447',
              -- Rivaroxaban
              '1114198','1232082','1232086','2059015',
              '1114202','1232084','1232088','2059017',
              -- Dabigatran
              '1037179','1723476','1037045','1037181','1723478','1037049',
              -- Edoxaban
              '1599543','1599551','1599555','1599549','1599553','1599557',
              -- Warfarin (representative subset)
              '855288','855302','855312','855318','855324','855332','855338',
              '855344','855296','855350'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o7.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o7.index_date
    ) oac ON o.PATID = oac.PATID

    -- ----- Concurrent statin -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o8 ON p.PATID = o8.PATID
      WHERE p.RXNORM_CUI IN (
              -- atorvastatin 10/20/40/80 (SCD)
              '259255','617311','617314','617318',
              -- rosuvastatin 5/10/20/40 (SCD)
              '859747','859751','859755','859749',
              -- simvastatin 10/20/40/80 (SCD)
              '198211','314231','152923','312961'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o8.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o8.index_date
    ) statin ON o.PATID = statin.PATID

    -- ----- Concurrent ACEi / ARB -----
    LEFT JOIN (
      SELECT DISTINCT p.PATID
      FROM CDW.dbo.PRESCRIBING p
      INNER JOIN #outcomes o9 ON p.PATID = o9.PATID
      WHERE p.RXNORM_CUI IN (
              -- lisinopril 5/10/20/40 (SCD)
              '314076','314077','197884','314078',
              -- losartan 25/50/100 (SCD)
              '979467','979471','979480',
              -- valsartan 40/80/160/320 (SCD)
              '349483','349199','349201','349485'
            )
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE)
            >= DATEADD(day, -", config$lookback_rx, ", o9.index_date)
        AND COALESCE(p.RX_START_DATE, p.RX_ORDER_DATE) <= o9.index_date
    ) acei_arb ON o.PATID = acei_arb.PATID

    -- ----- Enrollment (exactly 1 row per patient via ROW_NUMBER) -----
    LEFT JOIN (
      SELECT PATID, ENR_BASIS FROM (
        SELECT enr.PATID, enr.ENR_BASIS,
               ROW_NUMBER() OVER (PARTITION BY enr.PATID ORDER BY enr.ENR_START_DATE DESC) AS rn
        FROM CDW.dbo.ENROLLMENT enr
        INNER JOIN #outcomes o10 ON enr.PATID = o10.PATID
        WHERE o10.index_date BETWEEN enr.ENR_START_DATE
          AND COALESCE(enr.ENR_END_DATE, '9999-12-31')
      ) ranked WHERE rn = 1
    ) enr ON o.PATID = enr.PATID
    ;
  ")

  return(list(
    eligibility  = eligibility_sql,
    treatment    = treatment_sql,
    landmark     = landmark_sql,
    outcome      = outcome_sql,
    confounders  = confounders_sql
  ))
}
```


## 3. Execute SQL and Pull Data

```{r}
#| label: helper-count

count_temp <- function(con, tbl) {
  res <- dbGetQuery(con, paste("SELECT COUNT(DISTINCT PATID) AS n FROM", tbl))
  res$n[1]
}
```

```{r}
#| label: pull-cohort

pull_analytic_cohort <- function(con, config) {
  sql_parts <- build_cohort_sql(config)
  consort <- list()

  message("=== Executing Cohort SQL ===")

  message("  Step 1: Building eligible population (newly diagnosed AF, age >= 80)...")
  dbExecute(con, sql_parts$eligibility)

  consort_raw <- dbGetQuery(con, "SELECT * FROM #consort_counts")
  names(consort_raw) <- tolower(names(consort_raw))
  consort$incident_af  <- consort_raw$n_incident_af[1]
  consort$eligible     <- consort_raw$n_eligible[1]
  try(dbExecute(con, "DROP TABLE IF EXISTS #consort_counts"), silent = TRUE)

  message(sprintf("    -> Incident AF:       %d", consort$incident_af))
  message(sprintf("    -> Eligible (age>=80): %d", consort$eligible))

  message("  Step 2: Assigning treatment (rhythm vs rate control)...")
  dbExecute(con, sql_parts$treatment)
  consort$treatment <- count_temp(con, "#treatment")
  message(sprintf("    -> #treatment: %d patients", consort$treatment))

  message("  Step 3: Applying landmark filter (survived 12 months)...")
  dbExecute(con, sql_parts$landmark)
  consort$landmark <- count_temp(con, "#landmark")
  message(sprintf("    -> #landmark: %d patients", consort$landmark))

  message("  Step 4: Ascertaining outcomes...")
  dbExecute(con, sql_parts$outcome)
  consort$outcomes <- count_temp(con, "#outcomes")
  message(sprintf("    -> #outcomes: %d patients", consort$outcomes))

  # Retrieve DEATH_CAUSE completeness check
  death_check <- dbGetQuery(con, "SELECT * FROM #death_cause_check")
  names(death_check) <- tolower(names(death_check))
  consort$n_deaths_in_followup   <- death_check$n_deaths_in_followup[1]
  consort$n_deaths_with_cause    <- death_check$n_deaths_with_cause[1]
  try(dbExecute(con, "DROP TABLE IF EXISTS #death_cause_check"), silent = TRUE)

  message(sprintf("    -> Deaths in follow-up: %d, with cause code: %d",
                  consort$n_deaths_in_followup, consort$n_deaths_with_cause))

  message("  Step 5: Extracting confounders and building analytic cohort...")
  dbExecute(con, sql_parts$confounders)
  cohort <- dbGetQuery(con, "SELECT * FROM #analytic_cohort")

  names(cohort) <- tolower(names(cohort))

  consort$analytic      <- nrow(cohort)
  consort$n_rhythm      <- sum(cohort$treatment == 1, na.rm = TRUE)
  consort$n_rate        <- sum(cohort$treatment == 0, na.rm = TRUE)

  message(sprintf(
    "  Analytic cohort: %d patients (%d rhythm control, %d rate control)",
    consort$analytic, consort$n_rhythm, consort$n_rate
  ))

  for (tbl in c("#eligible", "#treatment", "#landmark", "#outcomes",
                "#analytic_cohort")) {
    try(dbExecute(con, paste("DROP TABLE IF EXISTS", tbl)), silent = TRUE)
  }

  attr(cohort, "consort") <- consort
  return(cohort)
}
```


## 3a. CONSORT Flow Diagram

```{r}
#| label: consort-text

print_consort_table <- function(consort) {
  steps <- tibble::tibble(
    Step = c("1a. Incident AF patients (first I48.x, 365-day washout)",
             "1b. Eligible (age >= 80, active in CDW)",
             "2.  Treatment assigned (rhythm vs rate control)",
             "3.  Survived 12-month landmark",
             "4.  Outcomes ascertained",
             "5.  Analytic cohort (with confounders)"),
    N    = c(consort$incident_af,
             consort$eligible,
             consort$treatment,
             consort$landmark,
             consort$outcomes,
             consort$analytic),
    Excluded = c(NA_integer_,
                 consort$incident_af - consort$eligible,
                 consort$eligible    - consort$treatment,
                 consort$treatment   - consort$landmark,
                 consort$landmark    - consort$outcomes,
                 consort$outcomes    - consort$analytic)
  )
  message("\n=== CONSORT Flow (text) ===")
  print(steps, n = Inf)
  invisible(steps)
}
```

```{r}
#| label: consort-diagram

render_consort_diagram <- function(consort) {

  grid::grid.newpage()
  fmt <- function(x) format(x, big.mark = ",")

  box_w  <- grid::unit(0.50, "npc")
  box_h  <- grid::unit(0.055, "npc")
  excl_w <- grid::unit(0.28, "npc")
  excl_h <- grid::unit(0.040, "npc")
  x_main <- 0.38
  x_excl <- 0.82

  y_pos <- seq(0.92, 0.14, length.out = 7)

  draw_box <- function(x, y, w, h, label, fill = "white") {
    grid::grid.rect(x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    width = w, height = h,
                    gp = grid::gpar(fill = fill, col = "grey30", lwd = 1.5))
    grid::grid.text(label, x = grid::unit(x, "npc"), y = grid::unit(y, "npc"),
                    gp = grid::gpar(fontsize = 9, fontface = "bold"))
  }

  draw_arrow <- function(x1, y1, x2, y2) {
    grid::grid.lines(x = grid::unit(c(x1, x2), "npc"),
                     y = grid::unit(c(y1, y2), "npc"),
                     arrow = grid::arrow(length = grid::unit(0.015, "npc"),
                                         type = "closed"),
                     gp = grid::gpar(fill = "grey30", col = "grey30"))
  }

  draw_excl <- function(y_from, y_to, n, reason) {
    if (n <= 0) return(invisible(NULL))
    mid_y <- mean(c(y_from - 0.028, y_to + 0.028))
    draw_arrow(x_main + 0.25, mid_y, x_excl - 0.14, mid_y)
    draw_box(x_excl, mid_y, excl_w, excl_h,
             sprintf("Excluded: %s\n(%s)", fmt(n), reason),
             fill = "#FFF3F3")
  }

  grid::grid.text("CONSORT Flow Diagram --- Early Rhythm vs Rate Control in Elderly AF",
                  x = 0.5, y = 0.97,
                  gp = grid::gpar(fontsize = 12, fontface = "bold"))

  labels <- c(
    sprintf("1a. Incident AF\nn = %s", fmt(consort$incident_af)),
    sprintf("1b. Eligible (age >= 80)\nn = %s", fmt(consort$eligible)),
    sprintf("2. Treatment Assigned\nn = %s", fmt(consort$treatment)),
    sprintf("3. Survived 12-Month Landmark\nn = %s", fmt(consort$landmark)),
    sprintf("4. Outcomes Ascertained\nn = %s", fmt(consort$outcomes)),
    sprintf("5. Analytic Cohort\nn = %s", fmt(consort$analytic))
  )
  fills <- c(rep("white", 5), "#F0FFF0")

  for (i in seq_along(labels)) {
    draw_box(x_main, y_pos[i], box_w, box_h, labels[i], fill = fills[i])
    if (i < length(labels)) {
      draw_arrow(x_main, y_pos[i] - 0.028, x_main, y_pos[i + 1] + 0.028)
    }
  }

  excl_counts  <- c(consort$incident_af - consort$eligible,
                    consort$eligible    - consort$treatment,
                    consort$treatment   - consort$landmark,
                    consort$landmark    - consort$outcomes,
                    consort$outcomes    - consort$analytic)
  excl_reasons <- c("age < 80 / no active encounter",
                    "no rhythm or rate control Rx",
                    "died within 12 months",
                    "outcome ascertainment",
                    "confounder join")
  for (i in seq_along(excl_counts)) {
    draw_excl(y_pos[i], y_pos[i + 1], excl_counts[i], excl_reasons[i])
  }

  # Split into treatment arms
  draw_arrow(x_main - 0.10, y_pos[6] - 0.028, x_main - 0.18, y_pos[7] + 0.020)
  draw_arrow(x_main + 0.10, y_pos[6] - 0.028, x_main + 0.18, y_pos[7] + 0.020)

  draw_box(x_main - 0.20, y_pos[7], grid::unit(0.26, "npc"), box_h,
           sprintf("Rhythm Control\nn = %s", fmt(consort$n_rhythm)),
           fill = "#F0F0FF")
  draw_box(x_main + 0.20, y_pos[7], grid::unit(0.26, "npc"), box_h,
           sprintf("Rate Control\nn = %s", fmt(consort$n_rate)),
           fill = "#F0F0FF")

  invisible(NULL)
}
```


## 4. Data Preparation

```{r}
#| label: prepare-cohort

prepare_cohort <- function(cohort, consort) {

  # Determine whether to use CV death or all-cause death in composite
  # Use CV death if >= 30% of deaths have a cause code; otherwise all-cause
  use_cv_death <- FALSE
  if (!is.null(consort$n_deaths_in_followup) &&
      consort$n_deaths_in_followup > 0) {
    cv_pct <- consort$n_deaths_with_cause / consort$n_deaths_in_followup
    message(sprintf("DEATH_CAUSE completeness: %d / %d = %.1f%%",
                    consort$n_deaths_with_cause,
                    consort$n_deaths_in_followup,
                    100 * cv_pct))
    if (cv_pct >= 0.30) {
      use_cv_death <- TRUE
      message("  -> Using CV death in composite (DEATH_CAUSE >= 30% populated)")
    } else {
      message("  -> Using ALL-CAUSE death in composite (DEATH_CAUSE < 30%)")
    }
  } else {
    message("No deaths in follow-up; death component is moot.")
  }

  cohort <- cohort |>
    mutate(
      sex_cat = factor(sex, levels = c("F", "M"),
                       labels = c("Female", "Male")),
      race_cat = case_when(
        race == "03" ~ "Black",
        race == "05" ~ "White",
        race == "02" ~ "Asian",
        race %in% c("01", "04", "06", "07") ~ "Other",
        TRUE ~ "Unknown"
      ) |> factor(),
      hispanic_cat = factor(
        if_else(hispanic == "Y", "Hispanic", "Non-Hispanic")
      ),
      enr_basis_cat = factor(
        case_when(
          enr_basis %in% c("I", "A") ~ enr_basis,
          TRUE ~ "Other/Unknown"
        )
      ),
      age_group = factor(
        if_else(age_at_index >= 85, "85+", "80-84"),
        levels = c("80-84", "85+")
      ),
      chadsvasc_cat = factor(
        if_else(chadsvasc_score >= 5, "CHA2DS2-VASc >= 5", "CHA2DS2-VASc < 5")
      ),
      treatment = as.integer(treatment),
      treatment_label = factor(
        if_else(treatment == 1, "Rhythm Control", "Rate Control"),
        levels = c("Rate Control", "Rhythm Control")
      )
    )

  # Recompute composite based on CV death vs all-cause death decision
  if (use_cv_death) {
    cohort <- cohort |>
      mutate(
        death_component = cv_death_event,
        composite_event = as.integer(
          stroke_event == 1 | hf_hosp_event == 1 | cv_death_event == 1
        )
      )
  } else {
    cohort <- cohort |>
      mutate(
        death_component = death_event,
        composite_event = composite_allcause_event
      )
  }

  # Ensure minimum follow-up time of 1 day
  cohort <- cohort |>
    mutate(
      time_to_composite = pmax(time_to_composite, 1),
      time_to_stroke    = pmax(time_to_stroke, 1),
      time_to_hf_hosp   = pmax(time_to_hf_hosp, 1),
      time_to_death     = pmax(time_to_death, 1),
      time_to_af_hosp   = pmax(time_to_af_hosp, 1)
    ) |>
    filter(!is.na(treatment))

  # Impute missing continuous covariates with median
  cohort <- cohort |>
    mutate(
      bmi         = if_else(is.na(bmi), median(bmi, na.rm = TRUE), bmi),
      bp_systolic = if_else(is.na(bp_systolic),
                            median(bp_systolic, na.rm = TRUE), bp_systolic),
      bp_diastolic = if_else(is.na(bp_diastolic),
                             median(bp_diastolic, na.rm = TRUE), bp_diastolic),
      egfr_value  = if_else(is.na(egfr_value),
                            median(egfr_value, na.rm = TRUE), egfr_value),
      hba1c       = if_else(is.na(hba1c),
                            median(hba1c, na.rm = TRUE), hba1c),
      hemoglobin  = if_else(is.na(hemoglobin),
                            median(hemoglobin, na.rm = TRUE), hemoglobin)
    )

  attr(cohort, "use_cv_death") <- use_cv_death
  return(cohort)
}
```


## 5. Causal Analysis Functions (IPW)

```{r}
#| label: confounders

confounders <- c(
  "age_at_index", "sex_cat", "race_cat", "hispanic_cat",
  "bmi", "bp_systolic", "bp_diastolic",
  "egfr_value", "hba1c", "hemoglobin",
  "chf", "hypertension", "diabetes", "ckd",
  "prior_stroke", "prior_tia", "vascular_disease",
  "prior_bleed", "copd", "dementia",
  "chadsvasc_score",
  "oac_use", "statin_use", "acei_arb_use",
  "enr_basis_cat",
  "index_year"
)

build_ps_formula <- function(confounders, cohort) {
  dropped <- character()
  keep    <- character()

  for (v in confounders) {
    col <- cohort[[v]]
    if (is.null(col)) {
      dropped <- c(dropped, v)
      next
    }
    if (is.factor(col) || is.character(col)) {
      n_levels <- length(unique(na.omit(col)))
      if (n_levels < 2) {
        dropped <- c(dropped, v)
        next
      }
    } else {
      if (sd(col, na.rm = TRUE) == 0 || all(is.na(col))) {
        dropped <- c(dropped, v)
        next
      }
    }
    keep <- c(keep, v)
  }

  if (length(dropped) > 0) {
    message(sprintf("  PS model: dropped %d constant/single-level variable(s): %s",
                    length(dropped), paste(dropped, collapse = ", ")))
  }
  message(sprintf("  PS model: using %d confounder(s)", length(keep)))

  as.formula(paste("treatment ~", paste(keep, collapse = " + ")))
}
```

```{r}
#| label: ipw-analysis

run_ipw_analysis <- function(cohort, confounders, config) {
  message("\n=== IPW Analysis ===")

  # --- Guard: treatment must have at least 2 arms ---
  n_arms <- length(unique(cohort$treatment))
  if (n_arms < 2) {
    stop(sprintf(
      "Cannot run IPW: treatment variable has only %d unique value(s) (%s). ",
      n_arms, paste(unique(cohort$treatment), collapse = ", ")),
      "Check the CONSORT diagram -- one treatment arm may have been ",
      "completely filtered out during eligibility.", call. = FALSE)
  }
  message(sprintf("Treatment arms: Rhythm control (n=%d), Rate control (n=%d)",
                  sum(cohort$treatment == 1), sum(cohort$treatment == 0)))

  # --- Build PS formula, dropping single-level factors ---
  ps_formula <- build_ps_formula(confounders, cohort)

  # --- Fit propensity score model and compute weights ---
  message("Fitting propensity score model (logistic regression)...")
  weights <- weightit(
    ps_formula,
    data     = cohort,
    method   = "glm",
    estimand = config$estimand
  )

  # --- Balance diagnostics ---
  message("Checking covariate balance...")
  bal <- bal.tab(weights, stats = c("m", "v"), thresholds = c(m = 0.1))
  print(bal)

  p_love <- love.plot(weights, threshold = 0.1, abs = TRUE,
                      title = "Covariate Balance: Rhythm vs Rate Control (IPW)")

  cohort$ps <- weights$ps
  p_ps <- ggplot(cohort, aes(x = ps, fill = treatment_label)) +
    geom_density(alpha = 0.5) +
    labs(x = "Propensity Score (P[Rhythm Control])", y = "Density",
         title = "Propensity Score Distribution by Treatment Group",
         fill = "Treatment") +
    theme_minimal()

  ess_fn <- function(w) sum(w)^2 / sum(w^2)
  message(sprintf("Effective sample size: Rhythm=%.1f, Rate=%.1f",
                  ess_fn(weights$weights[cohort$treatment == 1]),
                  ess_fn(weights$weights[cohort$treatment == 0])))

  cohort$ipw <- weights$weights

  # --- Primary: Composite (CV death/all-cause death + stroke + HF hosp) ---
  use_cv_death <- attr(cohort, "use_cv_death")
  death_label <- if (isTRUE(use_cv_death)) "CV death" else "all-cause death"
  message(sprintf("\n--- Primary Outcome: Composite (%s + stroke + HF hosp) ---",
                  death_label))
  fit_composite <- coxph(
    Surv(time_to_composite, composite_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_composite))

  # --- Secondary: Stroke ---
  message("\n--- Secondary Outcome: Stroke (I63.x + I61.x) ---")
  fit_stroke <- coxph(
    Surv(time_to_stroke, stroke_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_stroke))

  # --- Secondary: HF hospitalization ---
  message("\n--- Secondary Outcome: HF Hospitalization ---")
  fit_hf_hosp <- coxph(
    Surv(time_to_hf_hosp, hf_hosp_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_hf_hosp))

  # --- Secondary: All-cause mortality ---
  message("\n--- Secondary Outcome: All-Cause Mortality ---")
  fit_death <- coxph(
    Surv(time_to_death, death_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_death))

  # --- Secondary: AF-related hospitalization ---
  message("\n--- Secondary Outcome: AF-Related Hospitalization ---")
  fit_af_hosp <- coxph(
    Surv(time_to_af_hosp, af_hosp_event) ~ treatment,
    data = cohort, weights = ipw, robust = TRUE
  )
  print(summary(fit_af_hosp))

  # --- Weighted KM curves ---
  km_composite <- survfit(
    Surv(time_to_composite, composite_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_composite <- ggsurvplot(
    km_composite, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = sprintf("Composite (%s + Stroke + HF Hosp): Rhythm vs Rate (IPW)",
                    death_label),
    xlab = "Days from 12-Month Landmark",
    ylab = "Event-Free Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  km_death <- survfit(
    Surv(time_to_death, death_event) ~ treatment_label,
    data = cohort, weights = ipw
  )
  p_km_death <- ggsurvplot(
    km_death, data = cohort,
    risk.table = TRUE, pval = TRUE,
    title = "All-Cause Mortality: Rhythm vs Rate Control (IPW)",
    xlab = "Days from 12-Month Landmark",
    ylab = "Survival Probability",
    palette = c("#E7B800", "#2E9FDF")
  )

  return(list(
    weights = weights,
    balance = bal,
    fits = list(
      composite = fit_composite,
      stroke    = fit_stroke,
      hf_hosp   = fit_hf_hosp,
      death     = fit_death,
      af_hosp   = fit_af_hosp
    ),
    plots = list(
      love         = p_love,
      ps_dist      = p_ps,
      km_composite = p_km_composite,
      km_death     = p_km_death
    ),
    cohort = cohort
  ))
}
```


## 6. Subgroup Analyses

```{r}
#| label: subgroup-analysis

run_subgroup_analyses <- function(cohort, confounders, config) {
  message("\n=== Subgroup Analyses ===")

  subgroups <- list(
    "Age 80-84"           = cohort$age_group == "80-84",
    "Age >= 85"           = cohort$age_group == "85+",
    "Female"              = cohort$sex_cat == "Female",
    "Male"                = cohort$sex_cat == "Male",
    "CHA2DS2-VASc < 5"   = cohort$chadsvasc_score < 5,
    "CHA2DS2-VASc >= 5"  = cohort$chadsvasc_score >= 5,
    "Heart failure"       = cohort$chf == 1,
    "No heart failure"    = cohort$chf == 0,
    "CKD"                 = cohort$ckd == 1,
    "No CKD"              = cohort$ckd == 0
  )

  results <- tibble::tibble(
    Subgroup = character(),
    N = integer(),
    N_rhythm = integer(),
    N_rate = integer(),
    N_events = integer(),
    HR = numeric(),
    CI_lo = numeric(),
    CI_hi = numeric(),
    P_value = numeric()
  )

  for (sg_name in names(subgroups)) {
    sg_data <- cohort[subgroups[[sg_name]], ]

    # Guard: need 2 arms and at least 10 patients per arm
    if (length(unique(sg_data$treatment)) < 2 ||
        sum(sg_data$treatment == 1) < 10 ||
        sum(sg_data$treatment == 0) < 10) {
      message(sprintf("  Skipping '%s': insufficient patients in one arm", sg_name))
      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_rhythm = sum(sg_data$treatment == 1),
        N_rate = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
      next
    }

    # Re-fit PS within subgroup
    tryCatch({
      ps_formula <- build_ps_formula(confounders, sg_data)
      w <- weightit(ps_formula, data = sg_data, method = "glm",
                    estimand = config$estimand)
      sg_data$ipw <- w$weights

      fit <- coxph(Surv(time_to_composite, composite_event) ~ treatment,
                   data = sg_data, weights = ipw, robust = TRUE)
      s <- summary(fit)

      results <- bind_rows(results, tibble::tibble(
        Subgroup = sg_name,
        N = nrow(sg_data),
        N_rhythm = sum(sg_data$treatment == 1),
        N_rate = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = s$conf.int[1, 1],
        CI_lo = s$conf.int[1, 3],
        CI_hi = s$conf.int[1, 4],
        P_value = s$coefficients[1, "Pr(>|z|)"]
      ))
      message(sprintf("  %s: HR=%.2f (%.2f-%.2f), N=%d",
                      sg_name, s$conf.int[1,1], s$conf.int[1,3],
                      s$conf.int[1,4], nrow(sg_data)))
    }, error = function(e) {
      message(sprintf("  Error in '%s': %s", sg_name, e$message))
      results <<- bind_rows(results, tibble::tibble(
        Subgroup = sg_name, N = nrow(sg_data),
        N_rhythm = sum(sg_data$treatment == 1),
        N_rate = sum(sg_data$treatment == 0),
        N_events = sum(sg_data$composite_event),
        HR = NA_real_, CI_lo = NA_real_, CI_hi = NA_real_, P_value = NA_real_
      ))
    })
  }

  return(results)
}
```


## 7. Sensitivity Analyses

```{r}
#| label: sensitivity-analysis

run_sensitivity <- function(ipw_results, cohort, confounders, config) {
  message("\n=== Sensitivity Analyses ===")

  sensitivity <- list()

  # --- 7a. E-value for composite outcome ---
  message("\n--- E-value Analysis (Composite Outcome) ---")
  fit_composite <- ipw_results$fits$composite
  s <- summary(fit_composite)
  hr  <- s$conf.int[1, 1]
  lo  <- s$conf.int[1, 3]
  hi  <- s$conf.int[1, 4]

  evals <- evalues.HR(hr, lo = lo, hi = hi, rare = TRUE)
  sensitivity$evalue_composite <- evals
  message("E-value for composite outcome:")
  print(evals)

  # --- 7b. E-value for all-cause mortality ---
  message("\n--- E-value Analysis (All-Cause Mortality) ---")
  fit_death <- ipw_results$fits$death
  s_death <- summary(fit_death)
  hr_d  <- s_death$conf.int[1, 1]
  lo_d  <- s_death$conf.int[1, 3]
  hi_d  <- s_death$conf.int[1, 4]

  evals_death <- evalues.HR(hr_d, lo = lo_d, hi = hi_d, rare = TRUE)
  sensitivity$evalue_death <- evals_death
  message("E-value for all-cause mortality:")
  print(evals_death)

  # --- 7c. PS trimming (5th-95th percentile) ---
  message("\n--- PS Trimming (5th-95th Percentile) ---")
  ps_lo <- quantile(cohort$ps, 0.05, na.rm = TRUE)
  ps_hi <- quantile(cohort$ps, 0.95, na.rm = TRUE)
  cohort_trimmed <- cohort |>
    filter(ps >= ps_lo & ps <= ps_hi)

  message(sprintf("  Trimmed from %d to %d patients (PS range: %.3f-%.3f)",
                  nrow(cohort), nrow(cohort_trimmed), ps_lo, ps_hi))

  if (length(unique(cohort_trimmed$treatment)) < 2) {
    message("WARNING: After PS trimming, only 1 arm remains. Skipping.")
    sensitivity$trimmed_fit <- NULL
  } else {
    ps_formula_trim <- build_ps_formula(confounders, cohort_trimmed)
    w_trim <- weightit(ps_formula_trim, data = cohort_trimmed,
                       method = "glm", estimand = config$estimand)
    cohort_trimmed$ipw <- w_trim$weights

    fit_trim <- coxph(
      Surv(time_to_composite, composite_event) ~ treatment,
      data = cohort_trimmed, weights = ipw, robust = TRUE
    )
    sensitivity$trimmed_fit <- fit_trim
    message("Trimmed analysis (composite):")
    print(summary(fit_trim))
  }

  # --- 7d. Post-Epic era only (2021+) ---
  message("\n--- Post-Epic Era (2021+) ---")
  cohort_epic <- cohort |> filter(index_year >= 2021)

  if (length(unique(cohort_epic$treatment)) < 2 ||
      min(table(cohort_epic$treatment)) < 10) {
    message("WARNING: Insufficient patients in post-Epic era. Skipping.")
    sensitivity$epic_fit <- NULL
  } else {
    ps_formula_epic <- build_ps_formula(confounders, cohort_epic)
    w_epic <- weightit(ps_formula_epic, data = cohort_epic,
                       method = "glm", estimand = config$estimand)
    cohort_epic$ipw <- w_epic$weights

    fit_epic <- coxph(
      Surv(time_to_composite, composite_event) ~ treatment,
      data = cohort_epic, weights = ipw, robust = TRUE
    )
    sensitivity$epic_fit <- fit_epic
    message(sprintf("Post-Epic analysis (N=%d):", nrow(cohort_epic)))
    print(summary(fit_epic))
  }

  # --- 7e. All-cause death composite (if primary used CV death) ---
  use_cv_death <- attr(cohort, "use_cv_death")
  if (isTRUE(use_cv_death)) {
    message("\n--- Sensitivity: All-Cause Death in Composite ---")
    cohort_ac <- cohort |>
      mutate(
        composite_event_ac = composite_allcause_event
      )

    fit_ac <- coxph(
      Surv(time_to_composite, composite_event_ac) ~ treatment,
      data = cohort_ac, weights = ipw, robust = TRUE
    )
    sensitivity$allcause_composite_fit <- fit_ac
    message("All-cause death composite:")
    print(summary(fit_ac))
  } else {
    sensitivity$allcause_composite_fit <- NULL
  }

  return(sensitivity)
}
```


## 8. Summary Helper

```{r}
#| label: summary-helper

format_hr <- function(fit) {
  s <- summary(fit)
  sprintf("HR = %.2f (95%% CI: %.2f-%.2f), p = %.4f",
          s$conf.int[1, 1], s$conf.int[1, 3], s$conf.int[1, 4],
          s$coefficients[1, "Pr(>|z|)"])
}
```


<!-- ============================================================= -->
<!-- PART 2: Execution & Results (Sections 9+)                      -->
<!-- Each section calls its function and displays results inline    -->
<!-- ============================================================= -->


## 9. Connect and Pull Cohort

```{r}
#| label: exec-connect

con <- connect_cdw()
cohort_raw <- pull_analytic_cohort(con, config)
consort <- attr(cohort_raw, "consort")
dbDisconnect(con)
```

### CONSORT Flow (Text)

```{r}
#| label: exec-consort-text

consort_table <- print_consort_table(consort)
knitr::kable(consort_table, caption = "CONSORT Patient Flow")
```

### CONSORT Flow Diagram

```{r}
#| label: fig-consort
#| fig-cap: "CONSORT flow diagram showing patient attrition at each cohort-building step."
#| fig-width: 10
#| fig-height: 12

render_consort_diagram(consort)
```

### Empty Cohort Guard

```{r}
#| label: exec-empty-guard

if (nrow(cohort_raw) == 0) {
  message("*** STOPPING: Analytic cohort has 0 patients. ***")
  message("Review the CONSORT diagram above to identify where patients were lost.")
  knitr::knit_exit()
}
```


## 10. Prepare Data and Table 1

```{r}
#| label: exec-prepare

cohort <- prepare_cohort(cohort_raw, consort)

message(sprintf("Prepared cohort: %d patients (%d rhythm control, %d rate control)",
                nrow(cohort),
                sum(cohort$treatment == 1),
                sum(cohort$treatment == 0)))
```

### Treatment Arms Guard

```{r}
#| label: exec-arms-guard

n_arms <- length(unique(cohort$treatment))
if (n_arms < 2) {
  message(sprintf("*** STOPPING: treatment has only %d unique value(s). ***", n_arms))
  message("Cannot proceed with IPW analysis.")
  knitr::knit_exit()
}
```

### Table 1: Baseline Characteristics

```{r}
#| label: exec-table1

table1_data <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = as.character(n()),
    `Age (mean +/- SD)` = sprintf("%.1f +/- %.1f", mean(age_at_index), sd(age_at_index)),
    `Age 80-84 (%)` = sprintf("%.1f", 100 * mean(age_group == "80-84")),
    `Age >= 85 (%)` = sprintf("%.1f", 100 * mean(age_group == "85+")),
    `Female (%)` = sprintf("%.1f", 100 * mean(sex_cat == "Female")),
    `White (%)` = sprintf("%.1f", 100 * mean(race_cat == "White")),
    `Black (%)` = sprintf("%.1f", 100 * mean(race_cat == "Black")),
    `Hispanic (%)` = sprintf("%.1f", 100 * mean(hispanic_cat == "Hispanic")),
    `eGFR (mean)` = sprintf("%.1f", mean(egfr_value, na.rm = TRUE)),
    `BMI (mean)` = sprintf("%.1f", mean(bmi, na.rm = TRUE)),
    `SBP (mean)` = sprintf("%.0f", mean(bp_systolic, na.rm = TRUE)),
    `HbA1c (mean)` = sprintf("%.1f", mean(hba1c, na.rm = TRUE)),
    `Hemoglobin (mean)` = sprintf("%.1f", mean(hemoglobin, na.rm = TRUE)),
    `CHA2DS2-VASc (mean)` = sprintf("%.1f", mean(chadsvasc_score, na.rm = TRUE)),
    `CHF (%)` = sprintf("%.1f", 100 * mean(chf)),
    `Hypertension (%)` = sprintf("%.1f", 100 * mean(hypertension)),
    `Diabetes (%)` = sprintf("%.1f", 100 * mean(diabetes)),
    `CKD (%)` = sprintf("%.1f", 100 * mean(ckd)),
    `Prior stroke (%)` = sprintf("%.1f", 100 * mean(prior_stroke)),
    `Prior TIA (%)` = sprintf("%.1f", 100 * mean(prior_tia)),
    `Prior bleeding (%)` = sprintf("%.1f", 100 * mean(prior_bleed)),
    `COPD (%)` = sprintf("%.1f", 100 * mean(copd)),
    `Dementia (%)` = sprintf("%.1f", 100 * mean(dementia)),
    `OAC use (%)` = sprintf("%.1f", 100 * mean(oac_use)),
    `Statin (%)` = sprintf("%.1f", 100 * mean(statin_use)),
    `ACEi/ARB (%)` = sprintf("%.1f", 100 * mean(acei_arb_use)),
    .groups = "drop"
  ) |>
  pivot_longer(-treatment_label, names_to = "Characteristic", values_to = "Value") |>
  pivot_wider(names_from = treatment_label, values_from = Value)

knitr::kable(table1_data, caption = "Table 1: Baseline Characteristics by Treatment Group")
```

### Event Rates

```{r}
#| label: exec-events

events <- cohort |>
  group_by(treatment_label) |>
  summarise(
    N = n(),
    `Composite events` = sum(composite_event),
    `Stroke events` = sum(stroke_event),
    `HF hosp events` = sum(hf_hosp_event),
    `Deaths (all-cause)` = sum(death_event),
    `CV deaths` = sum(cv_death_event),
    `AF hosp events` = sum(af_hosp_event),
    .groups = "drop"
  )
knitr::kable(events, caption = "Outcome Event Counts by Treatment Group")
```


## 11. IPW Analysis

```{r}
#| label: exec-ipw

ipw_results <- run_ipw_analysis(cohort, confounders, config)
cohort <- ipw_results$cohort
```

### Love Plot (Covariate Balance)

```{r}
#| label: fig-love-plot
#| fig-cap: "Absolute standardized mean differences before and after IPW."
#| fig-width: 10
#| fig-height: 8

print(ipw_results$plots$love)
```

### Propensity Score Distribution

```{r}
#| label: fig-ps-dist
#| fig-cap: "Propensity score distributions by treatment group."
#| fig-width: 8
#| fig-height: 5

print(ipw_results$plots$ps_dist)
```


## 12. Cox Model Results and Kaplan-Meier Curves

### Summary of All Outcomes

```{r}
#| label: exec-results-summary

use_cv_death <- attr(cohort, "use_cv_death")
death_label <- if (isTRUE(use_cv_death)) "CV death" else "all-cause death"

outcome_summary <- tibble::tibble(
  Outcome = c(sprintf("Composite (%s + stroke + HF hosp)", death_label),
              "Stroke (ischemic + hemorrhagic)",
              "HF hospitalization",
              "All-cause mortality",
              "AF-related hospitalization"),
  Result = c(
    format_hr(ipw_results$fits$composite),
    format_hr(ipw_results$fits$stroke),
    format_hr(ipw_results$fits$hf_hosp),
    format_hr(ipw_results$fits$death),
    format_hr(ipw_results$fits$af_hosp)
  )
)
knitr::kable(outcome_summary,
             caption = "Weighted Cox PH Results: Rhythm Control (treatment=1) vs Rate Control (ref)")
```

### Kaplan-Meier: Composite Outcome

```{r}
#| label: fig-km-composite
#| fig-cap: "Weighted Kaplan-Meier curves for the composite outcome."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_composite)
```

### Kaplan-Meier: All-Cause Mortality

```{r}
#| label: fig-km-death
#| fig-cap: "Weighted Kaplan-Meier curves for all-cause mortality."
#| fig-width: 9
#| fig-height: 7

print(ipw_results$plots$km_death)
```


## 13. Subgroup Analyses

```{r}
#| label: exec-subgroups

subgroup_results <- run_subgroup_analyses(cohort, confounders, config)
knitr::kable(subgroup_results |>
               mutate(across(where(is.numeric), ~round(., 3))),
             caption = "Subgroup Analysis: Composite Outcome (Rhythm vs Rate Control)")
```

### Forest Plot

```{r}
#| label: fig-forest
#| fig-cap: "Forest plot of subgroup analyses for the composite outcome."
#| fig-width: 10
#| fig-height: 8

sg_plot_data <- subgroup_results |> filter(!is.na(HR))

if (nrow(sg_plot_data) > 0) {
  ggplot(sg_plot_data, aes(x = HR, y = reorder(Subgroup, HR))) +
    geom_point(size = 3) +
    geom_errorbarh(aes(xmin = CI_lo, xmax = CI_hi), height = 0.2) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
    scale_x_log10() +
    labs(x = "Hazard Ratio (Rhythm Control vs Rate Control)",
         y = NULL,
         title = "Subgroup Analyses: Composite Outcome",
         subtitle = "HR < 1 favors rhythm control") +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 10))
} else {
  message("No subgroup analyses had sufficient data for plotting.")
}
```


## 14. Sensitivity Analyses

```{r}
#| label: exec-sensitivity

sensitivity_results <- run_sensitivity(ipw_results, cohort, confounders, config)
```

### E-value Results

```{r}
#| label: exec-evalue

message("=== E-value: Composite Outcome ===")
print(sensitivity_results$evalue_composite)

message("\n=== E-value: All-Cause Mortality ===")
print(sensitivity_results$evalue_death)
```

### PS Trimming Results

```{r}
#| label: exec-trimmed

if (!is.null(sensitivity_results$trimmed_fit)) {
  message("Trimmed composite result:")
  message(format_hr(sensitivity_results$trimmed_fit))
} else {
  message("PS trimming analysis was skipped (insufficient arms).")
}
```

### Post-Epic Era Results

```{r}
#| label: exec-epic

if (!is.null(sensitivity_results$epic_fit)) {
  message("Post-Epic (2021+) composite result:")
  message(format_hr(sensitivity_results$epic_fit))
} else {
  message("Post-Epic analysis was skipped (insufficient patients).")
}
```

### All-Cause Death Composite (if primary used CV death)

```{r}
#| label: exec-allcause-sensitivity

if (!is.null(sensitivity_results$allcause_composite_fit)) {
  message("All-cause death composite (sensitivity):")
  message(format_hr(sensitivity_results$allcause_composite_fit))
} else {
  message("Primary analysis already used all-cause death; no sensitivity needed.")
}
```


## 15. Summary

```{r}
#| label: exec-summary

use_cv_death <- attr(cohort, "use_cv_death")
death_label <- if (isTRUE(use_cv_death)) "CV death" else "all-cause death"

message("\n", strrep("=", 70))
message("PROTOCOL 02 SUMMARY: Early Rhythm vs Rate Control in Elderly AF")
message(strrep("=", 70))
message(sprintf("Analytic cohort: %d patients", nrow(cohort)))
message(sprintf("  Rhythm control: %d", sum(cohort$treatment == 1)))
message(sprintf("  Rate control:   %d", sum(cohort$treatment == 0)))
message(sprintf("Landmark: %d days (12 months) from AF diagnosis",
                config$classification_days))
message(sprintf("Follow-up: %d days (3 years) from landmark",
                config$followup_days))
message(sprintf("Study period: %s to %s", config$study_start, config$study_end))
message(sprintf("Death component in composite: %s", death_label))
message("")
message(sprintf("PRIMARY OUTCOME (Composite: %s + stroke + HF hosp):", death_label))
message(sprintf("  %s", format_hr(ipw_results$fits$composite)))
message("")
message("SECONDARY OUTCOMES:")
message(sprintf("  Stroke:          %s", format_hr(ipw_results$fits$stroke)))
message(sprintf("  HF hosp:         %s", format_hr(ipw_results$fits$hf_hosp)))
message(sprintf("  All-cause death: %s", format_hr(ipw_results$fits$death)))
message(sprintf("  AF hosp:         %s", format_hr(ipw_results$fits$af_hosp)))
message("")
message("Interpretation: HR < 1 favors rhythm control; HR > 1 favors rate control.")
message("This emulates EAST-AFNET 4 in octogenarians (>= 80), a population")
message("systematically underrepresented in that trial.")
```
