---
title: "CDW_DB_Profiler"
bibliography: references.bib
format: html
editor: visual
params:
  db_config: "databases/local/secure_pcornet_cdw.yaml"
  save_temp_tables: false
---

## Load R Libraries for DB access

Among other things, this provides access to dbply and dbconnect, both needed to connect to an MS SQL DB. If this fails, you probably need to install the packages in your version of R.

```{r load-libraries}
library(tidyverse)
library(DBI)
library(dbplyr)
library(pointblank)
library(SqlRender)
library(gtsummary)
library(yaml)
library(here)
options(encoding = "UTF-8") # Needed since winterm Locale defaults to 1252 instead of UTF-8

```

## Load configuration

Reads the database YAML config so output paths, DSN, and schema all come from
one source of truth. To profile a different offline DB, copy this .qmd and
change the `db_config` param at the top.

```{r load-config}
#| code-fold: true
# Shared scaffolding: with_db_connection, load_profiler_config,
# schema_name_from_cfg, make_suppress, make_safe_query, preflight_check,
# done_banner. Same helpers used by every profilers/<cdm>.qmd.
source(here::here("profilers/_helpers.R"))

# `params` is only defined during `quarto render`. Fall back to the default
# for interactive "Run All Chunks" in RStudio.
db_config_path <- if (exists("params") && !is.null(params$db_config)) {
  params$db_config
} else {
  "databases/local/secure_pcornet_cdw.yaml"
}

config <- load_profiler_config(db_config_path)
cfg     <- config$cfg
paths   <- config$paths
cdw_schema_name <- schema_name_from_cfg(cfg)
```

## Preflight connectivity check

Fail fast with a readable message if the DSN is wrong, before kicking off the
hours-long profile queries.

```{r preflight}
#| code-fold: true
preflight_check(cfg$connection$r_code, label = cfg$id %||% "cdw")
```

This function:

1.  **Extracts comprehensive schema information:**

    -   All tables in the specified schema
    -   Column names, data types, and nullability
    -   Primary keys (marked with \[PRIMARY KEY\])
    -   Foreign key relationships (marked with \[FK -\> table.column\])

2.  **Optional sample data** to help understand content and values

3.  **Machine-readable format** that's easy for Claude to parse when writing SQL

4.  **Can write to a file** for easy uploading to Claude

The output format looks like:

## Table: DEMOGRAPHIC

### Columns:

-   PATID: varchar(50) (NOT NULL) \[PRIMARY KEY\]
-   BIRTH_DATE: date (NULL)
-   SEX: varchar(2) (NULL) ...

## Table: ENCOUNTER

### Columns:

-   ENCOUNTERID: varchar(50) (NOT NULL) \[PRIMARY KEY\]
-   PATID: varchar(50) (NOT NULL) \[FK -\> DEMOGRAPHIC.PATID\] ...

```{r fn-extract-schema}
#| code-fold: true
extract_schema_for_claude <- function(con, 
                                      schema_name = "dbo",
                                      include_sample_values = FALSE,
                                      sample_size = 5,
                                      output_file = NULL) {
  
  library(DBI)
  library(dplyr)
  
  # Get list of tables AND views. Workers need both: duplicate-records
  # studies, for example, rely on curated views (e.g.
  # MasterPatientIndex.dbo.vSourceRecordsUID) that don't appear in
  # INFORMATION_SCHEMA.TABLES when TABLE_TYPE is restricted to 'BASE TABLE'.
  tables <- dbGetQuery(con, sprintf("
    SELECT TABLE_NAME,
           TABLE_TYPE
    FROM INFORMATION_SCHEMA.TABLES
    WHERE TABLE_SCHEMA = '%s'
      AND TABLE_TYPE IN ('BASE TABLE', 'VIEW')
    ORDER BY TABLE_TYPE, TABLE_NAME
  ", schema_name))
  
  schema_text <- paste0(
    "# Database Schema Information\n",
    "# Generated: ", Sys.time(), "\n",
    "# Schema: ", schema_name, "\n\n"
  )
  
  for (i_row in seq_len(nrow(tables))) {
    table_name <- tables$TABLE_NAME[i_row]
    table_type <- tables$TABLE_TYPE[i_row]
    header_kind <- if (table_type == "VIEW") "View" else "Table"

    # Get column information
    columns <- dbGetQuery(con, sprintf("
      SELECT
        COLUMN_NAME,
        DATA_TYPE,
        CHARACTER_MAXIMUM_LENGTH,
        IS_NULLABLE,
        COLUMN_DEFAULT
      FROM INFORMATION_SCHEMA.COLUMNS
      WHERE TABLE_SCHEMA = '%s'
        AND TABLE_NAME = '%s'
      ORDER BY ORDINAL_POSITION
    ", schema_name, table_name))
    
    # Get primary keys
    pks <- dbGetQuery(con, sprintf("
      SELECT COLUMN_NAME
      FROM INFORMATION_SCHEMA.KEY_COLUMN_USAGE
      WHERE TABLE_SCHEMA = '%s'
        AND TABLE_NAME = '%s'
        AND CONSTRAINT_NAME LIKE 'PK_%%'
    ", schema_name, table_name))
    
    # Get foreign keys
    fks <- dbGetQuery(con, sprintf("
      SELECT 
        fk.COLUMN_NAME,
        pk.TABLE_NAME AS REFERENCES_TABLE,
        pk.COLUMN_NAME AS REFERENCES_COLUMN
      FROM INFORMATION_SCHEMA.REFERENTIAL_CONSTRAINTS rc
      JOIN INFORMATION_SCHEMA.KEY_COLUMN_USAGE fk 
        ON rc.CONSTRAINT_NAME = fk.CONSTRAINT_NAME
      JOIN INFORMATION_SCHEMA.KEY_COLUMN_USAGE pk 
        ON rc.UNIQUE_CONSTRAINT_NAME = pk.CONSTRAINT_NAME
      WHERE fk.TABLE_SCHEMA = '%s'
        AND fk.TABLE_NAME = '%s'
    ", schema_name, table_name))
    
    # Build table / view section.
    schema_text <- paste0(schema_text, "## ", header_kind, ": ",
                          table_name, "\n\n")
    
    # Add columns
    schema_text <- paste0(schema_text, "### Columns:\n")
    for (i in 1:nrow(columns)) {
      col <- columns[i, ]
      type_info <- col$DATA_TYPE
      if (!is.na(col$CHARACTER_MAXIMUM_LENGTH)) {
        type_info <- paste0(type_info, "(", col$CHARACTER_MAXIMUM_LENGTH, ")")
      }
      
      nullable <- ifelse(col$IS_NULLABLE == "YES", "NULL", "NOT NULL")
      
      pk_marker <- ifelse(col$COLUMN_NAME %in% pks$COLUMN_NAME, " [PRIMARY KEY]", "")
      
      fk_info <- ""
      if (col$COLUMN_NAME %in% fks$COLUMN_NAME) {
        fk_row <- fks[fks$COLUMN_NAME == col$COLUMN_NAME, ]
        fk_info <- paste0(" [FK -> ", fk_row$REFERENCES_TABLE, ".", fk_row$REFERENCES_COLUMN, "]")
      }
      
      schema_text <- paste0(
        schema_text,
        "- ", col$COLUMN_NAME, ": ", type_info, " (", nullable, ")", 
        pk_marker, fk_info, "\n"
      )
    }
    
    # Add sample values if requested
    if (include_sample_values) {
      tryCatch({
        sample_data <- dbGetQuery(con, sprintf(
          "SELECT TOP %d * FROM %s.%s",
          sample_size, schema_name, table_name
        ))
        
        if (nrow(sample_data) > 0) {
          schema_text <- paste0(schema_text, "\n### Sample Data:\n")
          schema_text <- paste0(schema_text, "```\n")
          schema_text <- paste0(schema_text, capture.output(print(sample_data)), collapse = "\n")
          schema_text <- paste0(schema_text, "\n```\n")
        }
      }, error = function(e) {
        schema_text <<- paste0(schema_text, "\n### Sample Data: [Error retrieving samples]\n")
      })
    }
    
    schema_text <- paste0(schema_text, "\n")
  }
  
  # Add summary statistics
  schema_text <- paste0(
    schema_text,
    "\n---\n",
    "## Summary\n",
    "Total tables: ", nrow(tables), "\n"
  )
  
  # Write to file if requested
  if (!is.null(output_file)) {
    writeLines(schema_text, output_file)
    message("Schema written to: ", output_file)
  }
  
  return(schema_text)
}

# Usage examples:
# Basic usage
# schema_info <- extract_schema_for_claude(con)
# cat(schema_info)

# With sample data
# schema_info <- extract_schema_for_claude(con, include_sample_values = TRUE, sample_size = 3)

# Write to file
# extract_schema_for_claude(con, output_file = "database_schema.txt")

# For a different schema
# extract_schema_for_claude(con, schema_name = "myschema", output_file = "schema.txt")
```

The actual schema dump is invoked from the unified run section at the end of
this document so it shares the connection-lifecycle helper.

## CDW Data Profile for AutoTTE

The schema dump tells AutoTTE what columns exist but not what's actually *in*
them. This profiler produces a machine-readable data profile with aggregate
counts, coding system distributions, temporal coverage, and column completeness
— everything the worker agent needs for realistic feasibility assessment and
correct SQL generation, with no PHI.

**PHI protection rules applied throughout:**

- All outputs are aggregates — no individual rows or PATIDs
- Small cell suppression: any count of distinct patients \< 11 is replaced with `<11`
- Date ranges reported as year-month only
- No free-text or high-cardinality ID values exported

```{r fn-suppress-small-cells}
#| code-fold: true

# Small cell suppression helper — replaces counts < 11 with "<11"
# Applied to all patient counts before they are written to the profile.
suppress <- function(n, threshold = 11) {
  ifelse(is.na(n), NA_character_,
         ifelse(n < threshold, "<11", format(n, big.mark = ",")))
}
```

```{r fn-profile-cdw}
#| code-fold: true

profile_cdw_for_autoTTE <- function(con,
                                    schema_name = "dbo",
                                    output_file = "CDW_data_profile.md",
                                    suppress_threshold = 11) {

  # ------------------------------------------------------------------
  # Helper: run a query and return a data frame (or empty on error)
  # ------------------------------------------------------------------
  safe_query <- function(sql, label = "") {
    tryCatch(
      dbGetQuery(con, sql),
      error = function(e) {
        message("  [SKIP] ", label, ": ", conditionMessage(e))
        data.frame()
      }
    )
  }

  # Helper: suppress small counts in a character vector
  sup <- function(n) {
    ifelse(is.na(n), NA_character_,
           ifelse(n < suppress_threshold, "<11", format(n, big.mark = ",")))
  }

  out <- character()
  ln  <- function(...) out <<- c(out, paste0(...))

  ln("# CDW Data Profile for AutoTTE")
  ln("# Generated: ", Sys.time())
  ln("# Schema: ", schema_name)
  ln("# NOTE: This file contains ONLY aggregate counts. No PHI (PATIDs, dates")
  ln("#       of service, etc.) is included. Counts < ", suppress_threshold,
     " are suppressed.")
  ln("")

  # ================================================================
  # SECTION 1: Overall CDW size and temporal coverage
  # ================================================================
  ln("---")
  ln("## 1. Overall CDW Size and Temporal Coverage")
  ln("")

  # Total patients
  total_pts <- safe_query(sprintf(
    "SELECT COUNT(DISTINCT PATID) AS n FROM %s.DEMOGRAPHIC", schema_name),
    "DEMOGRAPHIC count")
  if (nrow(total_pts) > 0) {
    ln("Total distinct patients in DEMOGRAPHIC: ", sup(total_pts$n[1]))
    ln("")
  }

  # Per-table patient counts and date ranges
  # Define tables with their date column and patient ID column
  table_dates <- list(
    list(tbl = "ENCOUNTER",    date_col = "ADMIT_DATE"),
    list(tbl = "DIAGNOSIS",    date_col = "ADMIT_DATE"),
    list(tbl = "PROCEDURES",   date_col = "PX_DATE"),
    list(tbl = "PRESCRIBING",  date_col = "RX_ORDER_DATE"),
    list(tbl = "DISPENSING",   date_col = "DISPENSE_DATE"),
    list(tbl = "LAB_RESULT_CM", date_col = "RESULT_DATE"),
    list(tbl = "VITAL",        date_col = "MEASURE_DATE"),
    list(tbl = "MED_ADMIN",    date_col = "MEDADMIN_START_DATE"),
    list(tbl = "CONDITION",    date_col = "ONSET_DATE"),
    list(tbl = "DEATH",        date_col = "DEATH_DATE"),
    list(tbl = "ENROLLMENT",   date_col = "ENR_START_DATE")
  )

  ln("| Table | Distinct Patients | Rows | Earliest (YYYY-MM) | Latest (YYYY-MM) |")
  ln("|-------|-------------------|------|---------------------|-------------------|")

  for (td in table_dates) {
    res <- safe_query(sprintf("
      SELECT COUNT(DISTINCT PATID) AS n_patients,
             COUNT(*)              AS n_rows,
             FORMAT(MIN(%s), 'yyyy-MM') AS earliest,
             FORMAT(MAX(%s), 'yyyy-MM') AS latest
      FROM %s.%s
    ", td$date_col, td$date_col, schema_name, td$tbl), td$tbl)

    if (nrow(res) > 0 && !is.na(res$n_patients[1]) && res$n_patients[1] > 0) {
      ln("| ", td$tbl, " | ", sup(res$n_patients[1]), " | ",
         format(res$n_rows[1], big.mark = ","), " | ",
         ifelse(is.na(res$earliest[1]), "—", res$earliest[1]), " | ",
         ifelse(is.na(res$latest[1]), "—", res$latest[1]), " |")
    }
  }
  ln("")

  # ================================================================
  # SECTION 2: Year-by-year data volume (ENCOUNTER table)
  # ================================================================
  ln("---")
  ln("## 2. Year-by-Year Patient Volume (ENCOUNTER)")
  ln("")
  ln("This shows how many distinct patients have at least one encounter per year.")
  ln("")

  yearly <- safe_query(sprintf("
    SELECT YEAR(ADMIT_DATE)       AS yr,
           COUNT(DISTINCT PATID)  AS n_patients,
           COUNT(*)               AS n_encounters
    FROM %s.ENCOUNTER
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY YEAR(ADMIT_DATE)
    ORDER BY yr
  ", schema_name), "yearly encounters")

  if (nrow(yearly) > 0) {
    ln("| Year | Distinct Patients | Encounters |")
    ln("|------|-------------------|------------|")
    for (i in seq_len(nrow(yearly))) {
      ln("| ", yearly$yr[i], " | ", sup(yearly$n_patients[i]), " | ",
         format(yearly$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 3: Legacy Encounters, EHR Migration, and Record Duplication
  # ================================================================
  ln("---")
  ln("## 3. Legacy Encounters, EHR Migration, and Record Duplication")
  ln("")
  ln("### Background: AllScripts → Epic Migration")
  ln("")
  ln("This CDW contains data from TWO EHR eras:")
  ln("1. **AllScripts era** — original EHR data fed directly into the CDW.")
  ln("2. **Epic era** — after migration, Epic became the primary EHR feed.")
  ln("")
  ln("During the Epic go-live, some AllScripts records were imported INTO Epic.")
  ln("The Epic feed then sent those imported records back into the CDW, creating")
  ln("**duplicate records**: the original AllScripts version AND the re-imported")
  ln("Epic version of the same clinical event.")
  ln("")
  ln("These duplicates are identifiable via two columns in the ENCOUNTER table:")
  ln("- `RAW_ENC_TYPE = 'Legacy Encounter'` — the Epic-side label for imported")
  ln("  AllScripts encounters that were re-fed into the CDW.")
  ln("- `IS_LEGACY_IMPORT = 'Y'` — a CDW-level flag for the same records.")
  ln("")
  ln("**Default rule: ALWAYS filter out legacy encounters** unless you have a")
  ln("specific reason to keep them. Including them causes double-counting of")
  ln("encounters, diagnoses, procedures, and other linked records from the")
  ln("pre-Epic period.")
  ln("")
  ln("```sql")
  ln("-- Standard filter to exclude duplicate legacy encounters:")
  ln("INNER JOIN CDW.dbo.ENCOUNTER e ON ...")
  ln("  AND e.RAW_ENC_TYPE <> 'Legacy Encounter'")
  ln("```")
  ln("")

  # 3a. RAW_ENC_TYPE distribution (primary filter)
  ln("### 3a. RAW_ENC_TYPE Distribution")
  ln("")
  ln("This is the primary column for identifying legacy duplicates.")
  ln("")

  raw_enc_dist <- safe_query(sprintf("
    SELECT ISNULL(RAW_ENC_TYPE, '(NULL)') AS raw_enc_type,
           COUNT(DISTINCT PATID)  AS n_patients,
           COUNT(*)               AS n_encounters
    FROM %s.ENCOUNTER
    GROUP BY RAW_ENC_TYPE
    ORDER BY n_encounters DESC
  ", schema_name), "RAW_ENC_TYPE dist")

  if (nrow(raw_enc_dist) > 0) {
    ln("| RAW_ENC_TYPE | Patients | Encounters |")
    ln("|--------------|----------|------------|")
    for (i in seq_len(nrow(raw_enc_dist))) {
      ln("| ", raw_enc_dist$raw_enc_type[i], " | ",
         sup(raw_enc_dist$n_patients[i]), " | ",
         format(raw_enc_dist$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # 3b. IS_LEGACY_IMPORT distribution (secondary/corroborating flag)
  has_legacy_col <- safe_query(sprintf("
    SELECT TOP 1 COLUMN_NAME
    FROM INFORMATION_SCHEMA.COLUMNS
    WHERE TABLE_SCHEMA = '%s' AND TABLE_NAME = 'ENCOUNTER'
      AND COLUMN_NAME = 'IS_LEGACY_IMPORT'
  ", schema_name), "IS_LEGACY_IMPORT check")

  if (nrow(has_legacy_col) > 0) {
    legacy_dist <- safe_query(sprintf("
      SELECT ISNULL(IS_LEGACY_IMPORT, '(NULL)') AS legacy_flag,
             COUNT(DISTINCT PATID)  AS n_patients,
             COUNT(*)               AS n_encounters
      FROM %s.ENCOUNTER
      GROUP BY IS_LEGACY_IMPORT
      ORDER BY legacy_flag
    ", schema_name), "IS_LEGACY_IMPORT dist")

    if (nrow(legacy_dist) > 0) {
      ln("### 3b. IS_LEGACY_IMPORT Distribution")
      ln("")
      ln("| IS_LEGACY_IMPORT | Patients | Encounters |")
      ln("|------------------|----------|------------|")
      for (i in seq_len(nrow(legacy_dist))) {
        ln("| ", legacy_dist$legacy_flag[i], " | ",
           sup(legacy_dist$n_patients[i]), " | ",
           format(legacy_dist$n_encounters[i], big.mark = ","), " |")
      }
      ln("")
    }

    # Cross-tab RAW_ENC_TYPE vs IS_LEGACY_IMPORT to verify alignment
    cross_tab <- safe_query(sprintf("
      SELECT ISNULL(RAW_ENC_TYPE, '(NULL)') AS raw_enc_type,
             ISNULL(IS_LEGACY_IMPORT, '(NULL)') AS legacy_flag,
             COUNT(*) AS n_encounters
      FROM %s.ENCOUNTER
      WHERE RAW_ENC_TYPE = 'Legacy Encounter'
         OR IS_LEGACY_IMPORT = 'Y'
      GROUP BY RAW_ENC_TYPE, IS_LEGACY_IMPORT
      ORDER BY raw_enc_type, legacy_flag
    ", schema_name), "cross-tab legacy flags")

    if (nrow(cross_tab) > 0) {
      ln("### Cross-tab: RAW_ENC_TYPE vs IS_LEGACY_IMPORT")
      ln("")
      ln("Verifies whether the two legacy flags align.")
      ln("")
      ln("| RAW_ENC_TYPE | IS_LEGACY_IMPORT | Encounters |")
      ln("|--------------|------------------|------------|")
      for (i in seq_len(nrow(cross_tab))) {
        ln("| ", cross_tab$raw_enc_type[i], " | ",
           cross_tab$legacy_flag[i], " | ",
           format(cross_tab$n_encounters[i], big.mark = ","), " |")
      }
      ln("")
    }
  }

  # 3c. CDW_Source distribution on ENCOUNTER
  ln("### 3c. CDW_Source Distribution (ENCOUNTER)")
  ln("")
  ln("Shows which source system feeds each encounter. Multiple CDW_Source")
  ln("values in the same year range may indicate overlapping feeds.")
  ln("")

  cdw_src <- safe_query(sprintf("
    SELECT ISNULL(CDW_Source, '(NULL)') AS cdw_source,
           COUNT(DISTINCT PATID)  AS n_patients,
           COUNT(*)               AS n_encounters,
           FORMAT(MIN(ADMIT_DATE), 'yyyy-MM') AS earliest,
           FORMAT(MAX(ADMIT_DATE), 'yyyy-MM') AS latest
    FROM %s.ENCOUNTER
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY CDW_Source
    ORDER BY n_encounters DESC
  ", schema_name), "CDW_Source dist")

  if (nrow(cdw_src) > 0) {
    ln("| CDW_Source | Patients | Encounters | Earliest | Latest |")
    ln("|-----------|----------|------------|----------|--------|")
    for (i in seq_len(nrow(cdw_src))) {
      ln("| ", cdw_src$cdw_source[i], " | ",
         sup(cdw_src$n_patients[i]), " | ",
         format(cdw_src$n_encounters[i], big.mark = ","), " | ",
         ifelse(is.na(cdw_src$earliest[i]), "—", cdw_src$earliest[i]), " | ",
         ifelse(is.na(cdw_src$latest[i]), "—", cdw_src$latest[i]), " |")
    }
    ln("")
  }

  # 3d. Legacy vs non-legacy by year
  legacy_by_year <- safe_query(sprintf("
    SELECT YEAR(ADMIT_DATE) AS yr,
           CASE WHEN RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
                ELSE 'Non-Legacy' END AS legacy_flag,
           COUNT(DISTINCT PATID)  AS n_patients,
           COUNT(*)               AS n_encounters
    FROM %s.ENCOUNTER
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY YEAR(ADMIT_DATE),
             CASE WHEN RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
                  ELSE 'Non-Legacy' END
    ORDER BY yr, legacy_flag
  ", schema_name), "legacy by year")

  if (nrow(legacy_by_year) > 0) {
    ln("### 3d. Legacy vs Non-Legacy Encounters by Year")
    ln("")
    ln("Shows when the duplicate legacy records appear and when they stop.")
    ln("Years with BOTH Legacy and Non-Legacy rows indicate the overlap")
    ln("period where double-counting occurs if legacy records are not filtered.")
    ln("")
    ln("| Year | Category | Patients | Encounters |")
    ln("|------|----------|----------|------------|")
    for (i in seq_len(nrow(legacy_by_year))) {
      ln("| ", legacy_by_year$yr[i], " | ",
         legacy_by_year$legacy_flag[i], " | ",
         sup(legacy_by_year$n_patients[i]), " | ",
         format(legacy_by_year$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # 3e. Data completeness: legacy vs non-legacy
  ln("### 3e. Data Completeness: Legacy vs Non-Legacy")
  ln("")
  ln("Percentage of encounters with non-NULL values for key fields.")
  ln("Differences may indicate what data was carried over in the migration")
  ln("vs what was left behind.")
  ln("")

  completeness_cols <- c("DISCHARGE_DATE", "DISCHARGE_DISPOSITION",
                         "DISCHARGE_STATUS", "PAYER_TYPE_PRIMARY",
                         "FACILITY_LOCATION")
  ln("| Column | Legacy % Non-NULL | Non-Legacy % Non-NULL |")
  ln("|--------|-------------------|----------------------|")

  for (cc in completeness_cols) {
    comp_res <- safe_query(sprintf("
      SELECT
        CASE WHEN RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
             ELSE 'Non-Legacy' END AS legacy_flag,
        COUNT(*) AS total,
        SUM(CASE WHEN %s IS NOT NULL THEN 1 ELSE 0 END) AS n_nonnull
      FROM %s.ENCOUNTER
      GROUP BY CASE WHEN RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
                    ELSE 'Non-Legacy' END
    ", cc, schema_name), paste0("completeness ", cc))

    if (nrow(comp_res) >= 1) {
      leg_row <- comp_res[comp_res$legacy_flag == "Legacy", ]
      non_row <- comp_res[comp_res$legacy_flag == "Non-Legacy", ]
      leg_pct <- if (nrow(leg_row) > 0 && leg_row$total[1] > 0)
        round(100 * leg_row$n_nonnull[1] / leg_row$total[1], 1) else "—"
      non_pct <- if (nrow(non_row) > 0 && non_row$total[1] > 0)
        round(100 * non_row$n_nonnull[1] / non_row$total[1], 1) else "—"
      ln("| ", cc, " | ", leg_pct, "%% | ", non_pct, "%% |")
    }
  }
  ln("")

  # 3f. Linked tables: do legacy encounters have associated records?
  ln("### 3f. Linked Data Availability for Legacy Encounters")
  ln("")
  ln("Do legacy encounters have associated records in other tables?")
  ln("If legacy encounters DO have linked diagnoses/procedures, that confirms")
  ln("they are duplicates of the original AllScripts records (not empty shells).")
  ln("")

  linked_tables <- list(
    list(tbl = "DIAGNOSIS",    join_col = "ENCOUNTERID"),
    list(tbl = "PROCEDURES",   join_col = "ENCOUNTERID"),
    list(tbl = "PRESCRIBING",  join_col = "ENCOUNTERID"),
    list(tbl = "LAB_RESULT_CM", join_col = "ENCOUNTERID"),
    list(tbl = "VITAL",        join_col = "ENCOUNTERID")
  )

  ln("| Linked Table | Legacy Enc w/ Records | Non-Legacy Enc w/ Records |")
  ln("|--------------|----------------------|--------------------------|")

  for (lt in linked_tables) {
    link_res <- safe_query(sprintf("
      SELECT
        CASE WHEN e.RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
             ELSE 'Non-Legacy' END AS legacy_flag,
        COUNT(DISTINCT e.ENCOUNTERID) AS total_enc,
        COUNT(DISTINCT t.ENCOUNTERID) AS linked_enc
      FROM %s.ENCOUNTER e
      LEFT JOIN %s.%s t ON e.ENCOUNTERID = t.ENCOUNTERID
      GROUP BY CASE WHEN e.RAW_ENC_TYPE = 'Legacy Encounter' THEN 'Legacy'
                    ELSE 'Non-Legacy' END
    ", schema_name, schema_name, lt$tbl), paste0("linked ", lt$tbl))

    if (nrow(link_res) >= 1) {
      leg_row <- link_res[link_res$legacy_flag == "Legacy", ]
      non_row <- link_res[link_res$legacy_flag == "Non-Legacy", ]
      leg_pct <- if (nrow(leg_row) > 0 && leg_row$total_enc[1] > 0)
        paste0(round(100 * leg_row$linked_enc[1] / leg_row$total_enc[1], 1), "%%")
      else "—"
      non_pct <- if (nrow(non_row) > 0 && non_row$total_enc[1] > 0)
        paste0(round(100 * non_row$linked_enc[1] / non_row$total_enc[1], 1), "%%")
      else "—"
      ln("| ", lt$tbl, " | ", leg_pct, " | ", non_pct, " |")
    }
  }
  ln("")

  # 3g. EHR source system identification
  ln("### 3g. Patient EHR Source Coverage")
  ln("")
  ln("Patients with ALLSCRIPTS_PERSON_ID vs EPIC_PAT_ID in DEMOGRAPHIC,")
  ln("showing how many patients come from each source system. Patients with")
  ln("BOTH IDs are those whose AllScripts records were linked to Epic.")
  ln("")

  ehr_src <- safe_query(sprintf("
    SELECT
      SUM(CASE WHEN ALLSCRIPTS_PERSON_ID IS NOT NULL THEN 1 ELSE 0 END) AS n_allscripts,
      SUM(CASE WHEN EPIC_PAT_ID IS NOT NULL THEN 1 ELSE 0 END) AS n_epic,
      SUM(CASE WHEN ALLSCRIPTS_PERSON_ID IS NOT NULL
                AND EPIC_PAT_ID IS NOT NULL THEN 1 ELSE 0 END) AS n_both,
      COUNT(*) AS n_total
    FROM %s.DEMOGRAPHIC
  ", schema_name), "EHR source IDs")

  if (nrow(ehr_src) > 0 && !all(is.na(ehr_src))) {
    ln("| Category | Patients |")
    ln("|----------|----------|")
    ln("| Total in DEMOGRAPHIC | ", sup(ehr_src$n_total[1]), " |")
    ln("| Has ALLSCRIPTS_PERSON_ID | ", sup(ehr_src$n_allscripts[1]), " |")
    ln("| Has EPIC_PAT_ID | ", sup(ehr_src$n_epic[1]), " |")
    ln("| Has BOTH (linked across systems) | ", sup(ehr_src$n_both[1]), " |")
    ln("")
  }

  # ================================================================
  # SECTION 4: Diagnosis coding systems by year (ICD-9 vs ICD-10)
  # ================================================================
  ln("---")
  ln("## 4. Diagnosis Coding Systems by Year (ICD-9 vs ICD-10)")
  ln("")
  ln("DX_TYPE values: '09' = ICD-9-CM, '10' = ICD-10-CM, 'SM' = SNOMED, etc.")
  ln("The US transitioned from ICD-9 to ICD-10 on 2015-10-01. Knowing which")
  ln("years have which coding system is essential for writing correct SQL.")
  ln("")

  dx_by_year <- safe_query(sprintf("
    SELECT YEAR(ADMIT_DATE) AS yr,
           DX_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rows
    FROM %s.DIAGNOSIS
    WHERE ADMIT_DATE IS NOT NULL
    GROUP BY YEAR(ADMIT_DATE), DX_TYPE
    ORDER BY yr, DX_TYPE
  ", schema_name), "DX_TYPE by year")

  if (nrow(dx_by_year) > 0) {
    ln("| Year | DX_TYPE | Distinct Patients | Rows |")
    ln("|------|---------|-------------------|------|")
    for (i in seq_len(nrow(dx_by_year))) {
      ln("| ", dx_by_year$yr[i], " | ", dx_by_year$DX_TYPE[i], " | ",
         sup(dx_by_year$n_patients[i]), " | ",
         format(dx_by_year$n_rows[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 5: Procedure coding systems
  # ================================================================
  ln("---")
  ln("## 5. Procedure Coding Systems")
  ln("")

  px_types <- safe_query(sprintf("
    SELECT PX_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rows,
           FORMAT(MIN(PX_DATE), 'yyyy-MM') AS earliest,
           FORMAT(MAX(PX_DATE), 'yyyy-MM') AS latest
    FROM %s.PROCEDURES
    WHERE PX_DATE IS NOT NULL
    GROUP BY PX_TYPE
    ORDER BY n_patients DESC
  ", schema_name), "PX_TYPE")

  if (nrow(px_types) > 0) {
    ln("PX_TYPE values: 'CH' = CPT/HCPCS, '10' = ICD-10-PCS, '09' = ICD-9-CM, 'RE' = Revenue, etc.")
    ln("")
    ln("| PX_TYPE | Distinct Patients | Rows | Earliest | Latest |")
    ln("|---------|-------------------|------|----------|--------|")
    for (i in seq_len(nrow(px_types))) {
      ln("| ", px_types$PX_TYPE[i], " | ", sup(px_types$n_patients[i]), " | ",
         format(px_types$n_rows[i], big.mark = ","), " | ",
         px_types$earliest[i], " | ", px_types$latest[i], " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 6: Lab LOINC coverage (top 50 most common)
  # ================================================================
  ln("---")
  ln("## 6. Lab Results — Top 50 LOINCs by Patient Count")
  ln("")

  top_labs <- safe_query(sprintf("
    SELECT TOP 50
           LAB_LOINC,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_results,
           AVG(TRY_CAST(RESULT_NUM AS FLOAT)) AS mean_val,
           MIN(TRY_CAST(RESULT_NUM AS FLOAT)) AS min_val,
           MAX(TRY_CAST(RESULT_NUM AS FLOAT)) AS max_val
    FROM %s.LAB_RESULT_CM
    WHERE LAB_LOINC IS NOT NULL
    GROUP BY LAB_LOINC
    ORDER BY n_patients DESC
  ", schema_name), "top labs")

  if (nrow(top_labs) > 0) {
    ln("| LAB_LOINC | Patients | Results | Mean | Min | Max |")
    ln("|-----------|----------|---------|------|-----|-----|")
    for (i in seq_len(nrow(top_labs))) {
      r <- top_labs[i, ]
      ln("| ", r$LAB_LOINC, " | ", sup(r$n_patients), " | ",
         format(r$n_results, big.mark = ","), " | ",
         ifelse(is.na(r$mean_val), "—", round(r$mean_val, 2)), " | ",
         ifelse(is.na(r$min_val), "—", round(r$min_val, 2)), " | ",
         ifelse(is.na(r$max_val), "—", round(r$max_val, 2)), " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 7: Prescribing — Top 50 RXNORM_CUI by patient count
  # ================================================================
  ln("---")
  ln("## 7. Prescribing — Top 50 Medications (RXNORM_CUI) by Patient Count")
  ln("")

  top_rx <- safe_query(sprintf("
    SELECT TOP 50
           RXNORM_CUI,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_prescriptions,
           FORMAT(MIN(RX_ORDER_DATE), 'yyyy-MM') AS earliest,
           FORMAT(MAX(RX_ORDER_DATE), 'yyyy-MM') AS latest
    FROM %s.PRESCRIBING
    WHERE RXNORM_CUI IS NOT NULL AND RXNORM_CUI != ''
    GROUP BY RXNORM_CUI
    ORDER BY n_patients DESC
  ", schema_name), "top Rx")

  if (nrow(top_rx) > 0) {
    ln("| RXNORM_CUI | Patients | Prescriptions | Earliest | Latest |")
    ln("|------------|----------|---------------|----------|--------|")
    for (i in seq_len(nrow(top_rx))) {
      r <- top_rx[i, ]
      ln("| ", r$RXNORM_CUI, " | ", sup(r$n_patients), " | ",
         format(r$n_prescriptions, big.mark = ","), " | ",
         r$earliest, " | ", r$latest, " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 8: Demographic distributions
  # ================================================================
  ln("---")
  ln("## 8. Demographic Distributions")
  ln("")

  for (col in c("SEX", "RACE", "HISPANIC")) {
    dist <- safe_query(sprintf("
      SELECT %s AS val, COUNT(DISTINCT PATID) AS n_patients
      FROM %s.DEMOGRAPHIC
      GROUP BY %s
      ORDER BY n_patients DESC
    ", col, schema_name, col), paste0("DEMOGRAPHIC.", col))

    if (nrow(dist) > 0) {
      ln("### ", col)
      ln("")
      ln("| Value | Patients |")
      ln("|-------|----------|")
      for (i in seq_len(nrow(dist))) {
        ln("| ", ifelse(is.na(dist$val[i]), "(NULL)", dist$val[i]),
           " | ", sup(dist$n_patients[i]), " |")
      }
      ln("")
    }
  }

  # ================================================================
  # SECTION 9: Encounter type distribution
  # ================================================================
  ln("---")
  ln("## 9. Encounter Types")
  ln("")

  enc_types <- safe_query(sprintf("
    SELECT ENC_TYPE,
           COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_encounters
    FROM %s.ENCOUNTER
    GROUP BY ENC_TYPE
    ORDER BY n_patients DESC
  ", schema_name), "ENC_TYPE")

  if (nrow(enc_types) > 0) {
    ln("PCORnet ENC_TYPE: AV=Ambulatory, ED=Emergency, EI=ED-to-Inpatient,")
    ln("IP=Inpatient, IS=Institutional, OS=Outpt Surgery, OA=Other Ambulatory,")
    ln("TH=Telehealth, NI=No info, UN=Unknown, OT=Other")
    ln("")
    ln("| ENC_TYPE | Patients | Encounters |")
    ln("|----------|----------|------------|")
    for (i in seq_len(nrow(enc_types))) {
      ln("| ", ifelse(is.na(enc_types$ENC_TYPE[i]), "(NULL)", enc_types$ENC_TYPE[i]),
         " | ", sup(enc_types$n_patients[i]), " | ",
         format(enc_types$n_encounters[i], big.mark = ","), " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 10: Vital signs completeness
  # ================================================================
  ln("---")
  ln("## 10. Vital Signs Completeness")
  ln("")

  vital_cols <- c("HT", "WT", "ORIGINAL_BMI", "SYSTOLIC", "DIASTOLIC", "SMOKING")
  for (vc in vital_cols) {
    vc_res <- safe_query(sprintf("
      SELECT COUNT(DISTINCT PATID) AS n_patients,
             COUNT(*) AS n_rows
      FROM %s.VITAL
      WHERE %s IS NOT NULL
    ", schema_name, vc), paste0("VITAL.", vc))

    if (nrow(vc_res) > 0 && vc_res$n_patients[1] > 0) {
      ln("- **", vc, "**: ", sup(vc_res$n_patients[1]),
         " patients (", format(vc_res$n_rows[1], big.mark = ","), " measurements)")
    }
  }

  # Smoking value distribution
  smoking <- safe_query(sprintf("
    SELECT SMOKING, COUNT(DISTINCT PATID) AS n_patients
    FROM %s.VITAL
    WHERE SMOKING IS NOT NULL
    GROUP BY SMOKING
    ORDER BY n_patients DESC
  ", schema_name), "SMOKING values")

  if (nrow(smoking) > 0) {
    ln("")
    ln("### SMOKING Values")
    ln("")
    ln("| SMOKING | Patients |")
    ln("|---------|----------|")
    for (i in seq_len(nrow(smoking))) {
      ln("| ", smoking$SMOKING[i], " | ", sup(smoking$n_patients[i]), " |")
    }
  }
  ln("")

  # ================================================================
  # SECTION 11: DEATH table profile
  # ================================================================
  ln("---")
  ln("## 11. Death Records")
  ln("")

  death_res <- safe_query(sprintf("
    SELECT COUNT(DISTINCT PATID) AS n_patients,
           COUNT(*) AS n_rows
    FROM %s.DEATH
  ", schema_name), "DEATH count")

  if (nrow(death_res) > 0) {
    ln("Distinct patients with death records: ", sup(death_res$n_patients[1]))
    ln("Total rows (may have duplicates per patient): ",
       format(death_res$n_rows[1], big.mark = ","))
    ln("")
  }

  death_src <- safe_query(sprintf("
    SELECT DEATH_SOURCE, COUNT(DISTINCT PATID) AS n_patients
    FROM %s.DEATH
    WHERE DEATH_SOURCE IS NOT NULL
    GROUP BY DEATH_SOURCE
    ORDER BY n_patients DESC
  ", schema_name), "DEATH_SOURCE")

  if (nrow(death_src) > 0) {
    ln("### DEATH_SOURCE Values")
    ln("")
    ln("| DEATH_SOURCE | Patients |")
    ln("|--------------|----------|")
    for (i in seq_len(nrow(death_src))) {
      ln("| ", death_src$DEATH_SOURCE[i], " | ", sup(death_src$n_patients[i]), " |")
    }
    ln("")
  }

  # ---- DEATH_CAUSE sub-profile ----
  # Absence of this block made colonoscopy Protocol_01 learn at runtime that
  # DEATH_CAUSE_CODE='10' returns nothing. Report row/patient counts, coverage
  # vs DEATH, and the full code-system / source / type / confidence breakdowns
  # so downstream designs can gate cause-of-death endpoints at design time.
  dc_res <- safe_query(sprintf("
    SELECT COUNT(*) AS n_rows,
           COUNT(DISTINCT PATID) AS n_patients
    FROM %s.DEATH_CAUSE
  ", schema_name), "DEATH_CAUSE count")

  if (nrow(dc_res) > 0 && !is.na(dc_res$n_rows[1])) {
    ln("### DEATH_CAUSE Table")
    ln("")
    ln("Distinct patients with cause-of-death records: ", sup(dc_res$n_patients[1]))
    ln("Total rows: ", format(dc_res$n_rows[1], big.mark = ","))

    if (nrow(death_res) > 0 && death_res$n_patients[1] > 0) {
      cov <- dc_res$n_patients[1] / death_res$n_patients[1]
      ln("Coverage vs DEATH: ", sprintf("%.1f%%", 100 * cov),
         " of deceased patients have \u2265 1 DEATH_CAUSE row")
    }
    ln("")

    for (dim in list(
      list(col = "DEATH_CAUSE_CODE",       title = "DEATH_CAUSE_CODE (code system)"),
      list(col = "DEATH_CAUSE_SOURCE",     title = "DEATH_CAUSE_SOURCE"),
      list(col = "DEATH_CAUSE_TYPE",       title = "DEATH_CAUSE_TYPE"),
      list(col = "DEATH_CAUSE_CONFIDENCE", title = "DEATH_CAUSE_CONFIDENCE")
    )) {
      dist <- safe_query(sprintf("
        SELECT %s AS val,
               COUNT(*) AS n_rows,
               COUNT(DISTINCT PATID) AS n_patients
        FROM %s.DEATH_CAUSE
        GROUP BY %s
        ORDER BY n_rows DESC
      ", dim$col, schema_name, dim$col), dim$col)

      if (nrow(dist) > 0) {
        ln("#### ", dim$title)
        ln("")
        ln("| ", dim$col, " | Rows | Patients |")
        ln("|", strrep("-", nchar(dim$col) + 2), "|------|----------|")
        for (i in seq_len(nrow(dist))) {
          v <- dist$val[i]
          if (is.na(v) || v == "") v <- "NULL"
          ln("| ", v, " | ", sup(dist$n_rows[i]),
             " | ", sup(dist$n_patients[i]), " |")
        }
        ln("")
      }
    }

    # Storage-format read: dotted vs dotless. Mirrors the DX probe in
    # tools/diagnose_cdw_dx_format.qmd but for DEATH_CAUSE specifically,
    # whose varchar(8) storage is NOT covered by the DIAGNOSIS convention.
    fmt <- safe_query(sprintf("
      SELECT
        SUM(CASE WHEN CHARINDEX('.', DEATH_CAUSE) > 0 THEN 1 ELSE 0 END) AS n_dotted,
        SUM(CASE WHEN CHARINDEX('.', DEATH_CAUSE) = 0 THEN 1 ELSE 0 END) AS n_dotless,
        COUNT(*) AS n_total
      FROM %s.DEATH_CAUSE
      WHERE DEATH_CAUSE IS NOT NULL
    ", schema_name), "DEATH_CAUSE format")

    if (nrow(fmt) > 0 && !is.na(fmt$n_total[1]) && fmt$n_total[1] > 0) {
      ln("#### DEATH_CAUSE storage format (non-NULL rows)")
      ln("")
      ln("| Format | Rows |")
      ln("|--------|------|")
      ln("| Contains period (dotted) | ", sup(fmt$n_dotted[1]), " |")
      ln("| No period (dotless)      | ", sup(fmt$n_dotless[1]), " |")
      ln("")
    }
  }

  # ================================================================
  # SECTION 12: Column completeness (NULL rates) for key tables
  # ================================================================
  ln("---")
  ln("## 12. Column Completeness — Key Tables")
  ln("")
  ln("Percentage of rows where the column is NOT NULL.")
  ln("")

  completeness_tables <- list(
    list(tbl = "DEMOGRAPHIC", cols = c("BIRTH_DATE", "SEX", "RACE", "HISPANIC")),
    list(tbl = "PRESCRIBING", cols = c("RXNORM_CUI", "RX_ORDER_DATE",
                                        "RX_START_DATE", "RX_END_DATE",
                                        "RX_QUANTITY", "RX_DAYS_SUPPLY",
                                        "RX_DOSE_ORDERED")),
    list(tbl = "DIAGNOSIS",   cols = c("DX", "DX_TYPE", "ADMIT_DATE",
                                        "DX_DATE", "DX_SOURCE")),
    list(tbl = "LAB_RESULT_CM", cols = c("LAB_LOINC", "RESULT_NUM",
                                          "RESULT_QUAL", "RESULT_DATE",
                                          "NORM_RANGE_LOW", "NORM_RANGE_HIGH"))
  )

  for (ct in completeness_tables) {
    total_sql <- sprintf("SELECT COUNT(*) AS n FROM %s.%s", schema_name, ct$tbl)
    total_res <- safe_query(total_sql, paste0(ct$tbl, " total"))
    if (nrow(total_res) == 0 || total_res$n[1] == 0) next

    total_n <- total_res$n[1]
    ln("### ", ct$tbl, " (", format(total_n, big.mark = ","), " rows)")
    ln("")
    ln("| Column | % Non-NULL |")
    ln("|--------|-----------|")

    for (col in ct$cols) {
      col_res <- safe_query(sprintf("
        SELECT COUNT(*) AS n_nonnull
        FROM %s.%s
        WHERE %s IS NOT NULL
      ", schema_name, ct$tbl, col), paste0(ct$tbl, ".", col))

      if (nrow(col_res) > 0) {
        pct <- round(100 * col_res$n_nonnull[1] / total_n, 1)
        ln("| ", col, " | ", pct, "%% |")
      }
    }
    ln("")
  }

  # ================================================================
  # SECTION 13: Top 30 diagnosis codes (ICD-10 only, by patient count)
  # ================================================================
  ln("---")
  ln("## 13. Top 30 ICD-10 Diagnosis Codes by Patient Count")
  ln("")

  top_dx <- safe_query(sprintf("
    SELECT TOP 30
           DX,
           COUNT(DISTINCT PATID) AS n_patients
    FROM %s.DIAGNOSIS
    WHERE DX_TYPE = '10' AND DX IS NOT NULL
    GROUP BY DX
    ORDER BY n_patients DESC
  ", schema_name), "top ICD-10 DX")

  if (nrow(top_dx) > 0) {
    ln("| ICD-10 Code | Patients |")
    ln("|-------------|----------|")
    for (i in seq_len(nrow(top_dx))) {
      ln("| ", top_dx$DX[i], " | ", sup(top_dx$n_patients[i]), " |")
    }
    ln("")
  }

  # ================================================================
  # SECTION 14: Clinically important condition prevalence
  # ================================================================
  ln("---")
  ln("## 14. Prevalence of Clinically Important Conditions")
  ln("")
  ln("Patient counts for conditions commonly used in TTE study design.")
  ln("Counts use ICD-10 codes (DX_TYPE = '10'). Patients may appear in")
  ln("multiple categories.")
  ln("")

  conditions <- list(
    list(name = "Atrial fibrillation (I48.x)",  pattern = "I48%"),
    list(name = "Heart failure (I50.x)",         pattern = "I50%"),
    list(name = "Ischemic stroke (I63.x)",       pattern = "I63%"),
    list(name = "Type 2 diabetes (E11.x)",       pattern = "E11%"),
    list(name = "CKD stage 3 (N18.3x)",         pattern = "N18.3%"),
    list(name = "CKD stage 4 (N18.4)",          pattern = "N18.4%"),
    list(name = "CKD stage 5 (N18.5)",          pattern = "N18.5%"),
    list(name = "ESRD / dialysis (N18.6, Z99.2)", pattern = "N18.6%"),
    list(name = "Hypertension (I10-I16)",        pattern = "I1[0-6]%"),
    list(name = "COPD (J44.x)",                  pattern = "J44%"),
    list(name = "Sepsis (R65.2x, A41.x)",        pattern = "R65.2%"),
    list(name = "Dementia (F01-F03, G30)",        pattern = "F0[1-3]%"),
    list(name = "Liver cirrhosis (K74.x)",        pattern = "K74%"),
    list(name = "VTE / PE (I26.x, I82.x)",       pattern = "I26%"),
    list(name = "Major bleeding (various)",       pattern = "K92.0%"),
    list(name = "Obesity (E66.x)",               pattern = "E66%")
  )

  ln("| Condition | Patients |")
  ln("|-----------|----------|")

  for (cond in conditions) {
    # Use LIKE for simple patterns, handle multi-pattern conditions
    if (grepl(",", cond$name) || grepl("I1\\[", cond$pattern)) {
      # Multi-pattern: use separate queries for some conditions
      if (grepl("Sepsis", cond$name)) {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10'
            AND (DX LIKE 'R65.2%%' OR DX LIKE 'A41%%')
        ", schema_name), cond$name)
      } else if (grepl("ESRD", cond$name)) {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10'
            AND (DX LIKE 'N18.6%%' OR DX LIKE 'Z99.2%%')
        ", schema_name), cond$name)
      } else if (grepl("Dementia", cond$name)) {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10'
            AND (DX LIKE 'F01%%' OR DX LIKE 'F02%%' OR DX LIKE 'F03%%'
                 OR DX LIKE 'G30%%')
        ", schema_name), cond$name)
      } else if (grepl("Hypertension", cond$name)) {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10'
            AND (DX LIKE 'I10%%' OR DX LIKE 'I11%%' OR DX LIKE 'I12%%'
                 OR DX LIKE 'I13%%' OR DX LIKE 'I14%%' OR DX LIKE 'I15%%'
                 OR DX LIKE 'I16%%')
        ", schema_name), cond$name)
      } else if (grepl("VTE", cond$name)) {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10'
            AND (DX LIKE 'I26%%' OR DX LIKE 'I82%%')
        ", schema_name), cond$name)
      } else {
        cond_res <- safe_query(sprintf("
          SELECT COUNT(DISTINCT PATID) AS n_patients
          FROM %s.DIAGNOSIS
          WHERE DX_TYPE = '10' AND DX LIKE '%s'
        ", schema_name, cond$pattern), cond$name)
      }
    } else {
      cond_res <- safe_query(sprintf("
        SELECT COUNT(DISTINCT PATID) AS n_patients
        FROM %s.DIAGNOSIS
        WHERE DX_TYPE = '10' AND DX LIKE '%s'
      ", schema_name, cond$pattern), cond$name)
    }

    if (nrow(cond_res) > 0) {
      ln("| ", cond$name, " | ", sup(cond_res$n_patients[1]), " |")
    }
  }
  ln("")

  # ================================================================
  # SECTION 15: Key medication prevalence (PRESCRIBING)
  # ================================================================
  ln("---")
  ln("## 15. Key Medication Classes — Patient Counts")
  ln("")
  ln("Based on RXNORM_CUI in PRESCRIBING table. Counts are distinct patients")
  ln("with at least one prescription record.")
  ln("")

  # Use generic names via free-text RAW_RX_MED_NAME if available,
  # otherwise we search RXNORM_CUI ranges. Since RXNORM_CUI mappings
  # vary, use RAW_RX_MED_NAME LIKE for robustness.
  med_classes <- list(
    list(name = "Apixaban",     pattern = "%apixaban%"),
    list(name = "Rivaroxaban",  pattern = "%rivaroxaban%"),
    list(name = "Dabigatran",   pattern = "%dabigatran%"),
    list(name = "Edoxaban",     pattern = "%edoxaban%"),
    list(name = "Warfarin",     pattern = "%warfarin%"),
    list(name = "Amiodarone",   pattern = "%amiodarone%"),
    list(name = "Flecainide",   pattern = "%flecainide%"),
    list(name = "Sotalol",      pattern = "%sotalol%"),
    list(name = "Dronedarone",  pattern = "%dronedarone%"),
    list(name = "Dapagliflozin", pattern = "%dapagliflozin%"),
    list(name = "Empagliflozin", pattern = "%empagliflozin%"),
    list(name = "Canagliflozin", pattern = "%canagliflozin%"),
    list(name = "Metformin",    pattern = "%metformin%"),
    list(name = "Metoprolol",   pattern = "%metoprolol%"),
    list(name = "Diltiazem",    pattern = "%diltiazem%"),
    list(name = "Digoxin",      pattern = "%digoxin%"),
    list(name = "Semaglutide",  pattern = "%semaglutide%")
  )

  # Check if RAW_RX_MED_NAME exists
  has_raw <- safe_query(sprintf("
    SELECT TOP 1 COLUMN_NAME
    FROM INFORMATION_SCHEMA.COLUMNS
    WHERE TABLE_SCHEMA = '%s' AND TABLE_NAME = 'PRESCRIBING'
      AND COLUMN_NAME = 'RAW_RX_MED_NAME'
  ", schema_name), "RAW_RX_MED_NAME check")

  med_col <- if (nrow(has_raw) > 0) "RAW_RX_MED_NAME" else "RXNORM_CUI"

  ln("Search column: **", med_col, "**")
  ln("")
  ln("| Medication | Patients | Prescriptions |")
  ln("|------------|----------|---------------|")

  for (med in med_classes) {
    med_res <- safe_query(sprintf("
      SELECT COUNT(DISTINCT PATID) AS n_patients,
             COUNT(*) AS n_rx
      FROM %s.PRESCRIBING
      WHERE %s LIKE '%s'
    ", schema_name, med_col, med$pattern), med$name)

    if (nrow(med_res) > 0) {
      ln("| ", med$name, " | ", sup(med_res$n_patients[1]), " | ",
         format(med_res$n_rx[1], big.mark = ","), " |")
    }
  }
  ln("")

  # ================================================================
  # SECTION 16: Key LOINCs for common TTE confounders
  # ================================================================
  ln("---")
  ln("## 16. Key Lab LOINCs for TTE Confounders")
  ln("")
  ln("Patient counts for specific LOINCs commonly needed as confounders or")
  ln("eligibility criteria in target trial emulations.")
  ln("")

  key_loincs <- list(
    list(loinc = "2160-0", name = "Serum creatinine"),
    list(loinc = "48642-3", name = "eGFR (CKD-EPI)"),
    list(loinc = "33914-3", name = "eGFR (MDRD)"),
    list(loinc = "4548-4",  name = "HbA1c"),
    list(loinc = "2093-3",  name = "Total cholesterol"),
    list(loinc = "2571-8",  name = "Triglycerides"),
    list(loinc = "2085-9",  name = "HDL cholesterol"),
    list(loinc = "13457-7", name = "LDL cholesterol (calc)"),
    list(loinc = "718-7",   name = "Hemoglobin"),
    list(loinc = "777-3",   name = "Platelets"),
    list(loinc = "6299-2",  name = "BUN"),
    list(loinc = "2823-3",  name = "Potassium"),
    list(loinc = "2951-2",  name = "Sodium"),
    list(loinc = "1742-6",  name = "ALT"),
    list(loinc = "1920-8",  name = "AST"),
    list(loinc = "1975-2",  name = "Total bilirubin"),
    list(loinc = "6598-7",  name = "Troponin T"),
    list(loinc = "49563-0", name = "Troponin I (HS)"),
    list(loinc = "30313-1", name = "INR"),
    list(loinc = "33762-6", name = "NT-proBNP")
  )

  ln("| LOINC | Lab Test | Patients | Results |")
  ln("|-------|----------|----------|---------|")

  for (ll in key_loincs) {
    lab_res <- safe_query(sprintf("
      SELECT COUNT(DISTINCT PATID) AS n_patients,
             COUNT(*) AS n_results
      FROM %s.LAB_RESULT_CM
      WHERE LAB_LOINC = '%s'
    ", schema_name, ll$loinc), ll$name)

    if (nrow(lab_res) > 0) {
      ln("| ", ll$loinc, " | ", ll$name, " | ",
         sup(lab_res$n_patients[1]), " | ",
         format(lab_res$n_results[1], big.mark = ","), " |")
    }
  }
  ln("")

  # ================================================================
  # Write to file
  # ================================================================
  profile_text <- paste(out, collapse = "\n")

  if (!is.null(output_file)) {
    writeLines(profile_text, output_file)
    message("Data profile written to: ", output_file)
  }

  return(invisible(profile_text))
}
```

## MPI Data Profile

The MPI database holds patient-identity / cross-system linkage data. The
function below queries ONLY non-PHI aggregate columns: row counts, the
MPI_Src system inventory, source-system ID coverage, cross-system overlap,
and deceased-flag distribution.

**PHI exclusion list** — these columns are NEVER queried:
First/Last/Middle name, SSN, DateOfBirth/DoB, address fields, phone,
email, MaidenName, AKA, any free-text identifier column.

The MPI profile is appended to the same `data_profile` file as the CDW
profile so AutoTTE sees one consolidated profile per database.

```{r fn-profile-mpi}
#| code-fold: true

profile_mpi_for_autoTTE <- function(con,
                                    schema_name = "dbo",
                                    output_file,
                                    suppress_threshold = 11) {

  safe_query <- function(sql, label = "") {
    tryCatch(
      dbGetQuery(con, sql),
      error = function(e) {
        message("  [SKIP] ", label, ": ", conditionMessage(e))
        data.frame()
      }
    )
  }

  sup <- function(n) {
    ifelse(is.na(n), NA_character_,
           ifelse(n < suppress_threshold, "<11", format(n, big.mark = ",")))
  }

  out <- character()
  ln  <- function(...) out <<- c(out, paste0(...))

  ln("")
  ln("---")
  ln("# MPI Database (linkage and source coverage)")
  ln("# Generated: ", Sys.time())
  ln("# Schema: ", schema_name)
  ln("# NOTE: This section contains ONLY aggregate counts and source-system")
  ln("#       definitions. Patient-identifying columns (name, SSN, DOB,")
  ln("#       address, phone) are NEVER queried. Counts < ",
     suppress_threshold, " are suppressed.")
  ln("")

  # ---------------- M1. Row counts per MPI table ----------------
  ln("## M1. MPI Table Row Counts")
  ln("")
  ln("| Table | Rows |")
  ln("|-------|------|")
  for (tbl in c("EnterpriseRecords", "EnterpriseRecords_Ext", "Mpi", "MPI_Src")) {
    rc <- safe_query(sprintf(
      "SELECT COUNT(*) AS n FROM %s.%s", schema_name, tbl), tbl)
    if (nrow(rc) > 0 && !is.na(rc$n[1])) {
      ln("| ", tbl, " | ", format(rc$n[1], big.mark = ","), " |")
    }
  }
  ln("")

  # ---------------- M2. Source-system inventory (MPI_Src) ----------------
  # MPI_Src contains source-system *definitions*, not patient data.
  ln("## M2. Source-System Inventory (MPI_Src)")
  ln("")
  ln("Full dump of MPI_Src — this table holds source-system definitions,")
  ln("not patient records.")
  ln("")
  src <- safe_query(sprintf("SELECT * FROM %s.MPI_Src", schema_name), "MPI_Src")
  if (nrow(src) > 0) {
    ln("```")
    ln(paste(capture.output(print(src, row.names = FALSE)), collapse = "\n"))
    ln("```")
    ln("")
  }

  # ---------------- M3. Distinct Uid count ----------------
  ln("## M3. Distinct Patients (Uid)")
  ln("")
  uid <- safe_query(sprintf(
    "SELECT COUNT(DISTINCT Uid) AS n FROM %s.EnterpriseRecords", schema_name),
    "Uid count")
  if (nrow(uid) > 0 && !is.na(uid$n[1])) {
    ln("Distinct Uids in EnterpriseRecords: ", sup(uid$n[1]))
    ln("")
  }

  # ---------------- M4. Source-system ID coverage ----------------
  # Per-source-system non-NULL rate on EnterpriseRecords_Ext. Reports only
  # counts and percentages — the ID values themselves are never selected.
  ln("## M4. Source-System ID Coverage (EnterpriseRecords_Ext)")
  ln("")
  ln("Percentage of EnterpriseRecords_Ext rows with a non-NULL source ID.")
  ln("")

  source_id_cols <- c("EPIC_PAT_ID", "ALLSCRIPTS_PERSON_ID",
                      "MHH_MRN", "UTP_MRN")

  total_ext <- safe_query(sprintf(
    "SELECT COUNT(*) AS n FROM %s.EnterpriseRecords_Ext", schema_name),
    "ER_Ext total")

  if (nrow(total_ext) > 0 && total_ext$n[1] > 0) {
    ln("Total rows: ", format(total_ext$n[1], big.mark = ","))
    ln("")
    ln("| Source ID Column | Non-NULL Rows | % Coverage |")
    ln("|------------------|---------------|------------|")
    for (col in source_id_cols) {
      r <- safe_query(sprintf("
        SELECT COUNT(*) AS n
        FROM %s.EnterpriseRecords_Ext
        WHERE %s IS NOT NULL
      ", schema_name, col), paste0("ER_Ext.", col))
      if (nrow(r) > 0 && !is.na(r$n[1])) {
        pct <- round(100 * r$n[1] / total_ext$n[1], 1)
        ln("| ", col, " | ", sup(r$n[1]), " | ", pct, "% |")
      }
    }
    ln("")
  }

  # ---------------- M5. Cross-system overlap (EPIC vs ALLSCRIPTS) ----------------
  # 4-cell crosstab: who has EPIC only, ALLSCRIPTS only, both, neither.
  # No IDs leave the DB — only the four counts.
  ln("## M5. EPIC × ALLSCRIPTS Overlap")
  ln("")
  ln("Counts of EnterpriseRecords_Ext patients by which legacy/current EHR")
  ln("system they appear in. Drives expectations for legacy-encounter")
  ln("filtering in CDW queries.")
  ln("")

  overlap <- safe_query(sprintf("
    SELECT
      SUM(CASE WHEN EPIC_PAT_ID IS NOT NULL
                AND ALLSCRIPTS_PERSON_ID IS NOT NULL THEN 1 ELSE 0 END) AS both_,
      SUM(CASE WHEN EPIC_PAT_ID IS NOT NULL
                AND ALLSCRIPTS_PERSON_ID IS NULL THEN 1 ELSE 0 END) AS epic_only,
      SUM(CASE WHEN EPIC_PAT_ID IS NULL
                AND ALLSCRIPTS_PERSON_ID IS NOT NULL THEN 1 ELSE 0 END) AS allscripts_only,
      SUM(CASE WHEN EPIC_PAT_ID IS NULL
                AND ALLSCRIPTS_PERSON_ID IS NULL THEN 1 ELSE 0 END) AS neither
    FROM %s.EnterpriseRecords_Ext
  ", schema_name), "EPIC/ALLSCRIPTS overlap")

  if (nrow(overlap) > 0) {
    ln("| Category | Patients |")
    ln("|----------|----------|")
    ln("| Has EPIC and ALLSCRIPTS  | ", sup(overlap$both_[1]), " |")
    ln("| Has EPIC only            | ", sup(overlap$epic_only[1]), " |")
    ln("| Has ALLSCRIPTS only      | ", sup(overlap$allscripts_only[1]), " |")
    ln("| Has neither              | ", sup(overlap$neither[1]), " |")
    ln("")
  }

  # ---------------- M6. Deceased flag distribution ----------------
  # Only the boolean/categorical deceased flag is queried — no death dates.
  deceased <- safe_query(sprintf("
    SELECT ISNULL(CAST(Deceased AS varchar(20)), '(NULL)') AS deceased_val,
           COUNT(*) AS n
    FROM %s.EnterpriseRecords_Ext
    GROUP BY Deceased
    ORDER BY n DESC
  ", schema_name), "Deceased dist")

  if (nrow(deceased) > 0) {
    ln("## M6. Deceased Flag Distribution")
    ln("")
    ln("| Deceased | Patients |")
    ln("|----------|----------|")
    for (i in seq_len(nrow(deceased))) {
      ln("| ", deceased$deceased_val[i], " | ", sup(deceased$n[i]), " |")
    }
    ln("")
  }

  # Append (do not overwrite) so the CDW profile written earlier is preserved.
  profile_text <- paste(out, collapse = "\n")
  cat(profile_text, file = output_file, append = TRUE)
  message("MPI profile appended to: ", output_file)
  invisible(profile_text)
}
```

## Run everything

One unified run section. Each artifact opens its own short-lived connection
through `with_db_connection()` so a failure in any step still closes its
ODBC handle. Re-connecting between steps is cheap relative to the multi-hour
profile queries.

```{r run-cdw-schema}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  extract_schema_for_claude(con,
                            schema_name = cdw_schema_name,
                            output_file = paths$schema)
})
```

```{r run-cdw-profile}
#| code-fold: true
with_db_connection(cfg$connection$r_code, function(con) {
  profile_cdw_for_autoTTE(con,
                          schema_name = cdw_schema_name,
                          output_file = paths$data_profile)
})
```

```{r run-mpi-schema}
#| code-fold: true
with_db_connection(cfg$mpi_connection$r_code, function(con) {
  extract_schema_for_claude(con,
                            schema_name = "dbo",
                            output_file = paths$mpi_schema)
})
```

```{r run-staging-schemas}
#| code-fold: true
# Staging-database schema dumps (optional). For secure_pcornet_cdw this
# covers cdwstaging.gecbi, cdwstaging.allscripts, and CDWStaging.allscripts
# — the tables the duplicate-records pattern joins on MRN == mpi.lid.
# Each entry opens its own short-lived connection via its own USE <db>
# block so a DSN/permission problem on one staging DB doesn't abort the
# others. Entries whose DB / schema is unreachable are skipped with a
# [SKIP] log line and the downstream get_schema() call still works — the
# missing dumps are just absent from the concatenated result.
staging_entries <- cfg$staging_schema_dumps %||% list()
for (entry in staging_entries) {
  label   <- entry$label %||% "(unlabeled staging)"
  sn      <- entry$schema_name %||% "dbo"
  outpath <- entry$schema_dump_resolved
  r_code  <- entry$connection$r_code
  if (is.null(outpath) || !nzchar(outpath)) {
    message("[SKIP] staging dump ", label, ": no schema_dump path")
    next
  }
  if (is.null(r_code) || !nzchar(r_code)) {
    message("[SKIP] staging dump ", label, ": no connection.r_code")
    next
  }
  tryCatch(
    with_db_connection(r_code, function(con) {
      extract_schema_for_claude(con,
                                schema_name = sn,
                                output_file = outpath)
      message("Staging schema written [", label, " / ", sn, "] -> ", outpath)
    }),
    error = function(e) {
      message("[SKIP] staging dump ", label, " / ", sn, ": ",
              conditionMessage(e))
    }
  )
}
```

```{r run-mpi-profile}
#| code-fold: true
with_db_connection(cfg$mpi_connection$r_code, function(con) {
  profile_mpi_for_autoTTE(con,
                          schema_name = "dbo",
                          output_file = paths$data_profile)
})
```

```{r done-banner}
#| echo: false
message("==============================================================")
message("CDW schema written to:   ", paths$schema)
message("CDW data profile written to: ", paths$data_profile)
message("  (MPI section appended to the same file)")
message("MPI schema written to:   ", paths$mpi_schema)
message("==============================================================")
```