---
title: "CDW DX code storage format probe"
subtitle: "Are ICD-10 codes stored dotted, dotless, or mixed?"
format:
  html:
    toc: true
    code-fold: show
    embed-resources: true
execute:
  warning: false
  message: false
---

## Purpose

Determine how this CDW stores the `DX` column for ICD-10 codes. The PCORnet
CDM spec says dotless (`N1830`), but implementations vary depending on the
source feed. Analyses that emit exact-match IN clauses (or LIKE prefixes
with interior periods) silently match zero rows if they guess the wrong
format — a cohort of 0 looks like "no data for that question" rather than
a bug.

This notebook runs five quick probes and prints a verdict.

## Setup

```{r}
#| label: setup
library(DBI)
library(odbc)
con <- dbConnect(odbc::odbc(), "SQLODBCD17CDM")
stopifnot(DBI::dbIsValid(con))
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Total ICD-10 row count (denominator)

```{r}
#| label: row-count
DBI::dbGetQuery(con, "
  SELECT COUNT(*) AS icd10_rows
  FROM CDW.dbo.DIAGNOSIS
  WHERE DX_TYPE = '10'
")
```

## 2. Sample 20 ICD-10 codes (eyeball format)

```{r}
#| label: sample
DBI::dbGetQuery(con, "
  SELECT TOP 20 DX
  FROM CDW.dbo.DIAGNOSIS
  WHERE DX_TYPE = '10'
    AND DX IS NOT NULL
  ORDER BY NEWID()
")
```

## 3. Split by "has-period" across all ICD-10 rows

```{r}
#| label: period-split
split_all <- DBI::dbGetQuery(con, "
  SELECT
    SUM(CASE WHEN DX LIKE '%.%' THEN 1 ELSE 0 END) AS with_period,
    SUM(CASE WHEN DX LIKE '%.%' THEN 0 ELSE 1 END) AS without_period,
    COUNT(*)                                      AS total
  FROM CDW.dbo.DIAGNOSIS
  WHERE DX_TYPE = '10'
    AND DX IS NOT NULL
")
split_all
```

```{r}
#| label: verdict
total <- split_all$total
pct_dotted   <- round(100 * split_all$with_period    / total, 2)
pct_dotless  <- round(100 * split_all$without_period / total, 2)
cat(sprintf(
  "Dotted   (e.g. 'N18.30') : %s rows  (%.2f%%)\n",
  format(split_all$with_period, big.mark = ","), pct_dotted
))
cat(sprintf(
  "Dotless  (e.g. 'N1830' or 'E11')  : %s rows  (%.2f%%)\n",
  format(split_all$without_period, big.mark = ","), pct_dotless
))
cat("\n")
if (pct_dotless > 95) {
  cat("VERDICT: DOTLESS storage. Use IN ('N1830','N1831',...) patterns.\n")
} else if (pct_dotted > 95) {
  cat("VERDICT: DOTTED storage. Use IN ('N18.30','N18.31',...) patterns.\n")
} else {
  cat("VERDICT: MIXED storage. Emit BOTH dotted and dotless forms in every\n",
      "         IN list and in any LIKE prefix that contains a period.\n", sep = "")
}
```

Note: a three-character root like `E11` (diabetes mellitus) has no period
even in the "dotted" convention, which is why the verdict uses a 95%
threshold rather than 0 vs 100 split.

## 4. Split by "has-period" after excluding short (3-char) root codes

Three-character codes like `I10`, `E11`, `J18` never have periods in either
convention. Excluding them isolates codes that *would* have a period if the
CDW were dotted.

```{r}
#| label: period-split-long
DBI::dbGetQuery(con, "
  SELECT
    SUM(CASE WHEN DX LIKE '%.%' THEN 1 ELSE 0 END) AS with_period,
    SUM(CASE WHEN DX LIKE '%.%' THEN 0 ELSE 1 END) AS without_period,
    COUNT(*)                                      AS total
  FROM CDW.dbo.DIAGNOSIS
  WHERE DX_TYPE = '10'
    AND DX IS NOT NULL
    AND LEN(DX) >= 4
")
```

If this table shows near-100% `with_period` for rows with length ≥ 4, the
CDW is dotted. Near-100% `without_period` means dotless.

## 5. Targeted probe — CKD and dementia codes

Spot-check the exact codes Protocol 03 and Protocol 01 care about:

```{r}
#| label: targeted-probe
targeted <- DBI::dbGetQuery(con, "
  SELECT DX, COUNT(*) AS n
  FROM CDW.dbo.DIAGNOSIS
  WHERE DX_TYPE = '10'
    AND (
      DX LIKE 'N18%'
      OR DX LIKE 'G30%'
      OR DX LIKE 'G31%'
      OR DX LIKE 'Z99%'
      OR DX LIKE 'S72%'
    )
  GROUP BY DX
  HAVING COUNT(*) > 50
  ORDER BY n DESC
")
targeted
```

Scan the `DX` column of this table. You'll see exact strings like `N18.30`
vs `N1830` and can confirm the format at a glance.

## Cleanup

```{r}
#| label: cleanup
try(DBI::dbDisconnect(con), silent = TRUE)
```

## Next steps

- **Dotless only?** Future protocols should emit dotless codes directly.
  The `expand_icd_codes()` helper in the flu-ADRD protocols already
  handles this transparently — no action needed.
- **Dotted only?** Unusual but valid. Update `databases/conventions/secure_pcornet_cdw_conventions.md`
  to record this so future AutoTTE runs emit dotted codes by default.
- **Mixed?** Keep the both-forms expansion (current pattern). Document it
  in the conventions file as a warning.

Add the verdict from section 3 to
`databases/conventions/secure_pcornet_cdw_conventions.md` under "ICD-9/10
Transition" so the next protocol worker agent doesn't have to re-derive
the format.
